### 2021-02-12 documents/problemformulering.tex

\documentclass[]{article}
\usepackage[utf8]{inputenc}
%\usepackage[danish, english]{babel}
\usepackage[danish]{babel}
\usepackage{graphicx}
%\usepackage[intoc]{nomencl} % Nomenclature package
\usepackage[usenames]{xcolor}
\usepackage{times}
\usepackage{listings}
\usepackage{textcomp}
\usepackage{hyperref}
\usepackage{textpos}
%\usepackage[showboxes]{textpos}
\usepackage{indentfirst}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{ifthen}
\usepackage{float}
%\usepackage{mathtools}
\usepackage{pict2e}
%\usepackage{steinmetz}
\usepackage[export]{adjustbox}
\pgfplotsset{compat=1.13}
\usepackage{siunitx}
\usepackage{tabularx}
%
% adjustment to page format
\marginparwidth=0pt
\oddsidemargin=0pt
%\evensidemargin=0pt
\marginparsep=0pt
%\topmargin=-15 mm
\textwidth=168 mm
\textheight=230 mm
\parskip 0.27 em
\parindent=0pt
%\leftmargin=1cm

% Define DTU color
\definecolor{dtured}{RGB}{153,0,0}

\begin{document}
% formel header
\thispagestyle{empty}
\vspace*{-1.9cm}
\begin{textblock*}{18cm}[0,0](0cm,-2cm) %
  \noindent
  \includegraphics[width=4.5cm,valign=t]{tex_dtu_elektro_a}
  \hspace*{10.6cm}
  \includegraphics[width=1.5cm,valign=t]{tex_dtu_logo}
\end{textblock*}
\begin{textblock*}{19.0cm}(0cm,-2.5cm) %
\begin{center}
  {\color{dtured}\large31015 Fagprojekt}\\
  \normalsize
  s186083 Tjark Petersen\\
  s194006 Steffan Martin Kunoy\\
  s194027 Victor Alexander Hansen\\
\end{center}
\end{textblock*}
% rød streg
\vspace{1mm}
{\hspace*{-0.1cm}
\color{dtured}\noindent \rule{16.8cm}{5pt}}

 \section*{Identifikation}

 \begin{table}[H]
     \centering
     \begin{tabularx}{\textwidth}{|X|X|X|}
     \hline
          DTU Elektro&Forår 2021 & Gruppe: 7, ID: MEK2 \\\hline
          Kursus 31015 & Titel & Gruppemedlemmer \\\hline
          Fagprojekt - Elektroteknologi & ROAST Telemetri & \begin{tabular}{l} s186083 Tjark Petersen\\s194006 Steffan Martin Kunoy\\s194027 Victor Alexander Hansen \end{tabular}\\\hline
          Dokument:& Projektformulering & 1 side\\\hline 
          Version/Status: & 1. udgave &\today\\\hline
     \end{tabularx}
 \end{table}

 \section*{Introduktion til emne/problemstilling}

Dette projekt vil omhandle et telemetri modul som skal indgå i DTU Roadrunners Solar Team (ROAST). I 2023 deltager ROAST i Bridgestone World Solar Challenge, som er et løb på over 3000 km tværs gennem Australien. I løbet vil solbilen være ledsaget af en følgebil, som skal overvåge solbilens driftstatus og sende kommandoer til solbilen.

Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i en "black box". Et lignende modul skal også være til stede i følgebilen, så følgebilen omvendt kan reagere på denne data, enten automatisk eller ved menneskelig ageren, og sende kommandoer tilbage til solbilens CAN bus. Telemetri modulet spiller derfor en afgørende rolle, da man også skal tage højde for at afstanden mellem solbil og følgebil kan være op til 1 km alt afhængig af trafikforholdene.

\section*{Problemformulering}
Fjernmåling af data fra solbilen vil give ROAST en fordel i og med at support bilen kan indtage en mere aktiv rolle i styring og optimering af bilen. Det letter derfor på kørerens arbejdsbyrde, og fordeler ansvaret for hele ROAST-support holdet. Af denne grund vil vi afgrænse projektet på følgende punkter: 
\begin{itemize}
    \item Hvordan kan vi lave et modul der kan læse, sende og modtage sensor data og kommandoer i CAN bus netværket?
    \item Hvordan kan CAN data overføres trådløst i en afstand på en kilometer?
    \item Hvordan skal data behandles ved modtagelse i følgebilen?
    \item Hvordan kan data bedst gemmes lokalt (black box) og udlæses på et senere tidspunkt?
\end{itemize}

\end{document}

### 2021-02-25 documentation/projektplan/problemformulering.tex
Updates from Overleaf

--- 2021-02-12 documents/problemformulering.tex
@@
 \usepackage{textpos}
+\usepackage{pgfgantt}
+\usepackage{lscape}
 %\usepackage[showboxes]{textpos}
@@
 \definecolor{dtured}{RGB}{153,0,0}
-
+\title{Projektplan}
 \begin{document}
@@
   \noindent
-  \includegraphics[width=4.5cm,valign=t]{tex_dtu_elektro_a}
+  \includegraphics[width=4.5cm,valign=t]{documentation/resources/tex_dtu_elektro_a.pdf}
   \hspace*{10.6cm}
-  \includegraphics[width=1.5cm,valign=t]{tex_dtu_logo}
+  \includegraphics[width=1.5cm,valign=t]{documentation/resources/tex_dtu_logo.pdf}
 \end{textblock*}
@@
 
- \section*{Identifikation}
+\section{Identifikation}
 
@@
 
- \section*{Introduktion til emne/problemstilling}
+\section{Introduktion til emne/problemstilling}
 
@@
 
-\section*{Problemformulering}
+\section{Problemformulering}
 Fjernmåling af data fra solbilen vil give ROAST en fordel i og med at support bilen kan indtage en mere aktiv rolle i styring og optimering af bilen.
@@
 \begin{itemize}
+    \item Hvordan kan CAN data overføres trådløst i en afstand på en kilometer?
     \item Hvordan kan vi lave et modul der kan læse, sende og modtage sensor data og kommandoer i CAN bus netværket?
-    \item Hvordan kan CAN data overføres trådløst i en afstand på en kilometer?
     \item Hvordan skal data behandles ved modtagelse i følgebilen?
@@
 
+\section{Afgrænsning}
+Vi har i vores gruppe diskuteret flere forskellige umiddelbare løsningsformer for projektets produkt, heriblandt at lave en løsning på et PCB board, eller en løsning på et Teensy board/FPGA board.
+Vi har valgt at afgrænse os til at se på Teensy og FPGA løsninger, og ikke PCB boards.
+Dette skyldes at det hardware design vi kommer frem til som produkt kan redigeres mere fleksibelt undervejs på et Teensy og FPGA board, fremfor et PCB board som er mere special designet til formålet.\\
+Det skal dog siges at vi forventer at et PCB board design ville være mindre strømkrævende end de løsninger vi ser på, men vi har i dette projekt valgt at prioritere et godt design over strømbesparelse.
+\section{Målgruppe}
+Målgruppen for vores projekt er DTU ROAST holdet, som skal integrere telemetrimodulet i deres solbil. Efter planen skal bilen så deltage ved Bridgestone World Solar Challenge i 2023.
+\section{Løsningsmetode}
+\begin{enumerate}
+    \item Foreløbigt design af telemetrimodulet som blokdiagram. Fastslå specifikationer og krav til løsningen. 
+    \item Valg af komponenter til det indledende design. Bestilling af komponenter som ikke kan fremskaffes på DTU. 
+    \item Første prototype af telemetrimodul på breadboard. Skrive kode til at teste begrænset funktionalitet. 
+    \item Foreslå forbedringer til første prototype. evt. bestille nye komponenter. Optimer kode og design mere omfattende tests.
+    \item Anden prototype af telemetrimodul på breadboard. Gentag 3. og 4. indtil grundlæggende krav i 1. er opfyldt. 
+    \item Udvikle mere færdig løsning, dvs. kompakt, loddet og programmerbar. Gennemfør tests som i 5. 
+    \item Design af GUI til interaktion med PC. Test med USB forbindelse til support-bil modul. Styring via PC.  
+    \item Præsenter løsning for DTU ROAST holdet. Diskuter forbedringer og features. 
+    \item Integration med solbil, løbende tests og evaluering af projekt. 
+\end{enumerate}
+\section{Ressourcer}
+\begin{itemize}
+    \item Teensy 4.0/4.1 development board
+    \item RF transciever module
+    \item SD kort
+    \item MCP2515 CAN Module (Transceiver & Controller)
+\end{itemize}
+\section{Aktivitetsplan}
+
+\begin{ganttchart}{1}{20}
+    \gantttitle{Roast Telemetry}{20} \\
+    \gantttitlelist{9,10,11,12,14,15,16,17,18,19}{2} \\
+    \ganttgroup{Solar Car Unit}{1}{20} \\
+    \ganttbar{Task 1}{1}{2} \\
+    \ganttlinkedbar{Task 2}{3}{7} \ganttnewline
+    \ganttmilestone{Milestone}{7} \ganttnewline
+    \ganttbar{Final Task}{8}{12}
+    \ganttlink{elem2}{elem3}\\
+    \ganttgroup{Support Vehicle Unit}{1}{20} \\
+\end{ganttchart}
+\section{Referenceliste}
+\section{Grøn dyst?}
 \end{document}

### 2021-02-26 documentation/projektplan/problemformulering.tex

@@
 I 2023 deltager ROAST i Bridgestone World Solar Challenge, som er et løb på over 3000 km tværs gennem Australien.
+Solbilen skal køre en distance på 3 x 1200km, hvor bilen ikke har mulighed for at oplade.
+Den primære energikilde for bilen under løbet er derfor solenergi.\\
 I løbet vil solbilen være ledsaget af en følgebil, som skal overvåge solbilens driftstatus og sende kommandoer til solbilen.
+Dette er en yderst vigtig opgave, da bilerne skal køre tværs gennem Australien i offentlig trafik, hvor alt fra varme til dæktryk udfordrer solbilen.
+Følgebilen vil derfor have mulighed for at analysere og reagere på data fra solbilen, hvis køreren i solbilen er optaget af trafik og betjening af solbilen.
 
-Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i en "black box".
+Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i en "black box".
 Et lignende modul skal også være til stede i følgebilen, så følgebilen omvendt kan reagere på denne data, enten automatisk eller ved menneskelig ageren, og sende kommandoer tilbage til solbilens CAN bus.
@@
 Dette skyldes at det hardware design vi kommer frem til som produkt kan redigeres mere fleksibelt undervejs på et Teensy og FPGA board, fremfor et PCB board som er mere special designet til formålet.\\
-Det skal dog siges at vi forventer at et PCB board design ville være mindre strømkrævende end de løsninger vi ser på, men vi har i dette projekt valgt at prioritere et godt design over strømbesparelse.
+Det skal dog siges at vi forventer at et PCB board design ville være mindre strømkrævende end de løsninger vi ser på, men vi har i dette projekt valgt at prioritere et godt design over strømbesparelse, siden der formentlig er tale om forskel i størrelsesorden af nogen tiende af mA.
 \section{Målgruppe}
@@
     \gantttitlelist{9,10,11,12,14,15,16,17,18,19}{2} \\
+    \ganttbar{Møder}{1}{20} \\
+    \ganttbar{Rap. skrivning}{7}{8}
+    \ganttbar{}{18}{20} \ganttnewline
+    % Hvordan fjernes det mærkelige link på figur?
     \ganttgroup{Solar Car Unit}{1}{20} \\
-    \ganttbar{Task 1}{1}{2} \\
-    \ganttlinkedbar{Task 2}{3}{7} \ganttnewline
-    \ganttmilestone{Milestone}{7} \ganttnewline
-    \ganttbar{Final Task}{8}{12}
+    \ganttbar{Design af telemetrimodul}{1}{2} \\
+    \ganttlinkedbar{Prototype udvikling}{3}{7} \ganttnewline
+    \ganttmilestone{Krav opfyldt}{7} \ganttnewline
+    \ganttbar{Produkt udvikling}{8}{12}\\
+    \ganttbar{GUI}{8}{12}
     \ganttlink{elem2}{elem3}\\
@@
 \section{Referenceliste}
+https://worldsolarchallenge.org/
 \section{Grøn dyst?}

### 2021-02-26 documentation/projektplan/problemformulering.tex

@@
 
-\section{Identifikation}
-
 \begin{table}[H]
@@
 Dette projekt vil omhandle et telemetri modul som skal indgå i DTU Roadrunners Solar Team (ROAST).
-I 2023 deltager ROAST i Bridgestone World Solar Challenge, som er et løb på over 3000 km tværs gennem Australien.
+I 2023 deltager ROAST i Bridgestone World Solar Challenge, som er et løb på over 3000 km tværs gennem Australien \cite{wsc}.
 Solbilen skal køre en distance på 3 x 1200km, hvor bilen ikke har mulighed for at oplade.
@@
 
-Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i en "black box".
+Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i et "black box" hukommelseschip.
 Et lignende modul skal også være til stede i følgebilen, så følgebilen omvendt kan reagere på denne data, enten automatisk eller ved menneskelig ageren, og sende kommandoer tilbage til solbilens CAN bus.
@@
 \begin{itemize}
-    \item Teensy 4.0/4.1 development board
-    \item RF transciever module
-    \item SD kort
+    \item 2x Teensy 4.0/4.1 development board
+    \item 2x RF transciever module
+    \item SD kort modul
     \item MCP2515 CAN Module (Transceiver & Controller)
 \end{itemize}
+\begin{landscape}
 \section{Aktivitetsplan}
-
-\begin{ganttchart}{1}{20}
-    \gantttitle{Roast Telemetry}{20} \\
-    \gantttitlelist{9,10,11,12,14,15,16,17,18,19}{2} \\
-    \ganttbar{Møder}{1}{20} \\
-    \ganttbar{Rap. skrivning}{7}{8}
-    \ganttbar{}{18}{20} \ganttnewline
-    % Hvordan fjernes det mærkelige link på figur?
-    \ganttgroup{Solar Car Unit}{1}{20} \\
-    \ganttbar{Design af telemetrimodul}{1}{2} \\
-    \ganttlinkedbar{Prototype udvikling}{3}{7} \ganttnewline
-    \ganttmilestone{Krav opfyldt}{7} \ganttnewline
-    \ganttbar{Produkt udvikling}{8}{12}\\
-    \ganttbar{GUI}{8}{12}
-    \ganttlink{elem2}{elem3}\\
-    \ganttgroup{Support Vehicle Unit}{1}{20} \\
-\end{ganttchart}
+\hspace{1.5cm}\includegraphics[width=1.45\textwidth]{documentation/images/projectPlan.pdf}
+\end{landscape}
 \section{Referenceliste}
-https://worldsolarchallenge.org/
+\begin{thebibliography}{9}
+\bibitem{wsc}
+Bridgestone World Solar Challenge, 
+URL: https://worldsolarchallenge.org/
+\end{thebibliography}
 \section{Grøn dyst?}

### 2021-02-26 documentation/projektplan/problemformulering.tex
Updates from Overleaf
update shopping list

@@
 \usepackage[utf8]{inputenc}
-%\usepackage[danish, english]{babel}
-\usepackage[danish]{babel}
+\usepackage[danish,english]{babel}
 \usepackage{graphicx}
@@
 
-\section{Introduktion til emne/problemstilling}
+\section{Problem introduction}
 
-Dette projekt vil omhandle et telemetri modul som skal indgå i DTU Roadrunners Solar Team (ROAST).
+This project will concern a telemetry module for use by the DTU Roadrunners Solar Team (ROAST).
+In 2023 ROAST is set to participate in the Bridgestone World Solar Challenge, a race spanning 3000 km across Australia \cite{wsc}.
+The solar car must be able to drive 3 stretches of 1200 km without recharging.
+Thus, solar power will be the main energy source during the race.
+\\
+%Dette projekt vil omhandle et telemetri modul som skal indgå i DTU Roadrunners Solar Team (ROAST).
 I 2023 deltager ROAST i Bridgestone World Solar Challenge, som er et løb på over 3000 km tværs gennem Australien \cite{wsc}.
@@
 Den primære energikilde for bilen under løbet er derfor solenergi.\\
-I løbet vil solbilen være ledsaget af en følgebil, som skal overvåge solbilens driftstatus og sende kommandoer til solbilen.
+Throughout the race the solar car will be monitored by a support car, which can monitor the car's conditions and issue commands.
+This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
+The support car will therefore be able to analyze and react to the data stream being transmitted from the solar car, even when the driver is preoccupied by driving the car and navigating traffic.
+\\
+%I løbet vil solbilen være ledsaget af en følgebil, som skal overvåge solbilens driftstatus og sende kommandoer til solbilen.
 Dette er en yderst vigtig opgave, da bilerne skal køre tværs gennem Australien i offentlig trafik, hvor alt fra varme til dæktryk udfordrer solbilen.
 Følgebilen vil derfor have mulighed for at analysere og reagere på data fra solbilen, hvis køreren i solbilen er optaget af trafik og betjening af solbilen.
-
 Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i et "black box" hukommelseschip.
@@
 
-\section{Problemformulering}
-Fjernmåling af data fra solbilen vil give ROAST en fordel i og med at support bilen kan indtage en mere aktiv rolle i styring og optimering af bilen.
+\section{Problem description}
+Remote sensing of data from the solar car will give the ROAST team a competitive advantage as the support vehicle will take on a more active role in controlling and optimising the car's performance.
+This will ease the burden on the driver and engage a large part of the supporting crew.
+For these reasons we have chosen to focus our project on the following points:
+%Fjernmåling af data fra solbilen vil give ROAST en fordel i og med at support bilen kan indtage en mere aktiv rolle i styring og optimering af bilen.
 Det letter derfor på kørerens arbejdsbyrde, og fordeler ansvaret for hele ROAST-support holdet.
@@
 \begin{itemize}
-    \item Hvordan kan CAN data overføres trådløst i en afstand på en kilometer?
-    \item Hvordan kan vi lave et modul der kan læse, sende og modtage sensor data og kommandoer i CAN bus netværket?
-    \item Hvordan skal data behandles ved modtagelse i følgebilen?
-    \item Hvordan kan data bedst gemmes lokalt (black box) og udlæses på et senere tidspunkt?
+    \item How can we design a module that can read and write sensor data and commands from a CAN bus network? 
+    %\item Hvordan kan vi lave et modul der kan læse, sende og modtage sensor data og kommandoer i CAN bus netværket?
+    \item How can CAN data be transmitted over a distance of up to one kilometer? 
+    %\item Hvordan kan CAN data overføres trådløst i en afstand på en kilometer?
+    \item How does the support-car module process data upon receiving it from the solar car?
+    %\item Hvordan skal data behandles ved modtagelse i følgebilen?
+    \item What is the best method for storing data locally (black box) while being accesible at a later time? 
+    %\item Hvordan kan data bedst gemmes lokalt (black box) og udlæses på et senere tidspunkt?
 \end{itemize}
@@
 Målgruppen for vores projekt er DTU ROAST holdet, som skal integrere telemetrimodulet i deres solbil. Efter planen skal bilen så deltage ved Bridgestone World Solar Challenge i 2023.
-\section{Løsningsmetode}
+
+\section{Problem solution}
 \begin{enumerate}
-    \item Foreløbigt design af telemetrimodulet som blokdiagram. Fastslå specifikationer og krav til løsningen. 
-    \item Valg af komponenter til det indledende design. Bestilling af komponenter som ikke kan fremskaffes på DTU. 
-    \item Første prototype af telemetrimodul på breadboard. Skrive kode til at teste begrænset funktionalitet. 
-    \item Foreslå forbedringer til første prototype. evt. bestille nye komponenter. Optimer kode og design mere omfattende tests.
-    \item Anden prototype af telemetrimodul på breadboard. Gentag 3. og 4. indtil grundlæggende krav i 1. er opfyldt. 
-    \item Udvikle mere færdig løsning, dvs. kompakt, loddet og programmerbar. Gennemfør tests som i 5. 
-    \item Design af GUI til interaktion med PC. Test med USB forbindelse til support-bil modul. Styring via PC.  
-    \item Præsenter løsning for DTU ROAST holdet. Diskuter forbedringer og features. 
-    \item Integration med solbil, løbende tests og evaluering af projekt. 
+    \item Preliminary design of telemetry module in block diagram form. Establish specifications and requirements for the final product. 
+    %\item Foreløbigt design af telemetrimodulet som blokdiagram. Fastslå specifikationer og krav til løsningen. 
+    \item Select components for preliminary design. Ordering components not available from DTU. 
+    %\item Valg af komponenter til det indledende design. Bestilling af komponenter som ikke kan fremskaffes på DTU. 
+    \item Prototyping telemetry module on two breadboards. Write code to test limited functionality. 
+    %\item Første prototype af telemetrimodul på breadboard. Skrive kode til at teste begrænset funktionalitet. 
+    \item Improve on prototype. Order new components if needed. Optimize code and write more extensive tests. 
+    %\item Foreslå forbedringer til første prototype. evt. bestille nye komponenter. Optimer kode og design mere omfattende tests.
+    \item Improved prototype of telemetry module on breadboard. Repeat Steps 3. and 4. until fundamental requirements in Step 1. are satisfied. 
+    %\item Anden prototype af telemetrimodul på breadboard. Gentag 3. og 4. indtil grundlæggende krav i 1. er opfyldt. 
+    \item Develop more permanent solution, i.e. compact, soldered and configurable. Complete tests in Step 5.
+    %\item Udvikle mere færdig løsning, dvs. kompakt, loddet og programmerbar. Gennemfør tests som i 5.
+    \item Design GUI for interaction with PC. Test with USB connection to support-car module. Remote control via PC. 
+    %\item Design af GUI til interaktion med PC. Test med USB forbindelse til support-bil modul. Styring via PC.  
+    \item Present solution to DTU ROAST team. Discuss improvements and additional features. 
+    %\item Præsenter løsning for DTU ROAST holdet. Diskuter forbedringer og features. 
+    \item Integrate telemetry module with solar car. Ongoing testing of solution. Feedback and evaluate project. 
+    %\item Integration med solbil, løbende tests og evaluering af projekt. 
 \end{enumerate}
-\section{Ressourcer}
+\section{Resources}
 \begin{itemize}
-    \item 2x Teensy 4.0/4.1 development board
-    \item 2x RF transciever module
-    \item SD kort modul
-    \item MCP2515 CAN Module (Transceiver & Controller)
+    \item 1x Teensy 4.0 development board
+    \item 1x Teensy 4.1 development board
+    \item 2x nRF24L01 + PA/LNA Wireless Transceiver with antenna
+    \item 1x SD memory card
+    \item 2x MCP2551 CAN Module (Transceiver & Controller) m. breakout board
 \end{itemize}
@@
 \end{landscape}
+Meetings:\\
+We will have a weekly meeting with both the project coordinator Christian Kampp Kruuse and our advisor Martin Schoeberl at Friday.
+Report writing:\\
+
 \section{Referenceliste}

### 2021-02-27 documentation/projektplan/problemformulering.tex
Updates from Overleaf

@@
 \usepackage{tabularx}
-%
+
 % adjustment to page format
@@
 %\leftmargin=1cm
-
+\pagenumbering{arabic}
 % Define DTU color
@@
 Den primære energikilde for bilen under løbet er derfor solenergi.\\
+\\
 Throughout the race the solar car will be monitored by a support car, which can monitor the car's conditions and issue commands.
@@
 Følgebilen vil derfor have mulighed for at analysere og reagere på data fra solbilen, hvis køreren i solbilen er optaget af trafik og betjening af solbilen.
-Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i et "black box" hukommelseschip.
+\\
+The main function of the telemetry module is to read data from the solar car's onboard CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support car, while simultaneosly logging data locally in a "black box" memory chip.
+The support car is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus.
+As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering the distance between them may be up to 1 km depending on traffic conditions.
+%Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i et "black box" hukommelseschip.
 Et lignende modul skal også være til stede i følgebilen, så følgebilen omvendt kan reagere på denne data, enten automatisk eller ved menneskelig ageren, og sende kommandoer tilbage til solbilens CAN bus.
@@
 Dette skyldes at det hardware design vi kommer frem til som produkt kan redigeres mere fleksibelt undervejs på et Teensy og FPGA board, fremfor et PCB board som er mere special designet til formålet.\\
+\\
 Det skal dog siges at vi forventer at et PCB board design ville være mindre strømkrævende end de løsninger vi ser på, men vi har i dette projekt valgt at prioritere et godt design over strømbesparelse, siden der formentlig er tale om forskel i størrelsesorden af nogen tiende af mA.
-\section{Målgruppe}
-Målgruppen for vores projekt er DTU ROAST holdet, som skal integrere telemetrimodulet i deres solbil. Efter planen skal bilen så deltage ved Bridgestone World Solar Challenge i 2023.
+\section{Target audience}
+The target audience for our project is the DTU ROAST team, who will integrate the telemetry module in the solar car and support vehicle.
+Subsequently, the car is set to enter the 2023 Bridgestone World Solar Challenge.
 
@@
     \item 2x nRF24L01 + PA/LNA Wireless Transceiver with antenna
-    \item 1x SD memory card
-    \item 2x MCP2551 CAN Module (Transceiver & Controller) m. breakout board
+    \item 1x Micro SD memory card
+    \item 1x MCP2551 CAN Transceiver w/ breakout board
 \end{itemize}
@@
 
-\section{Referenceliste}
-\begin{thebibliography}{9}
+\section{References}
+\begingroup
+\renewcommand{\section}[2]{}%
+%\renewcommand{\chapter}[2]{}% for other classes
+\begin{thebibliography}{}
 \bibitem{wsc}
@@
 \end{thebibliography}
+\endgroup
 \section{Grøn dyst?}

### 2021-03-01 documentation/projektplan/problemformulering.tex
Updates from Overleaf

@@
 \usepackage{graphicx}
-%\usepackage[intoc]{nomencl} % Nomenclature package
 \usepackage[usenames]{xcolor}
@@
 \usepackage{lscape}
-%\usepackage[showboxes]{textpos}
 \usepackage{indentfirst}
@@
 \usepackage{float}
-%\usepackage{mathtools}
 \usepackage{pict2e}
-%\usepackage{steinmetz}
 \usepackage[export]{adjustbox}
@@
 \textwidth=168 mm
-\textheight=230 mm
+\textheight=210 mm
 \parskip 0.27 em
@@
 \title{Projektplan}
+
 \begin{document}
@@
 \begin{center}
-  {\color{dtured}\large31015 Fagprojekt}\\
+  {\color{dtured}\large31015 Introductory project}\\
   \normalsize
@@
      \hline
-          DTU Elektro&Forår 2021 & Gruppe: 7, ID: MEK2 \\\hline
-          Kursus 31015 & Titel & Gruppemedlemmer \\\hline
-          Fagprojekt - Elektroteknologi & ROAST Telemetri & \begin{tabular}{l} s186083 Tjark Petersen\\s194006 Steffan Martin Kunoy\\s194027 Victor Alexander Hansen \end{tabular}\\\hline
-          Dokument:& Projektformulering & 1 side\\\hline 
-          Version/Status: & 1. udgave &\today\\\hline
+          DTU Electro&Spring 2021 & Group: 7, ID: MEK2 \\\hline
+          Course 31015 & Title & Group members \\\hline
+          Introductory project - Electrotechnology & ROAST Telemetry & \begin{tabular}{l} s186083 Tjark Petersen\\s194006 Steffan Martin Kunoy\\s194027 Victor Alexander Hansen \end{tabular}\\\hline
+          Document:& Project plan & 4 pages\\\hline 
+          Version/Status: & 1. edition &\today\\\hline
      \end{tabularx}
@@
 \end{itemize}
+\newpage
+
 \begin{landscape}
-\section{Aktivitetsplan}
-\hspace{1.5cm}\includegraphics[width=1.45\textwidth]{documentation/images/projectPlan.pdf}
+\section{Activity Plan}
+\hspace{1.5cm}\includegraphics[width=1.35\textwidth]{documentation/images/projectPlan.pdf}
 \end{landscape}
+
+\newpage
 Meetings:\\
@@
 \endgroup
-\section{Grøn dyst?}
+\section{Green challenge?}
+
 \end{document}

### 2021-03-01 documentation/projektplan/problemformulering.tex
Updates from Overleaf

@@
   \noindent
-  \includegraphics[width=4.5cm,valign=t]{documentation/resources/tex_dtu_elektro_a.pdf}
-  \hspace*{10.6cm}
+  \includegraphics[width=6.5cm,valign=t]{documentation/resources/Stor-mobil-top_DTU_Elektro_UK.png}
+  \hspace*{8.6cm}
   \includegraphics[width=1.5cm,valign=t]{documentation/resources/tex_dtu_logo.pdf}
@@
 
-This project will concern a telemetry module for use by the DTU Roadrunners Solar Team (ROAST).
-In 2023 ROAST is set to participate in the Bridgestone World Solar Challenge, a race spanning 3000 km across Australia \cite{wsc}.
-The solar car must be able to drive 3 stretches of 1200 km without recharging.
-Thus, solar power will be the main energy source during the race.
-\\
+This project concerns a telemetry module for use by the DTU Roadrunners Solar Team (ROAST).
+In 2023, ROAST is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in Northern Territory to Adelaide in South Australia.
+The solar car have limited opportunities for recharging during the race, so the main source of energy during the race, is from solar energy.
+To this end, solar panels will cover the car, however, the solar car is only allowed to have a maximum of 4 square meters of solar panels.
+This emphasizes the need to create an energy-efficient solar car to win the race \cite{wsc}.\\
 %Dette projekt vil omhandle et telemetri modul som skal indgå i DTU Roadrunners Solar Team (ROAST).
 I 2023 deltager ROAST i Bridgestone World Solar Challenge, som er et løb på over 3000 km tværs gennem Australien \cite{wsc}.
-Solbilen skal køre en distance på 3 x 1200km, hvor bilen ikke har mulighed for at oplade.
 Den primære energikilde for bilen under løbet er derfor solenergi.\\
 \\
-Throughout the race the solar car will be monitored by a support car, which can monitor the car's conditions and issue commands.
+Throughout the race the solar car will be monitored by a support car, which can monitor the car's condition and issue commands.
 This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
@@
 \\
-The main function of the telemetry module is to read data from the solar car's onboard CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support car, while simultaneosly logging data locally in a "black box" memory chip.
+The main function of the telemetry module is to read data from the solar car's on-board CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support car, while simultaneously logging data locally in a "black box" memory chip.
 The support car is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus.
-As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering the distance between them may be up to 1 km depending on traffic conditions.
+As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering the distance between them may be up to 1 km depending on traffic conditions.\\
 %Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i et "black box" hukommelseschip.
@@
 Telemetri modulet spiller derfor en afgørende rolle, da man også skal tage højde for at afstanden mellem solbil og følgebil kan være op til 1 km alt afhængig af trafikforholdene.
+\\
+Therefore, the telemetry module must both serve the purpose of fulfilling the necessary specifications for communication, while also being as energy-efficient as possible.
+The project will mainly focus on the former that is, ensuring that the module meets the specifications with a solid solution.
 
@@
 
-\section{Afgrænsning}
-Vi har i vores gruppe diskuteret flere forskellige umiddelbare løsningsformer for projektets produkt, heriblandt at lave en løsning på et PCB board, eller en løsning på et Teensy board/FPGA board.
+\section{Problem scope}
+
+There are multiple solutions to the problem at hand that were discussed, among them implementing a whole solution on a PCB or more modular solutions involving prebuild microcontrollers like the Teensy\cite{teensy}.
+We chose to consider prebuild microcontroller based solutions and to not focus on implementing the final product on a PCB.
+The decision was based on us wanting to design a product which is modular and flexible and thus can be altered and improved during its life time.\\
+\\
+A PCB implementation is expected to achieve a higher efficiency due to reduced current draw compared to the implementations we are considering.
+We prioritized a more modular and flexible product over the potential power savings though since the difference in current draw is expected to be of the order of some tenths of milli-Ampère.
+
+%Vi har i vores gruppe diskuteret flere forskellige umiddelbare løsningsformer for projektets produkt, heriblandt at lave en løsning på et PCB board, eller en løsning på et Teensy board/FPGA board.
 Vi har valgt at afgrænse os til at se på Teensy og FPGA løsninger, og ikke PCB boards.
 Dette skyldes at det hardware design vi kommer frem til som produkt kan redigeres mere fleksibelt undervejs på et Teensy og FPGA board, fremfor et PCB board som er mere special designet til formålet.\\
-\\
-Det skal dog siges at vi forventer at et PCB board design ville være mindre strømkrævende end de løsninger vi ser på, men vi har i dette projekt valgt at prioritere et godt design over strømbesparelse, siden der formentlig er tale om forskel i størrelsesorden af nogen tiende af mA.
+%\\
+%Det skal dog siges at vi forventer at et PCB board design ville være mindre strømkrævende end de løsninger vi ser på, men vi har i dette projekt valgt at prioritere et godt design over strømbesparelse, siden der formentlig er tale om forskel i størrelsesorden af nogen tiende af mA.
 \section{Target audience}
@@
     \item 2x nRF24L01 + PA/LNA Wireless Transceiver with antenna
-    \item 1x Micro SD memory card
-    \item 1x MCP2551 CAN Transceiver w/ breakout board
+    \item 1x Micro SD memory card (32GB)
+    \item 2x MCP2551 CAN Transceiver w/ breakout board
 \end{itemize}
+At this stage, our telemetry modules, one for both the solar and support car, will consist of the components above.
+The size of the memory SD card is uncertain at this point, but we will use 32GB for testing purposes.\\
+\\
+In addition to the hardware ressources needed for this project, the DTU ROAST team provides a great source of insight for the project.
+Reports from different teams and last years telemetry project lies available, which helps laying the groundwork for our project \cite{ROAST}.
+
+\section{Activity Plan}
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth]{documentation/images/projectPlan.pdf}
+    \caption{The project activity plan}
+    \label{fig:my_label}
+\end{figure}
+\subsection*{Meetings}
+We will have a weekly meeting with both the project coordinator Christian Kampp Kruuse and our supervisor Martin Schoeberl on Fridays.
+In addition, meetings will be arranged with our second supervisor Vitaliy Zhurbenko if necessary.
+\subsection*{Report writing}
+We will begin drafting a project report once we have achieved an MVP (Minimum Viable Product). The report will include full documentation of all design choices in hardware and software. 
+\subsection*{Design}
+The design phase of the project will include designing prototypes to be implemented once the hardware components become available.
+Furthermore, basic software programs may be written and debugged during the design phase.
+We intend for all design, prototyping and testing of the Solar Car Unit (SC) and Support Vehicle Unit (SV) to run in parallel.
+\subsection*{Prototyping}
+During prototyping we want to test the basic functionality of our product.
+The setup should be configurable, which is why we plan to use breadboards and jumper wires.
+Basic tests will be written in software to make sure the product meets the basic requirements.
+\subsection*{Requirements fulfilled}
+Once the basic requirements have been fulfilled by the prototype we will begin rigorously testing to find ways to further optimise our design.
+This is to ensure that the product will not have any fundamental flaws.
+\subsection*{Product Development}
+This stage will concern the design of an MVP as well as later design iterations. 
+\subsection*{GUI}
+As a final touch we intend to create a basic GUI allowing PC users to interact with the telemetry module, as well as sending and receiving data on the CAN bus. 
+\subsection*{Field Testing and Validation}
+Finally, our product will be integrated with the ROAST solar car and support vehicle, allowing us to test it in operation.
+We will be able to gain feedback from the DTU ROAST team and evaluate the project.
 \newpage
+\section{System Specification}
 
-\begin{landscape}
-\section{Activity Plan}
-\hspace{1.5cm}\includegraphics[width=1.35\textwidth]{documentation/images/projectPlan.pdf}
-\end{landscape}
+In the following three subsections feature lists for each of the three systems are presented.
 
-\newpage
-Meetings:\\
-We will have a weekly meeting with both the project coordinator Christian Kampp Kruuse and our advisor Martin Schoeberl at Friday.
-Report writing:\\
+\subsection{Solar Car Unit (SC)}
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|l|l|}
+    \hline
+        \textbf{Feature ID} & \textbf{Description} \\ \hline
+        \textbf{SC.SD} & The system uses a SD card as a permanent storage of log files \\ \hline
+        SC.SD.write & Write a data structure to a file on the SD card \\ \hline
+        SC.SD.read & Read a file into a data structure \\ \hline
+         &  \\ \hline
+        \textbf{SC.RF} & The system needs to communicate via a RF module with the support vehicle \\ \hline
+        SC.RF.rxCmd & Receive a system configuration command from the support vehicle \\ \hline
+        SC.RF.rxCanCmd & Receive a message to be broadcasted via the solar car CAN bus \\ \hline
+        SC.RF.txResp & Transmit response to a received command \\ \hline
+        SC.RF.txLogUnit & Transmit a log unit to the support vehicle \\ \hline
+        SC.RF.txLogFile & Transmit a log file to the support vehicle \\ \hline
+         &  \\ \hline
+        \textbf{SC.CAN} & The system needs to communicate via the solar car CAN bus \\ \hline
+        SC.CAN.rx & Receive all messages broadcasted on the CAN bus \\ \hline
+        SC.CAN.tx & Transmit a message on the CAN bus \\ \hline
+         &  \\ \hline
+        \textbf{SC.BUF} & The system needs to buffer logged data in local flash memory until enough data to fill a file is collected \\ \hline
+        SC.BUF.add & Add a log unit to the buffer \\ \hline
+        SC.BUF.rd & Read a logging unit from the buffer into RAM \\ \hline
+        SC.BUF.rdToFile & Read all logged data from the buffer into a file \\ \hline
+        SC.BUF.clr & Empty the buffer \\ \hline
+    \end{tabular}
+\end{table}
+
+\subsection{Support Vehicle Unit (SV)}
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|l|l|}
+    \hline
+        \textbf{Feature ID} & \textbf{Description} \\ \hline
+        \textbf{SV.RF} & The system needs to communicate via RF with the solar car \\ \hline
+        SV.RF.txCmd & Transmit a system configuration command to the solar car \\ \hline
+        SV.RF.txCanCmd & Transmit a CAN message to the solar car \\ \hline
+        SV.RF.rxResp & Receive a response from the solar car unit \\ \hline
+        SV.RF.rxLogUnit & Receive a log unit from the solar car unit \\ \hline
+        SV.RF.rxLogFile & Receive a log file from the solar car unit \\ \hline
+         &  \\ \hline
+        \textbf{SV.UART} & The system needs to communicate with a computer via UART \\ \hline
+        SV.UART.rxCmd & Receive a command via UART \\ \hline
+        SV.UART.rxCanCmd & Receive a CAN bus message via UART \\ \hline
+        SV.UART.txLogUnit & Transmit a log unit to the computer \\ \hline
+        SV.UART.txLogFile & Transmit a log file to the computer \\ \hline
+    \end{tabular}
+\end{table}
+
+\subsection{Support Vehicle Computer Program (PR)}
+\begin{table}[H]
+    \centering
+    \begin{tabular}{|l|l|}
+    \hline
+        \textbf{Feature ID} & \textbf{Description} \\ \hline
+        \textbf{PR.UART} & The program needs to communicate with the support vehicle unit via UART \\ \hline
+        PR.UART.txCmd & Transmit a system configuration command via UART \\ \hline
+        PR.UART.txCanCmd & Transmit a CAN message via UART \\ \hline
+        PR.UART.rxLogUnit & Receive a log unit via UART \\ \hline
+        PR.UART.rxLogFile & Receive a log file via UART \\ \hline
+         &  \\ \hline
+        \textbf{PR.FILE} & The program needs to output received log files \\ \hline
+        PR.FILE.write & Write a file containing the last received log file \\ \hline
+         &  \\ \hline
+        \textbf{PR.GUI} & The program is interface via a GUI \\ \hline
+        PR.GUI.getConf & Get configuration parameters from GUI \\ \hline
+        PR.GUI.getCanCmd & Get a CAN message from the GUI \\ \hline
+        PR.GUI.monitor & Show incoming log data stream \\ \hline
+        PR.GUI.saveLog & Pass log file to file writer \\ \hline
+    \end{tabular}
+\end{table}
 
@@
 URL: https://worldsolarchallenge.org/
+\bibitem{ROAST}
+Reports and source files provided by DTU ROAST
+\bibitem{teensy}
+Teensy microcontroller,
+URL: https://www.pjrc.com/teensy/
 \end{thebibliography}
 \endgroup
-\section{Green challenge?}
 

### 2021-03-12 documentation/dear_diary/log_diary.tex

--- 2021-03-01 documentation/projektplan/problemformulering.tex
@@
 \begin{document}
-% formel header
-\thispagestyle{empty}
-\vspace*{-1.9cm}
-\begin{textblock*}{18cm}[0,0](0cm,-2cm) %
-  \noindent
-  \includegraphics[width=4.5cm,valign=t]{documentation/resources/tex_dtu_elektro_a.pdf}
-  \hspace*{10.6cm}
-  \includegraphics[width=1.5cm,valign=t]{documentation/resources/tex_dtu_logo.pdf}
-\end{textblock*}
-\begin{textblock*}{19.0cm}(0cm,-2.5cm) %
-\begin{center}
-  {\color{dtured}\large31015 Introductory project}\\
-  \normalsize
-  s186083 Tjark Petersen\\
-  s194006 Steffan Martin Kunoy\\
-  s194027 Victor Alexander Hansen\\
-\end{center}
-\end{textblock*}
-% rød streg
-\vspace{1mm}
-{\hspace*{-0.1cm}
-\color{dtured}\noindent \rule{16.8cm}{5pt}}
+This log shall be used as a technical diary we write throughout the course. It will be used to write the final report more easily.
 
-\begin{table}[H]
-     \centering
-     \begin{tabularx}{\textwidth}{|X|X|X|}
-     \hline
-          DTU Electro&Spring 2021 & Group: 7, ID: MEK2 \\\hline
-          Course 31015 & Title & Group members \\\hline
-          Introductory project - Electrotechnology & ROAST Telemetry & \begin{tabular}{l} s186083 Tjark Petersen\\s194006 Steffan Martin Kunoy\\s194027 Victor Alexander Hansen \end{tabular}\\\hline
-          Document:& Project plan & 4 pages\\\hline 
-          Version/Status: & 1. edition &\today\\\hline
-     \end{tabularx}
- \end{table}
+\section{05-03}
+We've set up PlatformIO in Visual Studio Code as the environment in which we write our code.
 
-\section{Problem introduction}
-
-This project will concern a telemetry module for use by the DTU Roadrunners Solar Team (ROAST).
-In 2023 ROAST is set to participate in the Bridgestone World Solar Challenge, a race spanning 3000 km across Australia \cite{wsc}.
-The solar car must be able to drive 3 stretches of 1200 km without recharging.
-Thus, solar power will be the main energy source during the race.
-\\
-%Dette projekt vil omhandle et telemetri modul som skal indgå i DTU Roadrunners Solar Team (ROAST).
-I 2023 deltager ROAST i Bridgestone World Solar Challenge, som er et løb på over 3000 km tværs gennem Australien \cite{wsc}.
-Solbilen skal køre en distance på 3 x 1200km, hvor bilen ikke har mulighed for at oplade.
-Den primære energikilde for bilen under løbet er derfor solenergi.\\
-\\
-Throughout the race the solar car will be monitored by a support car, which can monitor the car's conditions and issue commands.
-This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
-The support car will therefore be able to analyze and react to the data stream being transmitted from the solar car, even when the driver is preoccupied by driving the car and navigating traffic.
-\\
-%I løbet vil solbilen være ledsaget af en følgebil, som skal overvåge solbilens driftstatus og sende kommandoer til solbilen.
-Dette er en yderst vigtig opgave, da bilerne skal køre tværs gennem Australien i offentlig trafik, hvor alt fra varme til dæktryk udfordrer solbilen.
-Følgebilen vil derfor have mulighed for at analysere og reagere på data fra solbilen, hvis køreren i solbilen er optaget af trafik og betjening af solbilen.
-\\
-The main function of the telemetry module is to read data from the solar car's onboard CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support car, while simultaneosly logging data locally in a "black box" memory chip.
-The support car is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus.
-As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering the distance between them may be up to 1 km depending on traffic conditions.
-%Telemetri modulets opgave er at aflæse sensordata fra solbilens CAN (Controller Area Network) bus og sende dataen via en RF-transceiver til følgebilen såvel som at logge dataen lokalt i et "black box" hukommelseschip.
-Et lignende modul skal også være til stede i følgebilen, så følgebilen omvendt kan reagere på denne data, enten automatisk eller ved menneskelig ageren, og sende kommandoer tilbage til solbilens CAN bus.
-Telemetri modulet spiller derfor en afgørende rolle, da man også skal tage højde for at afstanden mellem solbil og følgebil kan være op til 1 km alt afhængig af trafikforholdene.
-
-\section{Problem description}
-Remote sensing of data from the solar car will give the ROAST team a competitive advantage as the support vehicle will take on a more active role in controlling and optimising the car's performance.
-This will ease the burden on the driver and engage a large part of the supporting crew.
-For these reasons we have chosen to focus our project on the following points:
-%Fjernmåling af data fra solbilen vil give ROAST en fordel i og med at support bilen kan indtage en mere aktiv rolle i styring og optimering af bilen.
-Det letter derfor på kørerens arbejdsbyrde, og fordeler ansvaret for hele ROAST-support holdet.
-Af denne grund vil vi afgrænse projektet på følgende punkter:
-\begin{itemize}
-    \item How can we design a module that can read and write sensor data and commands from a CAN bus network? 
-    %\item Hvordan kan vi lave et modul der kan læse, sende og modtage sensor data og kommandoer i CAN bus netværket?
-    \item How can CAN data be transmitted over a distance of up to one kilometer? 
-    %\item Hvordan kan CAN data overføres trådløst i en afstand på en kilometer?
-    \item How does the support-car module process data upon receiving it from the solar car?
-    %\item Hvordan skal data behandles ved modtagelse i følgebilen?
-    \item What is the best method for storing data locally (black box) while being accesible at a later time? 
-    %\item Hvordan kan data bedst gemmes lokalt (black box) og udlæses på et senere tidspunkt?
-\end{itemize}
-
-\section{Afgrænsning}
-Vi har i vores gruppe diskuteret flere forskellige umiddelbare løsningsformer for projektets produkt, heriblandt at lave en løsning på et PCB board, eller en løsning på et Teensy board/FPGA board.
-Vi har valgt at afgrænse os til at se på Teensy og FPGA løsninger, og ikke PCB boards.
-Dette skyldes at det hardware design vi kommer frem til som produkt kan redigeres mere fleksibelt undervejs på et Teensy og FPGA board, fremfor et PCB board som er mere special designet til formålet.\\
-\\
-Det skal dog siges at vi forventer at et PCB board design ville være mindre strømkrævende end de løsninger vi ser på, men vi har i dette projekt valgt at prioritere et godt design over strømbesparelse, siden der formentlig er tale om forskel i størrelsesorden af nogen tiende af mA.
-\section{Target audience}
-The target audience for our project is the DTU ROAST team, who will integrate the telemetry module in the solar car and support vehicle.
-Subsequently, the car is set to enter the 2023 Bridgestone World Solar Challenge.
-
-\section{Problem solution}
-\begin{enumerate}
-    \item Preliminary design of telemetry module in block diagram form. Establish specifications and requirements for the final product. 
-    %\item Foreløbigt design af telemetrimodulet som blokdiagram. Fastslå specifikationer og krav til løsningen. 
-    \item Select components for preliminary design. Ordering components not available from DTU. 
-    %\item Valg af komponenter til det indledende design. Bestilling af komponenter som ikke kan fremskaffes på DTU. 
-    \item Prototyping telemetry module on two breadboards. Write code to test limited functionality. 
-    %\item Første prototype af telemetrimodul på breadboard. Skrive kode til at teste begrænset funktionalitet. 
-    \item Improve on prototype. Order new components if needed. Optimize code and write more extensive tests. 
-    %\item Foreslå forbedringer til første prototype. evt. bestille nye komponenter. Optimer kode og design mere omfattende tests.
-    \item Improved prototype of telemetry module on breadboard. Repeat Steps 3. and 4. until fundamental requirements in Step 1. are satisfied. 
-    %\item Anden prototype af telemetrimodul på breadboard. Gentag 3. og 4. indtil grundlæggende krav i 1. er opfyldt. 
-    \item Develop more permanent solution, i.e. compact, soldered and configurable. Complete tests in Step 5.
-    %\item Udvikle mere færdig løsning, dvs. kompakt, loddet og programmerbar. Gennemfør tests som i 5.
-    \item Design GUI for interaction with PC. Test with USB connection to support-car module. Remote control via PC. 
-    %\item Design af GUI til interaktion med PC. Test med USB forbindelse til support-bil modul. Styring via PC.  
-    \item Present solution to DTU ROAST team. Discuss improvements and additional features. 
-    %\item Præsenter løsning for DTU ROAST holdet. Diskuter forbedringer og features. 
-    \item Integrate telemetry module with solar car. Ongoing testing of solution. Feedback and evaluate project. 
-    %\item Integration med solbil, løbende tests og evaluering af projekt. 
-\end{enumerate}
-\section{Resources}
-\begin{itemize}
-    \item 1x Teensy 4.0 development board
-    \item 1x Teensy 4.1 development board
-    \item 2x nRF24L01 + PA/LNA Wireless Transceiver with antenna
-    \item 1x Micro SD memory card
-    \item 1x MCP2551 CAN Transceiver w/ breakout board
-\end{itemize}
-\newpage
-
-\begin{landscape}
-\section{Activity Plan}
-\hspace{1.5cm}\includegraphics[width=1.35\textwidth]{documentation/images/projectPlan.pdf}
-\end{landscape}
-
-\newpage
-Meetings:\\
-We will have a weekly meeting with both the project coordinator Christian Kampp Kruuse and our advisor Martin Schoeberl at Friday.
-Report writing:\\
-
-\section{References}
-\begingroup
-\renewcommand{\section}[2]{}%
-%\renewcommand{\chapter}[2]{}% for other classes
-\begin{thebibliography}{}
-\bibitem{wsc}
-Bridgestone World Solar Challenge, 
-URL: https://worldsolarchallenge.org/
-\end{thebibliography}
-\endgroup
-\section{Green challenge?}
+\section{10-03}
+We've made an adjustment in the ressources and will be using a teensyboard 3.6 instead of a 4.1, since the 3.6 board is already in stock at DTU.
+It has a slower clock (180Mhz, compared to 4.1's 600Mhz), but we judge that this won't matter to the project solution.
 

### 2021-03-12 documentation/projektplan/problemformulering.tex

@@
 \section{Problem scope}
-
+% 11-03 tilføjet {
+As mentioned in the introduction, the car will be using the CAN standard for distributing data, which scopes how we gather data and interact with the car significantly.
+A solution therefore requires an interface between the CAN bus and the microcontroller.\\
+% }
+\\
 There are multiple solutions to the problem at hand that were discussed, among them implementing a whole solution on a PCB or more modular solutions involving prebuild microcontrollers like the Teensy\cite{teensy}.
@@
     \item 1x Teensy 4.0 development board
-    \item 1x Teensy 4.1 development board
+    \item 1x Teensy 3.6 development board
     \item 2x nRF24L01 + PA/LNA Wireless Transceiver with antenna

### 2021-03-12 documentation/projektplan/problemformulering.tex
Updates from Overleaf

@@
 In addition, meetings will be arranged with our second supervisor Vitaliy Zhurbenko if necessary.
-\subsection*{Report writing}
-We will begin drafting a project report once we have achieved an MVP (Minimum Viable Product). The report will include full documentation of all design choices in hardware and software. 
+\subsection*{Paper writing}
+We will begin drafting a project paper once we have achieved an MVP (Minimum Viable Product). A Github repository will be updated continuously as a full documentation of the project.
 \subsection*{Design}
@@
 URL: https://www.pjrc.com/teensy/
+\bibitem{C++}
+Website for learning C++
+https://www.learn-cpp.org/
 \end{thebibliography}

### 2021-03-25 code/Teensy/src/can.cpp
Very simple transmit and receive of CAN frame

@@ -1,0 +1,60 @@
+/* CAN bus C++ transmission receiving file
+ *
+ * @author Victor Alexander Hansen
+ * @author Steffan Martin Kunoy
+ * @author Tjark Petersen
+ * 
+ * For CAN instructions for Teensy board, see https://github.com/pierremolinaro/acan
+ * For DTU Roadrunners previous use of Teensy, see https://bitbucket.org/dtucar/autonomous-ecocar/src/master/Source/
+ * For more specific examples of how the CAN was implemented in the Ecocar, see
+ * https://bitbucket.org/dtucar/ecocar/src/master/Code/CanLib/canbus.cpp
+ * https://bitbucket.org/dtucar/ecocar/src/master/Code/CanLib/canbus.h
+ * https://bitbucket.org/dtucar/ecocar/src/master/Code/CanLib/canbusKeywords.h
+*/
+
+#include <Arduino.h>
+#include <ACAN.h>
+
+
+CANMessage frame;
+
+// Function for transmitting message through CAN0
+// Pre condition: CAN identifier must be specified by user
+// Post condition: transmits message saying "HELLO, x"
+void CANtransmit (uint8_t CanId){
+    frame.id = CanId; // Identifier, lower number = higher priority
+    frame.rtr = 0; // Remote Transmit Request, 0 = data frame, 1 = request frame
+    frame.len = 8; // Data Lenght Code, sets the length of Data field
+
+    // Actual message
+    frame.data[0] = 'H';
+    frame.data[1] = 'E';
+    frame.data[2] = 'L';
+    frame.data[3] = 'L';
+    frame.data[4] = 'O';
+    frame.data[5] = ',';
+    frame.data[6] = ' ';
+    frame.data[7] = CanId;
+
+    ACAN::can0.tryToSend(frame);
+}
+
+// Function for receiving message through CAN1
+// Pre condition: Listens for a data frame to receive
+// Post condition: Prints received data fram to serial monitor
+void CANreceive(){
+    ACAN::can1.receive(frame);
+    for (int i = 0; i < 8; i++){
+        Serial.print(frame.data[i]);
+    }
+}
+
+void setup(){
+    Serial.begin(9600); // Baud rate
+    ACANSettings settings (125 * 1000); // Sets wished bitrate
+}
+
+void loop(){
+    CANtransmit(0);
+    CANreceive();
+}

### 2021-03-25 code/Teensy/src/can.cpp
Clearer succes sign

@@ -42,11 +42,13 @@ void CANtransmit (uint8_t CanId){
 // Function for receiving message through CAN1
 // Pre condition: Listens for a data frame to receive
 // Post condition: Prints received data fram to serial monitor
-void CANreceive(){
-    ACAN::can1.receive(frame);
+bool CANreceive(){
+    const bool ok = ACAN::can1.receive(frame);
     for (int i = 0; i < 8; i++){
         Serial.print(frame.data[i]);
+        Serial.print(" ");
     }
+    return ok;
 }
 
 void setup(){
@@ -56,5 +58,7 @@ void setup(){
 
 void loop(){
     CANtransmit(0);
-    CANreceive();
+    if (CANreceive()){
+        Serial.print("Succes");
+    }
 }

### 2021-03-26 code/Teensy/lib/BlackBox/BlackBox.h
Blackbox v1

@@ -1,0 +1,251 @@
+#include <ChRt.h>
+#include <SdFat.h>
+#include <RingBuf.h>
+#include <Arduino.h>
+#include <TimeLib.h>
+#include "util.h"
+#include "logType.h"
+
+
+#define SD_CONFIG  SdioConfig(FIFO_SDIO)
+
+#define NUM_BUFFERS 4
+#define BUFFERS_SIZE 4
+
+
+template<class DT>
+class BlackBox: public SdFs
+{
+    private:
+
+        char fileName[64]; // the current file name
+        FsFile file; // the current file
+
+        RingBuf<FsFile, 512> sdBuffer; // a ring buffer to buffer writes to the SD card
+
+
+
+        DT logBufs[NUM_BUFFERS][BUFFERS_SIZE]; // for buffering new log data
+
+        msg_t freeBuffersQueue[NUM_BUFFERS]; // queue holding free buffers
+        MAILBOX_DECL(freeBuffers, &freeBuffersQueue, NUM_BUFFERS); // mailbox used to interact with free buffers
+
+        msg_t filledBuffersQueue[NUM_BUFFERS]; // queue holding filled buffers ready to be processed
+        MAILBOX_DECL(filledBuffers, &filledBuffersQueue, NUM_BUFFERS); // mailbox used to interact with filled buffers
+
+        DT *writeBuf; // pointer to the head of the buffer currently in use to record logs
+        uint32_t writeHead = 0; // index of the next used slot in the current buffer
+
+
+    //----------------------------------------------------------------------------------------------------------------------
+    /* Setup functions */
+    
+    /**
+     * @brief   For now the constructor does nothing
+     */
+    public: BlackBox(){}
+
+    /**
+     * @brief   Synchronizes remaining logs in buffer to SD and uninitializes it
+     */
+    public: ~BlackBox(){
+        endLogFile();
+        this->end();
+    }
+
+    /**
+     * @brief   Initializes the black box. The Teensy 3.6 real time clock is setup up and the buffer system is initialized.
+     * 
+     * @return              returns false if the initialization failed
+     */
+    public: bool init(){
+        // setup RTC
+        setSyncProvider(getTeensy3Time);
+        delay(100);
+        if (timeStatus()!= timeSet) {
+            return false;
+        }
+
+        // setup write buffer
+        writeBuf = logBufs[0];
+        writeHead = 0;
[…]

### 2021-03-26 code/Teensy/lib/BlackBox/example.cpp

@@ -1,0 +1,70 @@
+
+#include <ChRt.h>
+#include <Arduino.h>
+#include "BlackBox.h"
+#include "ErrorHandler.h"
+
+#define LOG_TYPE CanMessage
+
+BlackBox<LOG_TYPE> bb;
+
+THD_WORKING_AREA(waSdWorker,256);
+THD_FUNCTION(sdWorkerFunc, arg){
+  ((BlackBox<LOG_TYPE>*) arg)->sdWorker();
+}
+
+void chSetup(){
+  chSysInit();
+  chThdCreateStatic(waSdWorker, sizeof(waSdWorker), NORMALPRIO-1, sdWorkerFunc, &bb);
+}
+
+void setup() {
+
+  pinMode(LED_BUILTIN,OUTPUT);
+
+  Serial.begin(9600);
+
+  while(!Serial){}
+
+
+  if(!bb.init()){
+    Serial.println("Failed to initialize black box");
+  }
+
+  bb.startNewLogFile();
+
+
+  chBegin(chSetup);
+  
+}
+
+void loop() {
+  uint32_t counter = 0;
+
+  while(true){
+    while(!Serial.available()){
+      chThdSleepMilliseconds(100);
+    }
+    char rd = Serial.read();
+    if(rd == 13){
+      break;
+    }
+    CanMessage msg;
+    msg.id = counter;
+    msg.ext = true;
+    msg.rtr = true;
+    msg.len = 4;
+    msg.data = 1LL*counter*10;
+    bb.addNewLog(&msg);
+    bb.printCurrentBuffer();
+    chThdSleepMilliseconds(100);
+    Serial.flush();
+    counter++;
+  }
+
+  Serial.println("Exiting Program!");
+
+  bb.~BlackBox();
+
+  while(true){}
+}

### 2021-03-26 code/Teensy/lib/BlackBox/logType.h

@@ -1,0 +1,36 @@
+#include <Arduino.h>
+
+
+class logType {
+    public:
+        uint8_t secs;
+        uint8_t mins;
+        uint8_t hours;
+        virtual const String toString() = 0;
+        virtual void reset() = 0;
+};
+
+class CanMessage : public logType {
+    public:
+        uint32_t id;
+        bool ext;
+        bool rtr;
+        uint8_t len;
+        uint64_t data;
+        
+        const String toString(){
+            char tmp[200];
+            snprintf(tmp,200,"\"%02d-%02d-%02d\", %8" PRIx32 ",     %1d,     %1d,     %1d,   %19" PRId64 "",hours,mins,secs,id,ext,rtr,len,data);
+            return String(tmp);
+        }
+        static const String getHeader(){
+            return "  \"time\"  ,   \"id\"  , \"ext\", \"rtr\", \"len\",         \"data\"";
+        }
+        void reset(){
+            id = 0;
+            ext = 0;
+            rtr = 0;
+            len = 0;
+            data = 0LL;
+        }
+};

### 2021-03-26 code/Teensy/lib/BlackBox/util.h

@@ -1,0 +1,39 @@
+#include <Arduino.h>
+
+class ExecutionTimeMeasurer{
+    bool killed;
+    const char* name;
+    uint32_t startTime;
+    public:
+        ExecutionTimeMeasurer(const char* name);
+        ~ExecutionTimeMeasurer();
+        void kill();
+        bool isKilled();
+};
+
+#define MEASURE_EXEC_TIME(name) for (ExecutionTimeMeasurer x(name); !x.isKilled(); x.kill())
+
+class ActivityMeasurer{
+    bool killed;
+    uint32_t startTime;
+    public:
+        ActivityMeasurer();
+        ~ActivityMeasurer();
+        void kill();
+        bool isKilled();
+};
+
+#define SHOW_ACTIVITY for (ActivityMeasurer x; !x.isKilled(); x.kill())
+
+class ActivityAndExecutionTimeMeasurer{
+    bool killed;
+    const char* name;
+    uint32_t startTime;
+    public:
+        ActivityAndExecutionTimeMeasurer(const char* name);
+        ~ActivityAndExecutionTimeMeasurer();
+        void kill();
+        bool isKilled();
+};
+
+#define ACT_PLUS_MEAS(name) for (ActivityAndExecutionTimeMeasurer x(name); !x.isKilled(); x.kill())

### 2021-03-26 code/Teensy/src/can_hello.cpp
Update and rename can.cpp to can_hello.cpp
Added can0.begin and can1.begin to setup

@@ -54,6 +54,8 @@ bool CANreceive(){
 void setup(){
     Serial.begin(9600); // Baud rate
     ACANSettings settings (125 * 1000); // Sets wished bitrate
+    ACAN::can1.begin(settings);
+    ACAN::can0.begin(settings);
 }
 
 void loop(){

### 2021-03-26 code/Teensy/src/canRF.cpp
Added canRF.cpp

@@ -1,0 +1,76 @@
+/* CAN bus and RF communication
+ *
+ * @author Victor Alexander Hansen
+ * @author Steffan Martin Kunoy
+ * @author Tjark Petersen
+ * 
+ * For CAN instructions for Teensy board, see https://github.com/pierremolinaro/acan
+ * For DTU Roadrunners previous use of Teensy, see https://bitbucket.org/dtucar/autonomous-ecocar/src/master/Source/
+ * For more specific examples of how the CAN was implemented in the Ecocar, see
+ * https://bitbucket.org/dtucar/ecocar/src/master/Code/CanLib/canbus.cpp
+ * https://bitbucket.org/dtucar/ecocar/src/master/Code/CanLib/canbus.h
+ * https://bitbucket.org/dtucar/ecocar/src/master/Code/CanLib/canbusKeywords.h
+*/
+
+/* TODO:
+ * Kan payload til antenne effektiviseres? E.g. kan vi skrive en funktion som
+ * reducerer størrelsen af payload, men som stadig aflæses korrekt ved ankomst, ek-
+ * sempelvis ved en bit i front som er størrelse på payload?
+ * Add code for RF transmission, possibly new file
+*/
+
+#include <Arduino.h>
+#include <ACAN.h>
+
+
+CANMessage frame;
+
+// Function for transmitting message from RF24 to CAN
+void RFtoCAN (uint32_t payloadR){
+    frame.id = ((0x11 << 16) & payloadR); // Identifier
+    frame.rtr = 0; // Remote Transmit Request, 0 = data frame, 1 = request frame
+    frame.len = 8; // Data Lenght Code, sets the length of Data field
+
+    frame.data[0] = 0x11 & payloadR;
+    frame.data[1] = (0x11 << 2) & payloadR;
+    frame.data[2] = (0x11 << 4) & payloadR;
+    frame.data[3] = (0x11 << 6) & payloadR;
+    frame.data[4] = (0x11 << 8) & payloadR;
+    frame.data[5] = (0x11 << 10) & payloadR;
+    frame.data[6] = (0x11 << 12) & payloadR;
+    frame.data[7] = (0x11 << 14) & payloadR;
+
+    ACAN::can0.tryToSend(frame);
+}
+
+// Transforms a CAN message to antenna payload on the form
+// CanId,canData[7][6][5][4][3][2][1][0]
+uint32_t CANtoRF (){
+    uint32_t payloadT = 0x0;
+    int n = 0;
+
[…]

### 2021-03-26 documentation/dear_diary/log_diary.tex
Updates from Overleaf

@@
 
+\section{12-03}
+We've decided to divide the work among us.
+Tjark will be having the 3.6 Teensy board with the SD module, Steffan will be having the 4.0 Teensy board and Victor will work on the related software.
+As Tjark has a soldering station, he will take care of soldering the components, if this is not done in laboratory.
+
+More specifically, Steffan will look into RF, Tjark into SD storage and Victor at CANbus conncection.
+
 \end{document}

### 2021-06-03 code/SolarVehicle/lib/message/TelemetryMessages.h
Split functionality build flags + usb test message gen

@@ -142,7 +142,7 @@ class CanTelemetryMsg {
         return "\"time\",\"id\",\"rtr\",\"len\",\"data\"";
     }
 
-    #ifdef TEENSY36_BOARD
+    #ifdef CAN_ADD_TIME_STAMPS
 
     public: CanTelemetryMsg(CANMessage *msg){
         id = msg->id;
@@ -165,9 +165,9 @@ class CanTelemetryMsg {
         h = hour();
     }
 
-    #else
+    #endif
 
-
+    #ifdef CAN_SERIALIZE_TO_JSON
     public: const String toJSON(char *buf, uint32_t len){
       return snprintf(buf,len,"{\"cmd\":%d,\"can\":{\"id\":%d,\"rtr\":%s,\"len\":%d,\"data\":%llu,\"stamp\":{\"hour\":%d,\"minute\":%d,\"second\":%d}}}\n",cmd,id,rtr ? "true" : "false",this->len,data64,h,m,s);
     }

### 2021-06-03 code/TelemetryUI/src/main/scala/telemetryui/udp/UdpServerTest.scala
Add udp server to telemetryUI

@@ -1,0 +1,41 @@
+package telemetryui.udp
+
+import telemetryui.types.{CanFrame, TimeStamp}
+
+import java.net.{DatagramPacket, DatagramSocket, InetAddress}
+
+object UdpServerTest extends App {
+  (new UdpServerTest).run()
+}
+
+class UdpServerTest extends Thread {
+
+  val socket = new DatagramSocket(4445)
+  var buf = new Array[Byte](16)
+
+  var running = true
+
+  override def run(): Unit = {
+    while (running) {
+      val rPacket = new DatagramPacket(buf,buf.length)
+      socket.receive(rPacket)
+      val address = rPacket.getAddress
+      val port = rPacket.getPort
+      val received = new String(rPacket.getData,0,rPacket.getLength)
+      if(received.equals("start")){
+        println("Starting...")
+        for(i <- 0 until 10){
+          buf = CanFrame(101,false,4,i,TimeStamp(10,8,i)).toByteArray
+          val packet = new DatagramPacket(buf,buf.length,address,port)
+          println("sent...")
+          println(socket.isClosed)
+          println(socket.isConnected)
+          socket.send(packet)
+          Thread.sleep(1000)
+        }
+      }
+
+    }
+    socket.close()
+  }
+}

### 2021-06-03 code/matlab/udpClient.m
add matlab udp listener test

@@ -1,0 +1,27 @@
+%echoudp("on",3030);
+u = udpport("datagram","OutputDatagramSize",16);
+
+write(u,"start","uint8","127.0.0.1",4445);
+
+
+
+for i = 1:10
+    u.NumDatagramsAvailable
+    while u.NumDatagramsAvailable == 0
+    end
+    data = read(u,u.NumDatagramsAvailable,"uint8");
+    d = data.Data
+    can.id = bitor(bitshift(d(2),8),d(1),"uint16");
+    can.rtr = d(3);
+    can.len = d(4);
+    can.data64 = d(5);
+    for k = 1:7
+       can.data64 = bitor(can.data64,bitshift(d(k+5),8*k));
+    end
+    can.data = d(5:12);
+    can.t.hour = d(13);
+    can.t.min = d(14);
+    can.t.sec = d(15);
+    can
+    can.t
+end

### 2021-06-03 documentation/FinalPaper/paper.tex
add final paper IEEE template

\documentclass[journal]{IEEEtran}

\usepackage{cite}
\usepackage[pdftex]{graphicx}

\usepackage{url}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Bare Demo of IEEEtran.cls\\ for IEEE Journals}

\author{Victor Alexander Hansen, Steffan Martin Kunoy, Tjark Petersen}

\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The abstract goes here.
\end{abstract}

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
IEEE
\end{IEEEkeywords}

\section{Introduction}
% The very first letter is a 2 line initial drop letter followed
% by the rest of the first word in caps.
% 
% form to use if the first word consists of a single letter:
% \IEEEPARstart{A}{demo} file is ....
% 
% form to use if you need the single drop letter followed by
% normal text (unknown if ever used by the IEEE):
% \IEEEPARstart{A}{}demo file is ....
% 
% Some journals put the first two words in caps:
% \IEEEPARstart{T}{his demo} file is ....
% 
% Here we have the typical use of a "T" for an initial drop letter
% and "HIS" in caps to complete the first word.
\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
for IEEE journal papers produced under \LaTeX\ using
IEEEtran.cls version 1.8b and later.
% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
I wish you the best of success.

\hfill mds

\hfill August 26, 2015

\subsection{Subsection Heading Here}
Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

\section{Conclusion}
The conclusion goes here.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

\appendices
\section{Proof of the First Zonklar Equation}
Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
\section{}
Appendix two text goes here.

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,../bib/paper}

\end{document}

### 2021-06-04 code/SolarVehicle/lib/canRF/canRF.h
new pull

@@ -1,61 +1,0 @@
-/*
- @author Victor Alexander Hansen
- @author Steffan Martin Kunoy
- @author Tjark Petersen
-*/
-
-#include <ACAN.h>
-#include <SPI.h>
-
-#ifndef CANRF_H
-#define CANRF_H
-
-CANMessage frame;
-
-void setCan(int bitrate) {
-    ACANSettings settings (bitrate); // Sets wished bitrate
-    ACAN::can1.begin(settings);
-    ACAN::can0.begin(settings);
-}
-
-// Function for transmitting message from RF24 to CAN
-void RFtoCAN (char* payloadR){
-    frame.id = ((0x11 << 16) & atoi(payloadR)); // Identifier
-    frame.rtr = 0; // Remote Transmit Request, 0 = data frame, 1 = request frame
-    frame.len = 8; // Data Lenght Code, sets the length of Data field
-
-    frame.data[0] = 0x11 & atoi(payloadR);
-    frame.data[1] = (0x11 << 2) & atoi(payloadR);
-    frame.data[2] = (0x11 << 4) & atoi(payloadR);
-    frame.data[3] = (0x11 << 6) & atoi(payloadR);
-    frame.data[4] = (0x11 << 8) & atoi(payloadR);
-    frame.data[5] = (0x11 << 10) & atoi(payloadR);
-    frame.data[6] = (0x11 << 12) & atoi(payloadR);
-    frame.data[7] = (0x11 << 14) & atoi(payloadR);
-
-    //ACAN::can0.tryToSend(frame);
-}
-
-// Transforms a CAN message to antenna payload on the form
-// CanId,canData[7][6][5][4][3][2][1][0]
-String CANtoRF () {
-    int payloadT = 0x0;
-    int n = 0;
-
-    if (ACAN::can1.available()) {
-        ACAN::can1.receive(frame);
-    }
-    // Transforms canData to one integer
-    while(n < 8){
-        payloadT += (frame.data[n] << (n*2));
-        n++;
-    }
-    // Adds the canID to the most significant bits of the integer
-    payloadT += (frame.id << n*2);
-    char tmp[200];
-    String t = snprintf(tmp, 200, "%d", payloadT);
-
-    return t;
-}
-
-#endif

### 2021-06-04 code/SolarVehicle/lib/message/TelemetryMessages.h

@@ -1,3 +1,17 @@
+/**
+ * This file contains all the message types used in the telemetry system.
+ * 
+ * All messages have a length of 32 bytes where the first byte is encoding the
+ * type of the message.
+ * 
+ * All messages can be casted to the BaseTelemetryMsg which treats the data
+ * payload as raw bytes.
+ * 
+ * Message types are:
+ *  - CanTelemetryMsg:  Contains a CAN frame and a time stamp
+ *  - FileStreamMsg:    Can be used to stream characters from a file
+ */
+
 #ifndef __TELEMETRYMESSAGES_H__
 #define __TELEMETRYMESSAGES_H__
 
@@ -10,6 +24,9 @@
 
 #endif
 
+/**
+ * This type is used for indentifying message types
+ */
 typedef enum CMD: uint8_t {
   RECEIVED_CAN = 0x00, 
   BROADCAST_CAN = 0x01,
@@ -20,15 +37,9 @@ typedef enum CMD: uint8_t {
   FILE_STREAM_DATA = 0x06
 } cmd_t;
 
-
-
-
-
-
-
-
-
-
+/**
+ * The base type for all messages treating the payload data as raw bytes
+ */
 class BaseTelemetryMsg {
 
   public: cmd_t cmd;
@@ -47,11 +58,9 @@ class BaseTelemetryMsg {
 
 }__attribute__((packed));
 
-
-
-
-
-
+/**
+ * A message type used to stream characters from a file
+ */
 class FileStreamMsg {
   public: cmd_t cmd;
   public: char str[31];
@@ -66,13 +75,12 @@ class FileStreamMsg {
 
 }__attribute__((packed));
 
-
-
-
-
+/**
+ * A message type used to send a CAN frame and a associated time stamp
+ */
 class CanTelemetryMsg {
     public:
-      cmd_t cmd;          // offset = 0x00:      opcode of the message
+      cmd_t cmd;        // offset = 0x00:      type of the message
       // data payload
       uint16_t id;      // offset = 0x01-0x02: identifier accompanying the CAN frame
       uint8_t rtr;      // offset = 0x03:      indicator of whether the frame is requesting data from another device

### 2021-06-04 code/SolarVehicle/lib/util/ChUtil.h

@@ -1,3 +1,5 @@
+#ifndef __CH_UTIL__
+#define __CH_UTIL__
 #include "ChRt.h"
 #include "TelemetryMessages.h"
 
@@ -61,3 +63,5 @@ void fifoBuffer::fifoMoveTail()
 {
     _fifoTail = _fifoTail < (_fifoSize - 1) ? _fifoTail + 1 : 0; 
 }
+
+#endif

### 2021-06-04 code/SolarVehicle/src/solar-car/RfThd.h

@@ -1,0 +1,70 @@
+/**
+ * This file contains the function for the thread responsible for streaming CAN
+ * messages to the support vehicle.
+ */
+
+#ifndef __RF_THD_H__
+#define __RF_THD_H__
+
+#include <ChRt.h>
+
+#include "ThreadState.h"
+#include "BlackBox.h"
+#include "Fifo.h"
+#include "RFfunctions.h"
+
+#include "solar-car/Mutexes.h"
+
+/**
+ * A bundle used for passsing all relevant resources to the radio thread
+ */
+struct rfWorkerBundle{
+    Fifo<CanTelemetryMsg> *fifo;
+    ThreadState *state;
+};
+
+// the working area for the thread is 512 bytes
+THD_WORKING_AREA(waRfWorker,512);
+
+THD_FUNCTION(rfWorker, arg){
+
+  rfWorkerBundle *bundle = (rfWorkerBundle*) arg;
+  ThreadState *state = bundle->state;
+  Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
+
+  CanTelemetryMsg *msg;
+
+  uint32_t fifoTail = 0;
+
+  WITH_MTX(serialMtx){Serial.println("Starting RF Transmitting thread...");}
+
+  while(!state->terminate){
+   
+    fifo->waitForData();
+
+    if(state->pause){
+      state->suspend();
+    }
+
+    msg = fifo->get(fifoTail);
+
+    Serial.println("hello");
+
+    WITH_MTX(serialMtx){
+      Serial.print("rf worker msg nr. ");
+      Serial.println((uint32_t)msg->data64);
+    }
+    
+    WITH_MTX(rfMTX){
+      chSysLock();
+      RFtransmit((BaseTelemetryMsg*)msg,32);
+      chSysUnlock();
+    }
+    
+    fifo->advance(&fifoTail);
+        
+  }
+  
+}
+
+#endif

### 2021-06-04 code/SolarVehicle/src/solar-car/blackBoxThd.h

@@ -1,0 +1,69 @@
+/**
+ * This file contains the function for the thread responsible for saving CAN messages from the system FIFO into the blackbox
+ */
+
+#ifndef __BLACKBOX_THD_H__
+#define __BLACKBOX_THD_H__
+
+#include <ChRt.h>
+
+#include "ThreadState.h"
+#include "BlackBox.h"
+#include "Fifo.h"
+#include "MutexLocker.h"
+
+#include "solar-car/Mutexes.h"
+
+/**
+ * A bundle used for passsing all relevant resources to the radio thread
+ */
+struct BlackboxWorkerBundle{
+    Fifo<CanTelemetryMsg> *fifo;
+    ThreadState *state;
+    BlackBox *bb;
+};
+
+// the working area for the thread is 512 bytes
+THD_WORKING_AREA(blackBoxWorker,512);
+
+THD_FUNCTION(canWorkerFunc, arg){
+
+  BlackboxWorkerBundle *bundle = (BlackboxWorkerBundle*) arg;
+  ThreadState *state = bundle->state;
+  Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
+  BlackBox *bb = bundle->bb;
+
+  CanTelemetryMsg *msg;
+
+  uint32_t fifoTail = 0;
+
+  WITH_MTX(serialMtx){Serial.println("Starting listener...");}
+
+  while(!state->terminate){
+   
+    fifo->waitForData();
+
+    if(state->pause){
+      state->suspend();
+    }
+
+    msg = fifo->get(fifoTail);
+
+
+    if(Serial){
+      WITH_MTX(serialMtx){
+        char str[64];
+        msg->toString(str,64);
+        Serial.println(str);
+      }
+    }
+
+    fifo->signalRead();
+    fifo->advance(&fifoTail);
+        
+  }
+
+  
+}
+
+#endif

### 2021-06-04 code/SolarVehicle/src/solar-car/solarCarMain.cpp

@@ -1,0 +1,110 @@
+/**
+ * This file contains the main function for the solar car unit.
+ * 
+ * It sets up three worker threads:
+ *  - the CAN receiver worker, responsible for putting newly received messages in a FIFO
+ *  - the blackbox worker, responsible for adding messages from the FIFO to the logs
+ *  - the radio worker, responsible for streaming messages from the FIFO to the support vehicle
+ * 
+ * The main thread is responsible for listing for commands from the support vehicle and executing them.
+ * 
+ */
+
+#include <Arduino.h>
+#include <ACAN.h>
+#include <ChRt.h>
+
+#include "BlackBox.h"
+#include "RFfunctions.h"
+#include "TelemetryMessages.h"
+#include "Fifo.h"
+#include "ThreadState.h"
+
+#include "solar-car/blackBoxThd.h"
+#include "solar-car/CanReceiverThd.h"
+#include "solar-car/Mutexes.h"
+#include "solar-car/RfThd.h"
+
+/**
+ * Singleton containing global state variables
+ */
+class GlobalState {
+  public:
+    uint8_t isLogging = true;
+    uint8_t isTransmittingRT = true;
+} globalState;
+
+Fifo<CanTelemetryMsg> canFifo(32);
+BlackBox bb;
+ThreadState blackBoxWorkerState;
+ThreadState canReceiverState;
+ThreadState rfWorkerState;
+
+void chSetup();
+void setup(){
+
+  // initialize serial port
+  Serial.begin(9600);
+  while(!Serial){} //needs to be removed when headless!!!!!!!!!!!!!!!!!!!!
+
+  // setup built in LED
+  pinMode(LED_BUILTIN,OUTPUT);
+
+  Serial.println("Initializing...");
+
+  // setup CAN bus
+  //ACANSettings settings(125 * 1000);
+  //if(ACAN::can0.begin(settings)){Serial.println("CAN setup failed!");}
+
+  // setup radio module
+  RFinit();
+
+  canFifo.clear();
+
+  Serial.println("Starting...");  
+
+  chBegin(chSetup);
+
+}
+
+void chSetup(){
+
+  chSysInit();
+  
+  // create the three worker threads
+  BlackboxWorkerBundle blackBoxWorkerBundle = {.fifo = &canFifo, .state = &blackBoxWorkerState, .bb = &bb};
[…]

### 2021-06-04 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp

@@ -1,6 +1,10 @@
 #include "Fifo.h"
+<<<<<<< HEAD:code/SolarVehicle/src/mainTeensy40.cpp
 #include "util.h"
 #include "RFfunctions.cpp"
+=======
+#include "RFfunctions.h"
+>>>>>>> 375268759b6cf456ffd2f2b1200d62b716760b62:code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp
 
 // Initialize state machine for RXthread and TXthread
 ThreadState radioReceiverThreadState, radioTransmitterThreadState, serialReceiverThreadState, serialTransmitterThreadState; 

### 2021-06-04 code/SolarVehicle/lib/rf/RFfunctions.h
commit

@@ -30,7 +30,7 @@ void RFinit()
     radio.enableDynamicPayloads();
     radio.setAutoAck(true);
     radio.enableAckPayload();
-    //radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
+    radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
     radio.openWritingPipe(address[!radioNumber]);
     radio.openReadingPipe(1, address[radioNumber]);
     radio.startListening();                                     // Starts RX mode
@@ -67,11 +67,15 @@ bool RFreceive(BaseTelemetryMsg *received)
     uint8_t pipe;
     if (radio.available(&pipe)){                                // Check if transmitter is sending message
         radio.read(received, 32);                               // Read message, cannot be larger than 32 bytes (null-terminated)
+        char str[64]; 
+        received->toString(str, sizeof(str));
+        Serial.println(str); 
+        /*
         if (received->cmd == RECEIVED_CAN){
             char tmp[64];
             ((CanTelemetryMsg *)received)->toString(tmp, sizeof(tmp));
             Serial.println(tmp);                                // Print message
-        }
+        }*/
         radio.writeAckPayload(1, received, 32);                 // Send acknowledge payload
         return true;
     }

### 2021-06-04 code/SolarVehicle/lib/util/Fifo.h

@@ -97,6 +97,7 @@ class Fifo {
         return advance(_fifoTail) == _fifoHead;
     }
 
+    // Used to check whether fifo is empty
     public: bool fifoEmpty(){
         return _fifoTail == _fifoHead; 
     }

### 2021-06-04 code/SolarVehicle/src/mainArduino1.cpp

@@ -1,14 +1,0 @@
-#include "RFfunctions.cpp"
-
-void setup() 
-{
-  Serial.begin(9600);
-  RFinit();
-}
-
-void loop() 
-{
-  CanTelemetryMsg msg;
-  RFtransmit(msg.toMessage(), sizeof(msg));
-  delay(1000);
-}

### 2021-06-04 code/SolarVehicle/src/support-vehicle/Mutexes.h

@@ -1,0 +1,12 @@
+/**
+ * This file contains mutex declarations for all the resources shared by multiple threads
+ */
+
+#ifndef __MUTEXES_H__
+#define __MUTEXES_H__
+
+#include <ChRt.h>
+
+MUTEX_DECL(rfMTX);
+
+#endif

### 2021-06-04 code/SolarVehicle/src/support-vehicle/RfThd.h

@@ -7,12 +7,10 @@
 #define __RF_THD_H__
 
 #include <ChRt.h>
-
 #include "ThreadState.h"
 #include "Fifo.h"
 #include "RFfunctions.h"
-
-#include "solar-car/Mutexes.h"
+#include "support-vehicle/Mutexes.h"
 
 /**
  * A bundle used for passsing all relevant resources to the radio thread
@@ -23,19 +21,19 @@ struct rfWorkerBundle{
 };
 
 // the working area for the thread is 512 bytes
-THD_WORKING_AREA(waRfWorker,512);
+THD_WORKING_AREA(waRXthread,512);
 
-THD_FUNCTION(rfWorker, arg){
+THD_FUNCTION(RXthread, arg){
 
   rfWorkerBundle *bundle = (rfWorkerBundle*) arg;
   ThreadState *state = bundle->state;
   Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
 
-  CanTelemetryMsg *msg;
+  //CanTelemetryMsg *msg;
 
   uint32_t fifoTail = 0;
 
-  WITH_MTX(serialMtx){Serial.println("Starting RF Transmitting thread...");}
+  //WITH_MTX(serialMtx){Serial.println("Starting RF Transmitting thread...");}
 
   while(!state->terminate){
    
@@ -45,14 +43,45 @@ THD_FUNCTION(rfWorker, arg){
       state->suspend();
     }
 
-    msg = fifo->get(fifoTail);
+    //msg = fifo->fifoHead();
 
     Serial.println("hello");
-
-    WITH_MTX(serialMtx){
-      Serial.print("rf worker msg nr. ");
-      Serial.println((uint32_t)msg->data64);
+    
+    WITH_MTX(rfMTX){
+      chSysLock();
+      RFreceive(fifo->fifoHead(),32);
+      chSysUnlock();
     }
+    
+    fifo->fifoMoveHead();
+        
+  }
+  
+}
+// the working area for the thread is 512 bytes
+THD_WORKING_AREA(waTXthread,512);
+
+THD_FUNCTION(TXthread, arg){
+
+  rfWorkerBundle *bundle = (rfWorkerBundle*) arg;
+  ThreadState *state = bundle->state;
+  Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
+
+  CanTelemetryMsg *msg;
+
+  uint32_t fifoHead = 0;
+
+  while(!state->terminate){
+   
[…]

### 2021-06-04 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp

@@ -1,74 +1,39 @@
-#include "Fifo.h"
+/**
+ * This file contains the main function for the solar car unit.
+ * 
+ * It sets up three worker threads:
+ *  - the CAN receiver worker, responsible for putting newly received messages in a FIFO
+ *  - the blackbox worker, responsible for adding messages from the FIFO to the logs
+ *  - the radio worker, responsible for streaming messages from the FIFO to the support vehicle
+ * 
+ * The main thread is responsible for listing for commands from the support vehicle and executing them.
+ * 
+ */
+
+#include <Arduino.h>
 #include "RFfunctions.h"
+#include "TelemetryMessages.h"
+#include "support-vehicle/RfThd.h"
 
-// Declare fifoBuffer for RXthread and TXthread 
-Fifo<CanTelemetryMsg> RFinbox(32), RFoutbox(32);
+Fifo<CanTelemetryMsg> RFoutbox(32), RFinbox(32); 
+ThreadState RXthreadState, TXthreadState; 
 
-// 512 byte working stack for RXthread
-THD_WORKING_AREA(waRXthread, 512); 
+void chSetup(){
 
-THD_FUNCTION(RXthread, arg)
-{
-  (void)arg; 
-  radio.startListening(); 
-  while (true){
-    uint8_t pipe; 
-    if (radio.available(&pipe)){
-      radio.read(RFoutbox.fifoTail(), 32);                // Always write data to fifoTail
-      Serial.println("Message received");
-      radio.writeAckPayload(1, RFoutbox.fifoTail(), 32);      
-      RFoutbox.fifoMoveTail();                            // Increment fifoTail index
-      RFoutbox.signalWrite();                             // Signal data write
-    } else {
-      Serial.println("No message received");
-    }
-  }
-}
-
-// 512 byte working stack for TXthread
-THD_WORKING_AREA(waTXthread, 512);
-
-THD_FUNCTION(TXthread, arg)
-{
-  (void)arg; 
-  radio.stopListening();
-  while (true){
-    bool report = radio.write(RFinbox.fifoHead(), 32);    // Always read data to fifoHead
-    if (report){                                          // Message transmission success
-      Serial.print(F("Transmission successful! "));       
-      if (radio.isAckPayloadAvailable()){                 // Get ack payload if available
-        radio.read(RFinbox.fifoHead(), 32); 
-        Serial.print(F("Acknowledge received: "));
[…]

### 2021-06-04 documentation/FinalPaper/paper.tex

@@
-\documentclass[journal]{IEEEtran}
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
 
@@
 \usepackage[pdftex]{graphicx}
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 
@@
 
-\author{Victor Alexander Hansen, Steffan Martin Kunoy, Tjark Petersen}
+\author{\IEEEauthorblockN{Victor Alexander Hansen}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address}
+\and
+\IEEEauthorblockN{Steffan Martin Kunoy}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address}
+\and
+\IEEEauthorblockN{Tjark Petersen}
+\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
+\textit{name of organization (of Aff.)}\\
+City, Country \\
+email address}}
+%\author{Victor Alexander Hansen, Steffan Martin Kunoy, Tjark Petersen}
 

### 2021-06-04 documentation/dear_diary/log_diary.tex

@@
 
+\section{17-03}
+The components for our project arrives.
+
+\section{19-03}
+Our group is granted access to the login for the ecocar wiki page, https://dtucar.com/wiki/index.php?title=Main_Page
+
+\section{22-03}
+The first successful transmission of a signal between the two antennas, not as part of the final setup
+
+\section{25-03}
+The first code for use with CAN data is uploaded.
+Simple "HELLO" was printed at each attempt to transmit.
+Since antennas was not connected, nothing would happen, but nonetheless the code worked as expected.
+
+\section{26-03}
+Source filter to manage what files would be compiled to which boards, blackbox code is made.
+
+\section{07-04}
+The first physical meeting at DTU campus.
+Test transmission between the two Teensy modules with connected antennas successful.
+Attempt to create a test environment in PlatformIO's integrated Test environment unsuccessful.
+Issue created at Github, github.com/platformio/platformio-docs/issues/190.
+User GUI for computer in support car created in Maven project.
+After working on the project at building 324, we had a look around at the DTU ecocar workshop.
+
+\section{16-04}
+Test environment implemented in seperate source folder using AUnit library, this environment tests successfully.
+
+\section{23-04}
+Trouble with transmission and receiving between the two units, code revisited. Protocol for commands through the GUI implemented.
+
+\section{30-04}
+Test file almost finished for files in project
+
+\section{07-05}
+We've discussed earlier that data encryption during transmission would be a useful feature to have during the actual competition, and we therefore decided to make an encryption class that encrypts data.
+The main part of the class is done, but there are some trouble with the pointers.
+We've also worked on FIFO buffers for the module.
+
+\section{08-05}
+The encryption class is done :)
+
+\section{14-05}
+We've met briefly over Discord to discuss the plan going forward in the three week period.
+This will revolve around the entire process of sending, as in, should the logging to blackbox and transmission of data be on the same thread, should the support vehicle be able to request log files, in contrast to making log files itself, in case of transmission disruption etc.
+We believe we are nearing a point where we have made all the small pieces that must now work together.
+
 \end{document}

### 2021-06-04 code/SolarVehicle/lib/rf/RFfunctions.h
minor changes

@@ -24,7 +24,9 @@ static const byte address[][6] = {"00001", "00002"};    // TX/RX byte addresses
 // Takes no arguments.
 void RFinit()
 {
-    radio.begin();
+    if(!radio.begin()){
+        Serial.println("Radio not working!");
+    }
     radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate
     radio.enableDynamicPayloads();

### 2021-06-04 code/SolarVehicle/lib/util/Fifo.h

@@ -84,12 +84,12 @@ class Fifo {
 
     // Increments fifoHead pointer
     public: void fifoMoveHead(){
-        advance(_fifoHead); 
+        _fifoHead = advance(_fifoHead); 
     }
 
     // Increments fifoTail pointer
     public: void fifoMoveTail(){
-        advance(_fifoTail);
+        _fifoTail = advance(_fifoTail);
     }
 
     // Used to check whether fifo is full

### 2021-06-04 code/SolarVehicle/src/solar-car/CanReceiverThd.h

@@ -35,7 +35,7 @@ THD_FUNCTION(canReceiverThd, arg){
   CanTelemetryMsg *msg;
 
   uint32_t count = 0;
-  uint32_t fifoHead = 0;
+  uint32_t fifoWriteIndex = 0;
 
   WITH_MTX(serialMtx){Serial.println("Starting CAN receiver thread...");}
 
@@ -45,16 +45,17 @@ THD_FUNCTION(canReceiverThd, arg){
       state->suspend();
     }
 
-    msg = fifo->get(fifoHead);
+    msg = fifo->get(fifoWriteIndex);
 
     msg->cmd = RECEIVED_CAN;
     msg->data64 = count++;
     msg->id = 0x10A;
 
+    //WITH_MTX(serialMtx){Serial.println("Rx pushed to FIFO...");}
     fifo->signalWrite();
-    fifo->advance(&fifoHead);
+    fifo->advance(&fifoWriteIndex);
 
-    chThdSleepMilliseconds(500);
+    chThdSleepMilliseconds(5);
   }
 }
 

### 2021-06-04 code/SolarVehicle/src/solar-car/RfThd.h

@@ -24,7 +24,7 @@ struct rfWorkerBundle{
 };
 
 // the working area for the thread is 512 bytes
-THD_WORKING_AREA(waRfWorker,512);
+THD_WORKING_AREA(waRfWorker,1024);
 
 THD_FUNCTION(rfWorker, arg){
 
@@ -34,7 +34,7 @@ THD_FUNCTION(rfWorker, arg){
 
   CanTelemetryMsg *msg;
 
-  uint32_t fifoTail = 0;
+  uint32_t fifoReadIndex = 0;
 
   WITH_MTX(serialMtx){Serial.println("Starting RF Transmitting thread...");}
 
@@ -46,7 +46,7 @@ THD_FUNCTION(rfWorker, arg){
       state->suspend();
     }
 
-    msg = fifo->get(fifoTail);
+    msg = fifo->get(fifoReadIndex);
 
     Serial.println("hello");
 
@@ -57,11 +57,12 @@ THD_FUNCTION(rfWorker, arg){
     
     WITH_MTX(rfMTX){
       chSysLock();
-      RFtransmit((BaseTelemetryMsg*)msg,32);
+      Serial.println("Calling rftransmit");
+      MEASURE("RF"){RFtransmit((BaseTelemetryMsg*)msg,32);}
       chSysUnlock();
     }
     
-    fifo->advance(&fifoTail);
+    fifo->advance(&fifoReadIndex);
         
   }
   

### 2021-06-04 code/SolarVehicle/src/support-vehicle/RfThd.h
change can to base message

@@ -11,12 +11,13 @@
 #include "Fifo.h"
 #include "RFfunctions.h"
 #include "support-vehicle/Mutexes.h"
+#include "MutexLocker.h"
 
 /**
  * A bundle used for passsing all relevant resources to the radio thread
  */
 struct rfWorkerBundle{
-    Fifo<CanTelemetryMsg> *fifo;
+    Fifo<BaseTelemetryMsg> *fifo;
     ThreadState *state;
 };
 
@@ -27,7 +28,7 @@ THD_FUNCTION(RXthread, arg){
 
   rfWorkerBundle *bundle = (rfWorkerBundle*) arg;
   ThreadState *state = bundle->state;
-  Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
+  Fifo<BaseTelemetryMsg> *fifo = bundle->fifo;
 
   //CanTelemetryMsg *msg;
 
@@ -49,7 +50,7 @@ THD_FUNCTION(RXthread, arg){
     
     WITH_MTX(rfMTX){
       chSysLock();
-      RFreceive(fifo->fifoHead(),32);
+      RFreceive(fifo->fifoHead());
       chSysUnlock();
     }
     
@@ -65,9 +66,9 @@ THD_FUNCTION(TXthread, arg){
 
   rfWorkerBundle *bundle = (rfWorkerBundle*) arg;
   ThreadState *state = bundle->state;
-  Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
+  Fifo<BaseTelemetryMsg> *fifo = bundle->fifo;
 
-  CanTelemetryMsg *msg;
+  BaseTelemetryMsg *msg;
 
   uint32_t fifoHead = 0;
 
@@ -85,7 +86,7 @@ THD_FUNCTION(TXthread, arg){
     
     WITH_MTX(rfMTX){
       chSysLock();
-      RFtransmit((BaseTelemetryMsg*)msg,32);
+      RFtransmit(msg,32);
       chSysUnlock();
     }
     

### 2021-06-04 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp

@@ -15,7 +15,7 @@
 #include "TelemetryMessages.h"
 #include "support-vehicle/RfThd.h"
 
-Fifo<CanTelemetryMsg> RFoutbox(32), RFinbox(32); 
+Fifo<BaseTelemetryMsg> RFoutbox(32), RFinbox(32); 
 ThreadState RXthreadState, TXthreadState; 
 
 void chSetup(){

### 2021-06-07 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp
message

@@ -24,16 +24,18 @@ void chSetup(){
   rfWorkerBundle RXworkerBundle = {.fifo = &RFoutbox, .state = &RXthreadState};
   chThdCreateStatic(waRXthread,sizeof(waRXthread), NORMALPRIO + 1, RXthread, &RXworkerBundle);
   rfWorkerBundle TXworkerBundle = {.fifo = &RFinbox, .state = &TXthreadState};
-  chThdCreateStatic(waRXthread,sizeof(waTXthread), NORMALPRIO + 1, TXthread, &TXworkerBundle);
+  chThdCreateStatic(waTXthread,sizeof(waTXthread), NORMALPRIO + 2, TXthread, &TXworkerBundle);
 }
 
 void setup()
 {
   Serial.begin(9600); // initialize serial port
   RFinit(); 
+  //chBegin(chSetup);
 }
 
 void loop()
 {
-  
+  BaseTelemetryMsg received; 
+  RFreceive(&received); 
 }

### 2021-06-07 documentation/FinalPaper/paper.tex

@@
 
-\title{Bare Demo of IEEEtran.cls\\ for IEEE Journals}
+\title{Telemetry Module for Solar Vehicle}
 
-\author{\IEEEauthorblockN{Victor Alexander Hansen}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{Steffan Martin Kunoy}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}
-\and
-\IEEEauthorblockN{Tjark Petersen}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address}}
+\author{\IEEEauthorblockN{Victor Alexander Hansen s194027, Steffan Martin Kunoy s194006, Tjark Petersen, s186083}
+\IEEEauthorblockA{\textit{Department of Electrical Engineering} \\
+\textit{Technical Uniservity of Denmark}\\
+Kongens Lyngby, Denmark \\
+31015 Introductory project - electrotechnology\\
+s194027@student.dtu.dk, s194006@student.dtu.dk, s186083@student.dtu.dk}}
 %\author{Victor Alexander Hansen, Steffan Martin Kunoy, Tjark Petersen}
@@
 \begin{abstract}
-The abstract goes here.
+Abstract
 \end{abstract}
@@
 \begin{IEEEkeywords}
-IEEE
+ROAST, telemetry, CAN
 \end{IEEEkeywords}
@@
 \section{Introduction}
-% The very first letter is a 2 line initial drop letter followed
-% by the rest of the first word in caps.
-% 
-% form to use if the first word consists of a single letter:
-% \IEEEPARstart{A}{demo} file is ....
-% 
-% form to use if you need the single drop letter followed by
-% normal text (unknown if ever used by the IEEE):
-% \IEEEPARstart{A}{}demo file is ....
-% 
-% Some journals put the first two words in caps:
-% \IEEEPARstart{T}{his demo} file is ....
-% 
-% Here we have the typical use of a "T" for an initial drop letter
-% and "HIS" in caps to complete the first word.
-\IEEEPARstart{T}{his} demo file is intended to serve as a ``starter file''
-for IEEE journal papers produced under \LaTeX\ using
-IEEEtran.cls version 1.8b and later.
-% You must have at least 2 lines in the paragraph with the drop letter
-% (should never be an issue)
-I wish you the best of success.
 
-\hfill mds
+\IEEEPARstart{I}{n} 2023, DTU Roadrunners Solar Team, abbreviated ROAST, is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in Northern Territory to Adelaide in South Australia.
+The solar car have limited opportunities for recharging during the race, so the main source of energy during the race, is from solar energy.
+To this end, solar panels will cover the car, however, the solar car is only allowed to have a maximum of 4 square meters of solar panels.
+This emphasizes the need to create an energy-efficient solar car to win the race \cite{wsc}.\\
+Throughout the race the solar car will be monitored by a support car, which can monitor the car's condition and issue commands.
+This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
+The support car will therefore be able to analyze and react to the data stream being transmitted from the solar car, even when the driver is preoccupied by driving the car and navigating traffic.
+\\
+The main function of the telemetry module is to read data from the solar car's on-board CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support car, while simultaneously logging data locally in a "black box" memory chip.
+The support car is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus.
+As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering the distance between them may be up to 1 km depending on traffic conditions.\\
+Therefore, the telemetry module must both serve the purpose of fulfilling the necessary specifications for communication, while also being as energy-efficient as possible.
+The project will mainly focus on the former that is, ensuring that the module meets the specifications with a solid solution.
 
-\hfill August 26, 2015
+\subsection{Problem Statement}
+Remote sensing of data from the solar car will give the ROAST team a competitive advantage as the support vehicle will take on a more active role in controlling and optimising the car's performance.
+This will ease the burden on the driver and engage a large part of the supporting crew.
+For these reasons we have chosen to focus our project on the following points:
+\begin{itemize}
+    \item How can we design a module that can read and write sensor data and commands from a CAN bus network? 
+    \item How can CAN data be transmitted securely over a distance of up to one kilometer?
+    \item How does the support-car module process data upon receiving it from the solar car?
+    \item What is the best method for storing data locally (black box) while being accessible at a later time?
+    \item How can the CAN data be made accessible for further processing in e.g. Matlab?
+\end{itemize}
 
-\subsection{Subsection Heading Here}
-Subsection text here.
+\section{Methodology}
+Sprint method and verification of software
+\section{Design}
+\begin{itemize}
+    \item Block diagram
+    \item Components
+    \item Software stack
+    \item Final design
+\end{itemize}
+\subsection{CAN} % Victor
+\subsection{nRF24 \& Shockburst} % Steffan 
+\subsection{Message protocol}
 
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
+\section{Implementation}
 
-\subsubsection{Subsubsection Heading Here}
-Subsubsection text here.
+\subsection{Solar Car} 
+\subsubsection{Black Box} % tjark
+\subsubsection{CAN transceiver}
+\subsection{Support Vehicle}
+\subsubsection{GUI} % tjark
+\subsubsection{RF transceiver}
+
+\section{Verification} % Victor
+Throughout the project we have used the cross-platform IDE PlatformIO, which is a development platform for embedded systems. This has greatly increased the efficiency in our work due to the
+
+\section{Discussion}
 
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
-\appendices
-\section{Proof of the First Zonklar Equation}
-Appendix one text goes here.
-
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-\section{}
-Appendix two text goes here.
-
 % use section* for acknowledgment
@@
 
-The authors would like to thank...
+The authors would like to thank Christian Kampp Kruse for his assistance as coordinator for this project. The authors would also like to thank Claus from DTU Mekanik % Udbedr
 
@@
 
+\section{References}
+\begingroup
+\renewcommand{\section}[2]{}%
+%\renewcommand{\chapter}[2]{}% for other classes
+\begin{thebibliography}{}
+\bibitem{wsc}
+Bridgestone World Solar Challenge, 
+URL: https://worldsolarchallenge.org/
+\bibitem{ROAST}
+Reports and source files provided by DTU ROAST
+\bibitem{teensy}
+Teensy microcontroller,
+URL: https://www.pjrc.com/teensy/
+\bibitem{C++}
+Website for learning C++
+https://www.learn-cpp.org/
+\end{thebibliography}
+\endgroup
+
 \end{document}

### 2021-06-07 code/SolarVehicle/lib/rf/RFfunctions.h
added Poster

@@ -6,6 +6,10 @@
 #include <nRF24L01.h>
 #include "TelemetryMessages.h"
 
+#define IRQ_PIN 2 
+#define CE_PIN 9
+#define CSN_PIN 10
+
 #define COUNT 5     // Number of transmission retries
 #define DELAY 15    // Delay between retries (= DELAY * 250 us + 250 us)
 
@@ -17,22 +21,31 @@ bool radioNumber = 1;
 bool radioNumber = 1;
 #endif
 
-RF24 radio(9, 10);                                      // CE and CSN pins
+RF24 radio(CE_PIN, CSN_PIN);                            // CE and CSN pins
 static const byte address[][6] = {"00001", "00002"};    // TX/RX byte addresses
+bool mode; 
+
+void RFinterrupt()
+{
+    bool tx, fail, rx;
+    radio.whatHappened(tx, fail, rx); 
+
+}
 
 // RFinit() function
 // Takes no arguments.
 void RFinit()
 {
-    if(!radio.begin()){
-        Serial.println("Radio not working!");
-    }
+    if(!radio.begin()){Serial.println("Radio not working!");}
     radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate
     radio.enableDynamicPayloads();
     radio.setAutoAck(true);
     radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
+    pinMode(IRQ_PIN, INPUT);
+    attachInterrupt(digitalPinToInterrupt(IRQ_PIN), RFinterrupt, FALLING); 
+    radio.maskIRQ(1, 1, 0);
     radio.openWritingPipe(address[!radioNumber]);
     radio.openReadingPipe(1, address[radioNumber]);
     radio.startListening();                                     // Starts RX mode
@@ -42,6 +55,7 @@ void RFinit()
 // Takes char array and its size to send. size cannot be greater than 32 bytes (null-terminated)
 bool RFtransmit(BaseTelemetryMsg *msg, uint32_t size)
 {
+    mode = 1; 
     radio.stopListening();                                      // Starts TX mode
     bool report = radio.write(msg, size);                       // Send message and wait for acknowledge
     if (report){    // Checks if message was delivered
@@ -65,6 +79,7 @@ bool RFtransmit(BaseTelemetryMsg *msg, uint32_t size)
[…]

### 2021-06-07 code/SolarVehicle/lib/util/Fifo.h

@@ -53,12 +53,12 @@ class Fifo {
     }
 
     // signals that an element has been read and that the slot is free for new data
-    public: void signalRead(){
+    public: void signalSpace(){
         chSemSignal(&spaceSem);
     }
 
     // signals that new data is present
-    public: void signalWrite(){
+    public: void signalData(){
         chSemSignal(&dataSem);
     }
     
@@ -72,7 +72,7 @@ class Fifo {
         memset(fifo,0,size*sizeof(T));
     }
 
-    // Returns a pointer to element at fifoEnd index
+    // Returns a pointer to element at fifoHead index
     public: T *fifoHead(){
         return &fifo[_fifoHead];    
     }

### 2021-06-07 code/SolarVehicle/src/support-vehicle/RfThd.h

@@ -29,32 +29,23 @@ THD_FUNCTION(RXthread, arg){
   ThreadState *state = bundle->state;
   Fifo<BaseTelemetryMsg> *fifo = bundle->fifo;
 
-  //CanTelemetryMsg *msg;
-
-  //uint32_t fifoTail = 0;
-
-  //WITH_MTX(serialMtx){Serial.println("Starting RF Transmitting thread...");}
-
   while(!state->terminate){
-    //Serial.println("RXthread");
-    //fifo->waitForData();
+    fifo->waitForSpace();
 
     if(state->pause){
       state->suspend();
     }
 
-    //msg = fifo->fifoHead();
+    if(RFreceive(fifo->fifoHead())){
+      fifo->fifoMoveTail();
+      fifo->signalData();
+    }  
 
-    //Serial.println("hello");
-      bool receive = RFreceive(fifo->fifoHead());
-      if(receive){
-        fifo->fifoMoveTail();
-      }  
   }
   
 }
 // the working area for the thread is 1024 bytes
-THD_WORKING_AREA(waTXthread,1024);
+THD_WORKING_AREA(waTXthread, 1024);
 
 THD_FUNCTION(TXthread, arg){
 
@@ -62,9 +53,7 @@ THD_FUNCTION(TXthread, arg){
   ThreadState *state = bundle->state;
   Fifo<BaseTelemetryMsg> *fifo = bundle->fifo;
 
-  BaseTelemetryMsg *msg;
-
-  uint32_t fifoHead = 0;
+  CanTelemetryMsg msg; 
 
   while(!state->terminate){
     fifo->waitForData();
@@ -73,19 +62,26 @@ THD_FUNCTION(TXthread, arg){
       state->suspend();
     }
 
-    msg = fifo->get(fifoHead);
-
-    //Serial.println("good bye");
+    msg.randomize(); 
     
-    WITH_MTX(rfMTX){
-      chSysLock();
-      RFtransmit(msg,32);
-      chSysUnlock();
+    if (RFtransmit(msg.toMessage(), 32)){
+      fifo->fifoMoveHead();
+      fifo->signalSpace();
     }
-    
-    fifo->advance(&fifoHead);
-        
   }
   
+}
+
+THD_WORKING_AREA(waSerialThread, 1024);
+
+THD_FUNCTION(serialThread, arg)
+{
[…]

### 2021-06-07 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp

@@ -8,9 +8,10 @@
  * 
  * The main thread is responsible for listing for commands from the support vehicle and executing them.
  * 
- */
+ **/
 
 #include <Arduino.h>
+#include <SPI.h>
 #include "RFfunctions.h"
 #include "TelemetryMessages.h"
 #include "support-vehicle/RfThd.h"
@@ -18,24 +19,25 @@
 Fifo<BaseTelemetryMsg> RFoutbox(32), RFinbox(32); 
 ThreadState RXthreadState, TXthreadState; 
 
-void chSetup(){
-
+void chSetup()
+{
   chSysInit();
-  rfWorkerBundle RXworkerBundle = {.fifo = &RFoutbox, .state = &RXthreadState};
-  chThdCreateStatic(waRXthread,sizeof(waRXthread), NORMALPRIO + 1, RXthread, &RXworkerBundle);
   rfWorkerBundle TXworkerBundle = {.fifo = &RFinbox, .state = &TXthreadState};
   chThdCreateStatic(waTXthread,sizeof(waTXthread), NORMALPRIO + 2, TXthread, &TXworkerBundle);
+  rfWorkerBundle RXworkerBundle = {.fifo = &RFoutbox, .state = &RXthreadState};
+  chThdCreateStatic(waRXthread,sizeof(waRXthread), NORMALPRIO + 1, RXthread, &RXworkerBundle);
+  rfWorkerBundle serialWorkerBundle[2] = { TXworkerBundle, RXworkerBundle }; 
+  chThdCreateStatic(waSerialThread, sizeof(waSerialThread), NORMALPRIO, serialThread, &serialWorkerBundle); 
 }
 
 void setup()
 {
   Serial.begin(9600); // initialize serial port
   RFinit(); 
-  //chBegin(chSetup);
+  chBegin(chSetup);
 }
 
 void loop()
 {
-  BaseTelemetryMsg received; 
-  RFreceive(&received); 
+
 }

### 2021-06-07 code/SolarVehicle/lib/message/TelemetryMessages.h
Solar Car Update

@@ -152,11 +152,13 @@ class CanTelemetryMsg {
         stamp();
     }
     public: void update(CANMessage *msg){
-        id = msg->id;
-        rtr = msg->rtr;
-        len = msg->len;
-        data64 = msg->data64;
+        chSysLock();
+        this->id = msg->id;
+        this->rtr = msg->rtr;
+        this->len = msg->len;
+        this->data64 = msg->data64;
         stamp();
+        chSysUnlock();
     }
 
     public: void stamp(){

### 2021-06-07 code/SolarVehicle/lib/rf/RFfunctions.h

@@ -35,6 +35,7 @@ void RFinit()
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
     radio.openWritingPipe(address[!radioNumber]);
     radio.openReadingPipe(1, address[radioNumber]);
+    //radio.maskIRQ(1,1,0);
     radio.startListening();                                     // Starts RX mode
 }
 
@@ -69,9 +70,6 @@ bool RFreceive(BaseTelemetryMsg *received)
     uint8_t pipe;
     if (radio.available(&pipe)){                                // Check if transmitter is sending message
         radio.read(received, 32);                               // Read message, cannot be larger than 32 bytes (null-terminated)
-        char str[64]; 
-        received->toString(str, sizeof(str));
-        Serial.println(str); 
         /*
         if (received->cmd == RECEIVED_CAN){
             char tmp[64];

### 2021-06-07 code/SolarVehicle/src/solar-car/CanReceiverThd.h

@@ -23,7 +23,7 @@ struct CanReceiverBundle{
 };
 
 // the working area for the thread is 256 bytes
-THD_WORKING_AREA(waCanReceiver,256);
+THD_WORKING_AREA(waCanReceiver,1024);
 
 THD_FUNCTION(canReceiverThd, arg){
 
@@ -34,7 +34,6 @@ THD_FUNCTION(canReceiverThd, arg){
   CANMessage frame;
   CanTelemetryMsg *msg;
 
-  uint32_t count = 0;
   uint32_t fifoWriteIndex = 0;
 
   WITH_MTX(serialMtx){Serial.println("Starting CAN receiver thread...");}
@@ -45,17 +44,19 @@ THD_FUNCTION(canReceiverThd, arg){
       state->suspend();
     }
 
-    msg = fifo->get(fifoWriteIndex);
 
-    msg->cmd = RECEIVED_CAN;
-    msg->data64 = count++;
-    msg->id = 0x10A;
+    if(ACAN::can0.receive(frame)){
+      chSysLock();
+      msg = fifo->get(fifoWriteIndex);
+      msg->update(&frame);
+      fifo->signalWrite();
+      fifo->advance(&fifoWriteIndex);
+      Serial.print("Received CAN [");Serial.print(fifoWriteIndex);Serial.println("]");
+      chSysUnlock();
+    }
 
-    //WITH_MTX(serialMtx){Serial.println("Rx pushed to FIFO...");}
-    fifo->signalWrite();
-    fifo->advance(&fifoWriteIndex);
+    chThdYield();
 
-    chThdSleepMilliseconds(1000);
   }
 }
 

### 2021-06-07 code/SolarVehicle/src/solar-car/RfTxThd.h

@@ -3,21 +3,23 @@
  * messages to the support vehicle.
  */
 
-#ifndef __RF_THD_H__
-#define __RF_THD_H__
+#ifndef __RF_TX_THD_H__
+#define __RF_TX_THD_H__
 
 #include <ChRt.h>
 
 #include "ThreadState.h"
 #include "Fifo.h"
 #include "RFfunctions.h"
+#include "MutexLocker.h"
+#include "Measure.h"
 
 #include "solar-car/Mutexes.h"
 
 /**
  * A bundle used for passsing all relevant resources to the radio thread
  */
-struct rfWorkerBundle{
+struct rfTxWorkerBundle{
     Fifo<CanTelemetryMsg> *fifo;
     ThreadState *state;
 };
@@ -27,7 +29,7 @@ THD_WORKING_AREA(waRfWorker,1024);
 
 THD_FUNCTION(rfWorker, arg){
 
-  rfWorkerBundle *bundle = (rfWorkerBundle*) arg;
+  rfTxWorkerBundle *bundle = (rfTxWorkerBundle*) arg;
   ThreadState *state = bundle->state;
   Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
 
@@ -46,20 +48,11 @@ THD_FUNCTION(rfWorker, arg){
     }
 
     msg = fifo->get(fifoReadIndex);
-
-    Serial.println("hello");
-
-    WITH_MTX(serialMtx){
-      Serial.print("rf worker msg nr. ");
-      Serial.println((uint32_t)msg->data64);
-    }
     
-    WITH_MTX(rfMTX){
-      chSysLock();
-      Serial.println("Calling rftransmit");
-      MEASURE("RF"){RFtransmit((BaseTelemetryMsg*)msg,32);}
-      chSysUnlock();
-    }
+    chSysLock();
+    Serial.println("Calling rftransmit");
+    MEASURE("RF"){RFtransmit((BaseTelemetryMsg*)msg,32);}
+    chSysUnlock();
     
     fifo->advance(&fifoReadIndex);
         

### 2021-06-07 code/SolarVehicle/src/solar-car/SystemThd.h

@@ -1,0 +1,81 @@
+/**
+ * This file contains the function for the system thread, which receives commands from the
+ * radio module and configures the system accordingly
+ */
+
+#ifndef __SYSTEM_THD_H__
+#define __SYSTEM_THD_H__
+
+#include <ChRt.h>
+
+#include "ThreadState.h"
+#include "Fifo.h"
+#include "RFfunctions.h"
+#include "MutexLocker.h"
+#include "Measure.h"
+
+#include "solar-car/Mutexes.h"
+
+/**
+ * A bundle used for passsing all relevant resources to the radio thread
+ */
+struct systemThdBundle{
+    ThreadState *canReceiverState;
+    ThreadState *blackBoxWorkerState;
+    ThreadState *rfWorkerState;
+};
+
+// the working area for the thread is 1024 bytes
+THD_WORKING_AREA(waSystemThd,1024);
+
+THD_FUNCTION(systemThd, arg){
+
+  systemThdBundle *bundle = (systemThdBundle*) arg;
+  ThreadState *canReceiverState = bundle->canReceiverState;
+  ThreadState *rfWorkerState = bundle->rfWorkerState;
+  ThreadState *blackBoxWorkerState = bundle->blackBoxWorkerState;
+
+  BaseTelemetryMsg msg;
+
+  Serial.println("Starting System Thread...");
+
+  chThdSleepMicroseconds(100); // release in order to allow creation of other threads
+
+  while(true){
+   
+    if(RFreceive(&msg)){
+      char str[64];
+      msg.toString(str,64);
+      Serial.println(str);
+    }
+
+    if(Serial.available()){
+      uint8_t read = Serial.read();
+      switch (read){
+        case 13:
+          Serial.println("\n\nStopping...");
+          blackBoxWorkerState->terminate = true;
+          canReceiverState->terminate = true;
+          rfWorkerState->terminate = true;
+          while(true){}
+          break;
+
+        case 'p':
+          if(!canReceiverState->pause){
+            Serial.println("Pausing...");
+            canReceiverState->pause = true;
+          } else {
+            Serial.println("Resuming...");
+            canReceiverState->wakeUp();
+          }
+          break;
+        default:
+          break;
+      }
+    }
+        
+  }
+  
+}
+
+#endif

### 2021-06-07 code/SolarVehicle/src/solar-car/blackBoxThd.h

@@ -24,7 +24,7 @@ struct BlackboxWorkerBundle{
 };
 
 // the working area for the thread is 512 bytes
-THD_WORKING_AREA(blackBoxWorker,512);
+THD_WORKING_AREA(blackBoxWorker,1024);
 
 THD_FUNCTION(canWorkerFunc, arg){
 

### 2021-06-07 code/SolarVehicle/src/solar-car/solarCarMain.cpp

@@ -23,18 +23,11 @@
 #include "solar-car/blackBoxThd.h"
 #include "solar-car/CanReceiverThd.h"
 #include "solar-car/Mutexes.h"
-#include "solar-car/RfThd.h"
+#include "solar-car/RfTxThd.h"
+#include "solar-car/SystemThd.h"
 
-/**
- * Singleton containing global state variables
- */
-class GlobalState {
-  public:
-    uint8_t isLogging = true;
-    uint8_t isTransmittingRT = true;
-} globalState;
 
-Fifo<CanTelemetryMsg> canFifo(32);
+Fifo<CanTelemetryMsg> canFifo(64);
 BlackBox bb;
 ThreadState blackBoxWorkerState;
 ThreadState canReceiverState;
@@ -53,18 +46,19 @@ void setup(){
   Serial.println("Initializing...");
 
   // setup CAN bus
-  //ACANSettings settings(125 * 1000);
-  //if(ACAN::can0.begin(settings)){Serial.println("CAN setup failed!");}
+  ACANSettings settings(125 * 1000);
+  if(ACAN::can0.begin(settings) != 0){Serial.println("CAN setup failed!");}
 
   // setup radio module
   RFinit();
 
+  bb.init();
+
   canFifo.clear();
 
-  Serial.println("Starting...");  
+  Serial.println("Starting...");
 
   chBegin(chSetup);
-
 }
 
 void chSetup(){
@@ -73,38 +67,14 @@ void chSetup(){
   
   // create the three worker threads
   BlackboxWorkerBundle blackBoxWorkerBundle = {.fifo = &canFifo, .state = &blackBoxWorkerState, .bb = &bb};
-  chThdCreateStatic(blackBoxWorker, sizeof(blackBoxWorker), NORMALPRIO+1, canWorkerFunc, &blackBoxWorkerBundle);
-  rfWorkerBundle rfWorkerBundle = {.fifo = &canFifo, .state = &rfWorkerState};
-  chThdCreateStatic(waRfWorker,sizeof(waRfWorker),NORMALPRIO + 1, rfWorker, &rfWorkerBundle);
+  chThdCreateStatic(blackBoxWorker, sizeof(blackBoxWorker), NORMALPRIO + 3, canWorkerFunc, &blackBoxWorkerBundle);
+  rfTxWorkerBundle rfWorkerBundle = {.fifo = &canFifo, .state = &rfWorkerState};
[…]

### 2021-06-07 code/SolarVehicle/lib/message/TelemetryMessages.h
steffan changes

@@ -55,6 +55,12 @@ class BaseTelemetryMsg {
 
   public: uint8_t* toBytes(){return (uint8_t*)this;}
 
+  public: void randomize(){
+      this->cmd = (cmd_t) random(0, 0x99);   
+      for (int i = 0; i < 31; i++){
+        this->data[i] = random(0,0x99); 
+      }  
+  }
 
 }__attribute__((packed));
 
@@ -109,7 +115,6 @@ class CanTelemetryMsg {
     }
 
     public: BaseTelemetryMsg* toMessage(){return (BaseTelemetryMsg*)this;}
-
     
 
     public: void randomize(){

### 2021-06-07 code/SolarVehicle/lib/rf/RFfunctions.h

@@ -23,14 +23,6 @@ bool radioNumber = 1;
 
 RF24 radio(CE_PIN, CSN_PIN);                            // CE and CSN pins
 static const byte address[][6] = {"00001", "00002"};    // TX/RX byte addresses
-bool mode; 
-
-void RFinterrupt()
-{
-    bool tx, fail, rx;
-    radio.whatHappened(tx, fail, rx); 
-
-}
 
 // RFinit() function
 // Takes no arguments.
@@ -43,9 +35,9 @@ void RFinit()
     radio.setAutoAck(true);
     radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
-    pinMode(IRQ_PIN, INPUT);
-    attachInterrupt(digitalPinToInterrupt(IRQ_PIN), RFinterrupt, FALLING); 
-    radio.maskIRQ(1, 1, 0);
+    //pinMode(IRQ_PIN, INPUT);
+    //attachInterrupt(digitalPinToInterrupt(IRQ_PIN), RFinterrupt, FALLING); 
+    //radio.maskIRQ(1, 1, 0);
     radio.openWritingPipe(address[!radioNumber]);
     radio.openReadingPipe(1, address[radioNumber]);
     radio.startListening();                                     // Starts RX mode
@@ -55,7 +47,6 @@ void RFinit()
 // Takes char array and its size to send. size cannot be greater than 32 bytes (null-terminated)
 bool RFtransmit(BaseTelemetryMsg *msg, uint32_t size)
 {
-    mode = 1; 
     radio.stopListening();                                      // Starts TX mode
     bool report = radio.write(msg, size);                       // Send message and wait for acknowledge
     if (report){    // Checks if message was delivered
@@ -79,14 +70,13 @@ bool RFtransmit(BaseTelemetryMsg *msg, uint32_t size)
 // Takes no arguments, prints received message to serial
 bool RFreceive(BaseTelemetryMsg *received)
 {
-    mode = 0;
     radio.startListening();                                     // Starts RX mode
     uint8_t pipe;
     if (radio.available(&pipe)){                                // Check if transmitter is sending message
         radio.read(received, 32);                               // Read message, cannot be larger than 32 bytes (null-terminated)
-        char str[64]; 
+        /*char str[64]; 
         received->toString(str, sizeof(str));
-        Serial.println(str); 
+        Serial.println(str); */
         radio.writeAckPayload(1, received, 32);                 // Send acknowledge payload
         return true;
     }

### 2021-06-07 code/SolarVehicle/lib/util/Fifo.h

@@ -2,7 +2,6 @@
 #define __FIFO_H__
 
 #include <ChRt.h>
-
 /**
  * A Fifo with built in semaphores
  */
@@ -11,16 +10,16 @@ class Fifo {
     private: 
         semaphore_t spaceSem, dataSem;
         T *fifo;
-        size_t _fifoHead, _fifoTail;  
+        size_t _head, _tail;  
         uint32_t size;
 
     public: Fifo(uint32_t size){
         this->size  = size;
+        this->_head   = 0;
+        this->_tail   = 0;
         spaceSem    = _SEMAPHORE_DATA(spaceSem,(cnt_t)size);
         dataSem     = _SEMAPHORE_DATA(dataSem,0);
         fifo        = new T[size];
-        _fifoHead   = 0;
-        _fifoTail   = 0;
     }
 
     public: ~Fifo(){
@@ -71,35 +70,43 @@ class Fifo {
     public: void clear(){
         memset(fifo,0,size*sizeof(T));
     }
-
-    // Returns a pointer to element at fifoHead index
-    public: T *fifoHead(){
-        return &fifo[_fifoHead];    
+    
+    // Prints the contents of the fifo buffer for debugging
+    public: void printContents(){
+        for (int i = 0; i < size; i++){
+            Serial.print(i + " : " + fifo[i]);
+        }
+        Serial.println();
     }
 
-    // Returns a pointer to element at fifoTail index
-    public: T *fifoTail(){
-        return &fifo[_fifoTail];
+    // Returns a pointer to element at head index
+    public: T *head(){
+        return &fifo[_head];    
+    }
+
+    // Returns a pointer to element at tail index
+    public: T *tail(){
+        return &fifo[_tail];
     }
 
     // Increments fifoHead pointer
-    public: void fifoMoveHead(){
-        _fifoHead = advance(_fifoHead); 
+    public: void moveHead(){
+        _head = advance(_head); 
     }
 
-    // Increments fifoTail pointer
-    public: void fifoMoveTail(){
-        _fifoTail = advance(_fifoTail);
+    // Increments tail pointer
+    public: void moveTail(){
+        _tail = advance(_tail);
     }
 
     // Used to check whether fifo is full
-    public: bool fifoFull(){
-        return advance(_fifoTail) == _fifoHead;
+    public: bool full(){
+        return advance(_tail) == _head;
     }
 
     // Used to check whether fifo is empty
-    public: bool fifoEmpty(){
[…]

### 2021-06-07 code/SolarVehicle/src/support-vehicle/Mutexes.h

@@ -1,12 +1,0 @@
-/**
- * This file contains mutex declarations for all the resources shared by multiple threads
- */
-
-#ifndef __MUTEXES_H__
-#define __MUTEXES_H__
-
-#include <ChRt.h>
-
-MUTEX_DECL(rfMTX);
-
-#endif

### 2021-06-07 code/SolarVehicle/src/support-vehicle/RfThd.h

@@ -1,87 +1,129 @@
-/**
- * This file contains the function for the thread responsible for streaming CAN
- * messages to the support vehicle.
- */
-
 #ifndef __RF_THD_H__
 #define __RF_THD_H__
 
 #include <ChRt.h>
 #include "ThreadState.h"
-#include "Fifo.h"
 #include "RFfunctions.h"
-#include "support-vehicle/Mutexes.h"
-#include "MutexLocker.h"
-/**
- * A bundle used for passsing all relevant resources to the radio thread
- */
-struct rfWorkerBundle{
+#include "Fifo.h"
+
+// A bundle used for passsing all relevant resources to each thread
+struct threadBundle
+{
     Fifo<BaseTelemetryMsg> *fifo;
     ThreadState *state;
 };
 
-// the working area for the thread is 1024 bytes
-THD_WORKING_AREA(waRXthread,1024);
+// the working area for the radioReceiverThread is 1024 bytes
+THD_WORKING_AREA(waRadioReceiverThread,1024);
 
-THD_FUNCTION(RXthread, arg){
+THD_FUNCTION(radioReceiverThread, arg){
 
-  rfWorkerBundle *bundle = (rfWorkerBundle*) arg;
-  ThreadState *state = bundle->state;
-  Fifo<BaseTelemetryMsg> *fifo = bundle->fifo;
+  threadBundle *radioReceiverBundle = (threadBundle*) arg;
+  ThreadState *radioReceiverState = radioReceiverBundle->state;
+  Fifo<BaseTelemetryMsg> *radioReceiverFifo = radioReceiverBundle->fifo;
+  while(!radioReceiverState->terminate){
+    Serial.println("radioReceiverThread");
+    //radioReceiverFifo->waitForSpace();
 
-  while(!state->terminate){
-    fifo->waitForSpace();
-
-    if(state->pause){
-      state->suspend();
+    if(radioReceiverState->pause){
+      radioReceiverState->suspend();
     }
 
-    if(RFreceive(fifo->fifoHead())){
-      fifo->fifoMoveTail();
-      fifo->signalData();
+    if(RFreceive(radioReceiverFifo->head())){
+      radioReceiverFifo->moveTail();
+      radioReceiverFifo->signalData();
     }  
-
-  }
-  
-}
-// the working area for the thread is 1024 bytes
-THD_WORKING_AREA(waTXthread, 1024);
-
[…]

### 2021-06-07 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp

@@ -1,43 +1,33 @@
-/**
- * This file contains the main function for the solar car unit.
- * 
- * It sets up three worker threads:
- *  - the CAN receiver worker, responsible for putting newly received messages in a FIFO
- *  - the blackbox worker, responsible for adding messages from the FIFO to the logs
- *  - the radio worker, responsible for streaming messages from the FIFO to the support vehicle
- * 
- * The main thread is responsible for listing for commands from the support vehicle and executing them.
- * 
- **/
-
 #include <Arduino.h>
 #include <SPI.h>
-#include "RFfunctions.h"
-#include "TelemetryMessages.h"
 #include "support-vehicle/RfThd.h"
 
 Fifo<BaseTelemetryMsg> RFoutbox(32), RFinbox(32); 
-ThreadState RXthreadState, TXthreadState; 
+ThreadState radioReceiverState, radioTransmitterState, serialReceiverState, serialTransmitterState; 
 
 void chSetup()
 {
   chSysInit();
-  rfWorkerBundle TXworkerBundle = {.fifo = &RFinbox, .state = &TXthreadState};
-  chThdCreateStatic(waTXthread,sizeof(waTXthread), NORMALPRIO + 2, TXthread, &TXworkerBundle);
-  rfWorkerBundle RXworkerBundle = {.fifo = &RFoutbox, .state = &RXthreadState};
-  chThdCreateStatic(waRXthread,sizeof(waRXthread), NORMALPRIO + 1, RXthread, &RXworkerBundle);
-  rfWorkerBundle serialWorkerBundle[2] = { TXworkerBundle, RXworkerBundle }; 
-  chThdCreateStatic(waSerialThread, sizeof(waSerialThread), NORMALPRIO, serialThread, &serialWorkerBundle); 
+  threadBundle radioReceiverBundle = {.fifo = &RFinbox, .state = &radioReceiverState};
+  chThdCreateStatic(waRadioReceiverThread, sizeof(waRadioReceiverThread), NORMALPRIO + 1, radioReceiverThread, &radioReceiverBundle);
+  threadBundle radioTransmitterBundle = {.fifo = &RFoutbox, .state = &radioTransmitterState};
+  chThdCreateStatic(waRadioTransmitterThread, sizeof(waRadioTransmitterThread), NORMALPRIO + 2, radioTransmitterThread, &radioTransmitterBundle);
[…]

### 2021-06-07 code/SolarVehicle/src/support-vehicle/RfThd.h
v2

@@ -21,6 +21,7 @@ THD_FUNCTION(radioReceiverThread, arg){
   threadBundle *radioReceiverBundle = (threadBundle*) arg;
   ThreadState *radioReceiverState = radioReceiverBundle->state;
   Fifo<BaseTelemetryMsg> *radioReceiverFifo = radioReceiverBundle->fifo;
+  chThdSleepMilliseconds(100); 
   while(!radioReceiverState->terminate){
     Serial.println("radioReceiverThread");
     //radioReceiverFifo->waitForSpace();
@@ -33,7 +34,6 @@ THD_FUNCTION(radioReceiverThread, arg){
       radioReceiverFifo->moveTail();
       radioReceiverFifo->signalData();
     }  
-    chThdSleepMilliseconds(1000);
   }
   
 }

### 2021-06-07 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp

@@ -29,5 +29,6 @@ void setup()
 
 void loop()
 {
+  
   /* Not used */
 }

### 2021-06-11 code/TelemetryUI/src/main/scala/telemetryui/types/CanFrame.scala
debug padto

@@ -8,6 +8,7 @@ case class CanFrame(id: BigInt, rtr: Boolean, len: Int, data: BigInt, stamp: Tim
   override def toString: String = s"($id: [$len]$data at $stamp)"
   def toByteArray : Array[Byte] = {
     //TODO: padTo seems to create arrays that vary in size
+    println(data.toByteArray.reverse.padTo(8,0.toByte).mkString(", "))
     (id.toByteArray.reverse.padTo(2,0.toByte) ++ Array((if(rtr)1 else 0).toByte) ++ Array(len.toByte) ++ data.toByteArray.reverse.padTo(8,0.toByte) ++
       stamp.toByteArray).padTo(31,0.toByte)
   }

### 2021-06-11 code/matlab/udpClient.m
update udpClient.m

@@ -1,5 +1,5 @@
 %echoudp("on",3030);
-u = udpport("datagram","OutputDatagramSize",32);
+u = udpport("datagram","OutputDatagramSize",16);
 
 write(u,"start","uint8","127.0.0.1",4445);
 
@@ -18,9 +18,9 @@ for i = 1:10
        can.data64 = bitor(can.data64,bitshift(d(k+5),8*k));
     end
     can.data = d(5:12);
-    can.t.hour = d(13);
+    can.t.hour = d(15);
     can.t.min = d(14);
-    can.t.sec = d(15);
+    can.t.sec = d(13);
     can
     can.t
 end

### 2021-06-11 code/TelemetryUI/src/main/scala/telemetryui/serial/SerialWorker.scala
mutex

@@ -3,6 +3,7 @@ package telemetryui.serial
 import com.fazecast.jSerialComm.SerialPort
 import telemetryui.types.CMD.RECEIVED_CAN
 import telemetryui.types.{CanFrame, TelemetryMessage}
+import java.util.concurrent.Semaphore
 
 import java.util.concurrent.ConcurrentLinkedQueue
 
@@ -12,11 +13,12 @@ class SerialWorker(port: SerialPort,
                    errListener: Seq[() => Unit]
                   ) extends Thread {
 
+  private val serialMutex = new Semaphore(1)
+
   this.setDaemon(true)
 
   // queue used for passing raw data from the event driven routine to a processing routine
   private val queue = new ConcurrentLinkedQueue[Array[Byte]]()
-  private val outBox = new ConcurrentLinkedQueue[Array[Byte]]()
 
   // class assembling the json string with basic error checking
   private val assembler = new JsonAssembler
@@ -31,7 +33,7 @@ class SerialWorker(port: SerialPort,
 
   override def run(): Unit = {
 
-    port.addDataListener(new SerialPortListener(port,queue))
+    port.addDataListener(new SerialPortListener(port,serialMutex,queue))
 
     while(running){
 
@@ -51,13 +53,9 @@ class SerialWorker(port: SerialPort,
   }
 
   def send(msg: TelemetryMessage): Unit = {
-    port.synchronized{
-      val bytes = msg.toByteArray
-      port.writeBytes(bytes,bytes.length)
-      /*
-      val out = port.getOutputStream
-      out.write(bytes)
-      out.close()*/
-    }
+    serialMutex.acquireUninterruptibly()
+    val bytes = msg.toByteArray
+    port.writeBytes(bytes,bytes.length)
+    serialMutex.release()
   }
 }

### 2021-06-11 code/TelemetryUI/src/main/scala/telemetryui/types/CanFrame.scala

@@ -2,22 +2,22 @@ package telemetryui.types
 
 import org.json4s.DefaultFormats
 import org.json4s.jackson.JsonMethods.parse
-
+import math.pow
 
 case class CanFrame(id: BigInt, rtr: Boolean, len: Int, data: BigInt, stamp: TimeStamp) {
   override def toString: String = s"($id: [$len]$data at $stamp)"
   def toByteArray : Array[Byte] = {
     //TODO: padTo seems to create arrays that vary in size
     println(data.toByteArray.reverse.padTo(8,0.toByte).mkString(", "))
-    (id.toByteArray.reverse.padTo(2,0.toByte) ++ Array((if(rtr)1 else 0).toByte) ++ Array(len.toByte) ++ data.toByteArray.reverse.padTo(8,0.toByte) ++
-      stamp.toByteArray).padTo(31,0.toByte)
+    (id.toByteArray.reverse.padTo(2,0.toByte) ++ Array((if(rtr)1 else 0).toByte) ++ Array(len.toByte) ++
+      (if((data & 0x8000000000000000L) == 0) data.toByteArray.reverse.padTo(8,0.toByte) else data.toByteArray.drop(1).reverse.padTo(8,0.toByte)) ++
+      stamp.toByteArray)//.padTo(31,0.toByte)
   }
 }
 
 object CanFrame {
   def apply(): CanFrame = {
     val rand = scala.util.Random
-    CanFrame(rand.nextInt(1000),rand.nextBoolean(),rand.nextInt(8),rand.nextInt(Int.MaxValue),TimeStamp())
+    CanFrame(rand.nextInt(pow(2,11).toInt),rand.nextBoolean(),rand.nextInt(8),0xFFFFFFFFFFFFFFFFL-1,TimeStamp())
   }
 }
-

### 2021-06-11 code/SolarVehicle/test/test_common_solar/test_solar.cpp
Test - added encryption test

@@ -21,11 +21,27 @@
 #include <sys/stat.h>
 #include <ctime>
 #include <fstream>
+#include <ChRt.h>
 
 #include "BlackBox.h"
+#include "TelemetryMessages.h"
+#include "Encryption.h"
+#include "RFfunctions.h"
+#include "MultiReaderFifo.h"
+#include "ThreadState.h"
+#include "solar-car/blackBoxThd.h"
+#include "solar-car/CanReceiverThd.h"
+#include "solar-car/Mutexes.h"
+#include "solar-car/RfTxThd.h"
+#include "solar-car/SystemThd.h"
 
 CANMessage frame;
 BlackBox box;
+Security security;
+MultiReaderFifo<CanTelemetryMsg> canFifo(64,2);
+ThreadState blackBoxWorkerState;
+ThreadState canReceiverState;
+ThreadState rfWorkerState;
 
 // void setUp(void) {
 // // set stuff up here
@@ -35,21 +51,14 @@ BlackBox box;
 // // clean stuff up here
 // }
 
-void test_add(void) {
+void test_sanity(void) {
     TEST_ASSERT_EQUAL(32, 25 + 7);
-}
-
-void test_subtract(void) {
     TEST_ASSERT_EQUAL(20, 23 - 3);
-}
-
-void test_multiplication(void) {
     TEST_ASSERT_EQUAL(50, 25 * 2);
-}
-
-void test_division(void) {
     TEST_ASSERT_EQUAL(32, 96 / 3);
 }
+
+// --------------------------------- setup --------------------------------- //
 
 // Should test that both CAN0 port and CAN1 port is active
 void test_can_x_begin(void) {
@@ -60,6 +69,114 @@ void test_can_x_begin(void) {
     TEST_ASSERT_TRUE(can1_begin);
 }
 
+// -------------------------- TelemetryMessages.h -------------------------- //
+
+//uint32_t toString(char *buf, uint32_t len, uint8_t base)
+
+//uint32_t toString(char *buf, uint32_t len)
+
+//static const String getHeader()
+
+// Should set data in CAN frame
+//void toCanFrame(CANMessage *msg)
+
+// Should update time stamp on CAN message
+//void update(CANMessage *msg)
+
+// Should set correct time in seconds, minutes and hours
+//void stamp()
+
+// ------------------------------ Encryption.h ----------------------------- //
+
+// Should test that an encryption array from Security class is made
[…]

### 2021-06-11 code/SolarVehicle/lib/util/MultiReaderFifo.h
update tasks and implement fifo write overtake handling

@@ -58,13 +58,16 @@ class MultiReaderFifo {
         chSemWait(&(dataSems[readerId]));
     }
 
-    public: void waitForData(){
+    public: void waitForData() {
         chSemWait(&(dataSems[0]));
     }
 
     public: void moveTail() {
         writeIndex = advance(writeIndex); // TODO: push read indices when they are overtaken
-
+        for(uint32_t i = 0; i < readers; i++) {
+            if(writeIndex == readIndices[i])
+                readIndices[i] = advance(readIndices[i]);
+        }
     }
 
     public: void moveHead(uint32_t readerId) {

### 2021-06-11 documentation/FinalPaper/paper.tex

@@
 
+% Regler ti skrivning: 
+% 1. Altid nutid
+% 2. Brug 1. persons flertal stedord (We/us/our/ours)
+% 3. Skriv formelt og uden sammentrækninger (don't do this) 
+%
+
 \begin{document}
@@
 \textit{Technical Uniservity of Denmark}\\
-Kongens Lyngby, Denmark \\
+Kgs. Lyngby, Denmark \\
 31015 Introductory project - electrotechnology\\
@@
 \IEEEPARstart{I}{n} 2023, DTU Roadrunners Solar Team, abbreviated ROAST, is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in Northern Territory to Adelaide in South Australia.
-The solar car have limited opportunities for recharging during the race, so the main source of energy during the race, is from solar energy.
+The solar car has limited opportunities for recharging during the race, so the main energy source during the race will be the sun.
 To this end, solar panels will cover the car, however, the solar car is only allowed to have a maximum of 4 square meters of solar panels.
 This emphasizes the need to create an energy-efficient solar car to win the race \cite{wsc}.\\
-Throughout the race the solar car will be monitored by a support car, which can monitor the car's condition and issue commands.
+Throughout the race the solar car will be monitored by a support vehicle, which can monitor the car's condition and should be able to issue commands to the solar car's internal systems.
 This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
@@
 \\
-The main function of the telemetry module is to read data from the solar car's on-board CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support car, while simultaneously logging data locally in a "black box" memory chip.
-The support car is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus.
-As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering the distance between them may be up to 1 km depending on traffic conditions.\\
+The main function of the telemetry module is to read data from the solar car's on-board CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support vehicle, while simultaneously logging data in a local storage acting as a "black box".
+The support vehicle is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus.
+As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering that the distance between them may be up to 1 km depending on traffic conditions.\\
 Therefore, the telemetry module must both serve the purpose of fulfilling the necessary specifications for communication, while also being as energy-efficient as possible.
-The project will mainly focus on the former that is, ensuring that the module meets the specifications with a solid solution.
+This project will mainly focus on the former, that is, ensuring that the module meets the specifications with a solid solution.
 
 \subsection{Problem Statement}
-Remote sensing of data from the solar car will give the ROAST team a competitive advantage as the support vehicle will take on a more active role in controlling and optimising the car's performance.
-This will ease the burden on the driver and engage a large part of the supporting crew.
+Remote sensing of data from the solar car will give the ROAST team a competitive advantage as the support vehicle will be able to take on a more active role in managing and optimising the car's performance.
+This will ease the burden on the driver and leverage the expertise of the supporting crew.
 For these reasons we have chosen to focus our project on the following points:
@@
     \item What is the best method for storing data locally (black box) while being accessible at a later time?
-    \item How can the CAN data be made accessible for further processing in e.g. Matlab?
+    \item How can the CAN data be made accessible in the support vehicle for further processing e.g. in Matlab?
 \end{itemize}
@@
 \section{Methodology}
-Sprint method and verification of software
+The project methodology has largely consisted of making a preliminary design to be completed in a series of design sprints.
+At first a detailed specification was drafted which then served as a reevaluation tool and an overall guidance each time a sprint was completed and when the goal for the next sprint had to be selected.
+Taking a starting point in this, we made a prototype setup of the hardware on breadboards as our first sprint.
+The following sprints would then mostly be conducted in software, such as making a first transmission and receiving of data, and finally a last hardware sprint where we soldered our hardware to a perf board.\\
+During the coding of the software functions, we also made verification and validation of our code design.
+In verification we would discuss code naming and functionality.
+To validate our code we made unit tests that would test the actual behaviour of our code against the intended behaviour.\\
+During the coding phase, we attempted to make our debugging more efficient with the use of a debugger.
+However, this would prove quite troublesome, as the Teensy board does not have an on-board debugger and in general is not very suited for debuggers.
+If we would insist on using a debugger such as st-link, we would have to hold or solder wires to the debug ports on the Teensy board's backside, which would complicate our prototype setup.
+Therefore, we decided not to spent more time on finding a debugger solution.
+This was a clear setback, as we would spent a lot of time debugging our code.
 \section{Design}
-\begin{itemize}
-    \item Block diagram
-    \item Components
-    \item Software stack
-    \item Final design
-\end{itemize}
+Overall specification?
+
+Block diagram
+
+\subsection{Components}
+
+\subsection{Software Stack}
+
+In this section the main software tools and libraries used throughout the project are introduced.
+
+\subsubsection{PlatformIO}
+The Teensy micro-processor line-up is integrated into the Arduino framework and implements as such the Arduino standard API.
+As a cross-platform build tool, PlatformIO is a more potent alternative to the standard Arduino IDE including more advanced features such as a static code analyzer and even a remote unit testing framework.
+
+\subsubsection{ChibiOS}
+An open-source RTOS for embedded systems including Teensy. Supports threads, semaphores, data-sharing and timers to ease the scheduling of different tasks by the program kernel.
+
+\subsubsection{Other Libraries}
+
 \subsection{CAN} % Victor
-\subsection{nRF24 \& Shockburst} % Steffan 
+The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer.
+This is possible with the special structure of the CAN message.\\
+\begin{figure}[h]
+    %\centering
+    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
+    \caption{Structure of a CAN frame}
+    \label{fig:CANframe}
+\end{figure}\\
+The frame uses an identifier to tell which messages goes to which unit.
+When two messages tries to access the bus, the one with the highest priority, lowest ID, will be granted access to the bus, while the other message has to wait.
+After the ID comes the RTR bit which determines whether the frame is set to be at sent or received by the node.
+The control field, sometimes called DLC, tells the amount of bytes used in the following frame field with actual data.
+The rest of the fields in fig.
+\ref{fig:CANframe} is not important concerning the nodes ability to communicate across the CAN network.\\
+The CAN bus itself is two twisted wires, CAN High and CAN Low.
+These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively.
+A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.
+\subsection{RF module} % Steffan 
+The nRF24L01+ LoRa RF module is used provide two-way communication between the solar car and support vehicle.
+module was chosen as it satisfied the project requirements on paper with a range exceeding 1000 m with and transmission rate of up to 2 Mbit/s.
+In addition, comprehensive driver libraries are provided, allowing us
+
 \subsection{Message protocol}
@@
 
-\section{Verification} % Victor
-Throughout the project we have used the cross-platform IDE PlatformIO, which is a development platform for embedded systems. This has greatly increased the efficiency in our work due to the
+\section{Verification and validation} % Victor
+As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design.
+We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.\\
+We performed software validation as we went along with the software coding, to make sure that the actual behaviour of the code was as the expected behaviour.
+This was done through the testing environment in PlatformIO.
+We used this to validate our code design with unit tests, testing both edge cases and random cases.
+We chose only to validate the code and functions we made ourselves, and not those that was imported from external libraries.\\
+Doing this provided us a test bench for our project, which greatly improved the credibility of our work, proving that our design works in general cases and not just the case we represent in our report.
 
@@
 
-The authors would like to thank Christian Kampp Kruse for his assistance as coordinator for this project. The authors would also like to thank Claus from DTU Mekanik % Udbedr
+The authors would like to thank Christian Kampp Kruse for his assistance as project coordinator. The authors would also like to thank Claus from the department of Mechanical Engineering at DTU.
 
@@
 https://www.learn-cpp.org/
+\bibitem{Aunit}
+Documentation,
+https://www.arduino.cc/reference/en/libraries/aunit/
 \end{thebibliography}

### 2021-06-13 documentation/FinalPaper/paper.tex

@@
 At first a detailed specification was drafted which then served as a reevaluation tool and an overall guidance each time a sprint was completed and when the goal for the next sprint had to be selected.
-Taking a starting point in this, we made a prototype setup of the hardware on breadboards as our first sprint.
-The following sprints would then mostly be conducted in software, such as making a first transmission and receiving of data, and finally a last hardware sprint where we soldered our hardware to a perf board.\\
-During the coding of the software functions, we also made verification and validation of our code design.
-In verification we would discuss code naming and functionality.
-To validate our code we made unit tests that would test the actual behaviour of our code against the intended behaviour.\\
-During the coding phase, we attempted to make our debugging more efficient with the use of a debugger.
-However, this would prove quite troublesome, as the Teensy board does not have an on-board debugger and in general is not very suited for debuggers.
-If we would insist on using a debugger such as st-link, we would have to hold or solder wires to the debug ports on the Teensy board's backside, which would complicate our prototype setup.
+Using the specification, we made a prototype setup of the hardware on breadboards as our first sprint.
+The following sprints would then mostly have software development objectives, such as achieving a first transmission and receiving of data.
+Finally, a last hardware sprint was conducted where we soldered our hardware to a perf board.\\
+During the development of the software functions, we also verified and validated our code design.
+In verification the code and its functionality was discussed.
+To validate our code we made unit tests that test the actual behaviour of our code against the intended behaviour.\\
+During the software development phase, we attempted to make our debugging more efficient with the use of a debugger.
+However, this proved quite troublesome, as the Teensy 3.6 board does not have an on-board debugger and in general is not very suited for debuggers.
+If we had insisted on using a debugger such as st-link, we would have had to hold or solder wires to the debug ports on the Teensy board's backside, which would complicate our prototype setup.
 Therefore, we decided not to spent more time on finding a debugger solution.
-This was a clear setback, as we would spent a lot of time debugging our code.
+This was a clear setback, as debugging could only be conducted by sending text over the serial connection.
 \section{Design}
@@
 \subsubsection{ChibiOS}
-An open-source RTOS for embedded systems including Teensy. Supports threads, semaphores, data-sharing and timers to ease the scheduling of different tasks by the program kernel.
+An open-source RTOS for embedded systems including Teensy. It supports threads, semaphores, data-sharing and timers and more to ease the scheduling of different tasks by the program kernel.
 
@@
 \subsection{RF module} % Steffan 
-The nRF24L01+ LoRa RF module is used provide two-way communication between the solar car and support vehicle.
-module was chosen as it satisfied the project requirements on paper with a range exceeding 1000 m with and transmission rate of up to 2 Mbit/s.
-In addition, comprehensive driver libraries are provided, allowing us
+The nRF24L01+ LoRa RF module is used to provide two-way communication between the solar car and support vehicle.
+It satisfies all the communications requirements a range exceeding 1000 m with and transmission rate of up to 2 Mbit/s.
+In addition, comprehensive driver libraries are provided, allowing us to raise the abstraction of our transmission protocol.
+\\
 
 \subsection{Message protocol}
+The nRF24L01 module uses the Enhanced Shockburst packet structure to transmit messages over the 2.4 GHz ISM band. The package
 
@@
 
-The authors would like to thank Christian Kampp Kruse for his assistance as project coordinator. The authors would also like to thank Claus from the department of Mechanical Engineering at DTU.
+The authors would like to thank Martin Schoeberl and Christian Kampp Kruse for their mentoring and assistance throughout the project.
+The authors would also like to thank Claus from the department of Mechanical Engineering at DTU.
 

### 2021-06-13 documentation/dear_diary/log_diary.tex

@@
 
+\section{07-06}
+As we've encountered increasingly difficulties with debugging the code for operation, we've decided to look more seriously into finding a suitable debug tool for the teensy board.
+However, as the Teensy board comes with no debug tool, we need to setup our own.
+At first, we've tried with a ft232r using ftdi debug tool, but because of difficulties with the compability of PlatformIO, we have abandoned this debug tool.
+PlatformIO itself suggests J-Link as a compatible debug tool, we will look into this and see if such a debug tool is available at DTU.
+
+\section{08-06}
+Christian has provided us with a St-link debugger and although it is not J-link, it should be able to work with the Teensy board, as it is an ARM processor.
+It seems that we must admit that debugging the Teensy is not likely to be possible for us. https://mcuoneclipse.com/2017/04/29/modifying-the-teensy-3-5-and-3-6-for-arm-swd-debugging/
+After finding out how to connect the st-link debugger to the board, we may find that the debugging will not work in spite of this.
+Also, since the pins must be connected to the backside of the board, which proves troublesome to our setup with the breadboard.
+(Why could the debug pins not be on the topside?!)
+
+\section{10-06}
+After several days of work, we performed an actual range test.
+The solar car unit was indoors, due to power supply requirements, (which will be nullified when we make a setup on perf board), and the support car unit moved outdoors.
+We observed that inside the building, the transmission was not affected by the range (about 30m), while as soon as the support module was outside the building, the transmission range dropped rapidly.
+With a glass window in between the two modules, the max range was about 5m.
+However, we were testing with the power amplifier on the lowest settings, so when we perform more tests on perf board, we have the option to increase this amplifier.
+
 \end{document}

### 2021-06-13 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp
Added poster folder

@@ -1,6 +1,38 @@
 #include <Arduino.h>
 #include <SPI.h>
 #include "support-vehicle/threads.h"
+#include "RSA.h"
+
+
+void rsa_print1()
+{
+  Security sec; 
+  BaseTelemetryMsg msg; char str[64]; 
+  msg.randomize(); 
+  msg.toString(str, sizeof(str)); 
+  Serial.println(str); 
+  sec.encrypt((uint16_t*)&msg, 32); 
+  msg.toString(str, sizeof(str)); 
+  Serial.println(str);
+  sec.decrypt((uint16_t*)&msg, 32); 
+  msg.toString(str, sizeof(str)); 
+  Serial.println(str);
+}
+
+void rsa_print2()
+{
+  RSA rsa; 
+  BaseTelemetryMsg msg; char str[64]; 
+  msg.randomize(); 
+  msg.toString(str, sizeof(str)); 
+  Serial.println(str); 
+  rsa.encrypt((uint8_t*)&msg, 32); 
+  msg.toString(str, sizeof(str)); 
+  Serial.println(str);
+  rsa.decrypt((uint8_t*)&msg, 32); 
+  msg.toString(str, sizeof(str)); 
+  Serial.println(str);
+}
 
 // ChibiOS setup function
 // Initializes all 4 threads with a working area, priority level, thread function and initial argument. 
@@ -18,25 +50,12 @@ void setup()
   while(!Serial){}
   RFinit(); 
   chBegin(chSetup);     // Initialize and start all 4 threads
-  //while (true){}
+  //rsa_print2();
+  while (true){}
 }
 
 void loop()
 {
-  /*
-  for (int i = 0; i < 256; i++){
-    uint8_t message = i; 
-    Serial.print("Original message: "); 
-    Serial.print(message); 
-    sec.encrypt(&message, 1); 
-    Serial.print("; Encrypted message: ");
-    Serial.print(message); 
-    sec.decrypt(&message, 1); 
-    Serial.print("; Decrypted message: ");
-    Serial.print(message); 
-    Serial.println();
-    delay(2000); 
-  }
   /*Serial.println(radio.available()); 
   delay(1000);
   /* Not used */

### 2021-06-13 Poster/main.tex

%======================================================================
%===  dtuposter - a class to make posters tha comply with the DTU CI
%
% Written and maintained in 2011-2014 
% by Jorrit Wronski (jowr@mek.dtu.dk)
%
%
%==========================================
%===  details and poster setup
\documentclass[
%    ,title     = {{Title test}}
%    ,author    = {{ThisIs MyName}}
%    ,subject   = {{This is the subject of my work}}
%    ,bgcolor   = dtulightgreen
%    ,highlight = dtuyellow
%    ,toplogo   = {{tex_dtu_aqua_b_uk}}
%    ,botlogo   = {{tex_dtu_bibliotek_b_uk}}
%    ,papersize = {{a0paper}}
%    ,colcount  = {{1column}}
%    ,longtitle
%    ,largecaption
%    ,draft
%    ,nocrop
%    ,english        % language
%    ,fleqn          % equations on the left
]{dtuposter}
%
%
%======================================================================
%===  Continue with packages
\usepackage[T1]{fontenc}        % special characters
%
%\usepackage[ansinew]{inputenc}  % Windows
%\usepackage[applemac]{inputenc} % MacOS
\usepackage[utf8x]{inputenc}    % Unicode, Linux
%
% 
%======================================================================
%=== Font definitions, DTU recommends Arial for posters
\usepackage{cmbright}
\usepackage{arevmath}
%\usepackage[scaled]{uarial} %Arial clone, set as default sf font - use "ua1" for direct access
%\usepackage{uarial} %Arial clone, set as default sf font - use "ua1" for direct access
%\usepackage[typeface=default,
%            sanstypeface=urwarial,
%            mathtypeface=arevmath
%           ]{typeface}
\renewcommand{\familydefault}{\sfdefault}
\usepackage{enumitem}
\setlist{nosep,leftmargin=*}
%
% 
%======================================================================
%=== Other useful packages
\usepackage{booktabs}
\usepackage{siunitx}
%
% 
%======================================================================
%=== The actual content starts here
\begin{document}
%
%
%======================================================================
%===  Make header for poster (title and authors)
\begin{dtuposterhead} %
\dtuposterauthor{First Author\textsuperscript{1,*}, Second Author\textsuperscript{2}}
\dtuposteraffil{\textsuperscript{1} Technical University of Denmark, Department of Mechanical Engineering, Kgs. Lyngby, Denmark}
\dtuposteraffil*{\textsuperscript{*} Corresponding author, Tel.: +45 4525 0000, Fax: +45 4588 0000, Email: name@mek.dtu.dk}
\dtuposteraffil{\textsuperscript{2} RWTH Aachen University, E.ON Energy Research Center, Aachen, Germany}
\end{dtuposterhead}
%
%
%======================================================================
%===  ... and the rest of the content
\begin{dtupostercontent}
\section{Introduction}
In general, it should be straight forward to use this poster class. Please
have a close look at this example file. I tried to exploit all special 
features and left a lot of comments in the code. After reading this 
guide, you are able to produce decent posters in the DTU layout with
just a few lines of \LaTeX\ code.

\section{Basic Options}
When loading this poster class, you can pass some options to adapt the 
file to your needs. There are two groups of options, key-value options 
and simple options. Key-value options are used by writing the option 
name, the key, followed by the value as in the examples below. 
Unfortunately, you have to
wrap it with curly brackets. Due to some problems with the option 
processing, it is highly recommended to use pairs of two curly brackets
for this purpose. The available key-value options and their 
default values are:
\begin{itemize}
\item \texttt{title     = \{\{This is the Poster Title\}\}}
\item \texttt{author    = \{\{This Ismyname\}\}}
\item \texttt{subject   = \{\{Subject of the Poster\}\}}
\item \texttt{bgcolor   = \{\{dtucoolgrey\}\}}
\item \texttt{highlight = \{\{dtured\}}
\item \texttt{toplogo   = \{\{tex\_dtu\_mekanik\_b\_uk\}}
\item \texttt{botlogo   = \{\{tex\_dtu\_elektro\_b\_uk\}\}}
\item \texttt{papersize = \{\{a0paper\}\}}
\item \texttt{colcount  = \{\{3columns\}\}}
\end{itemize}

\texttt{title}, \texttt{author} and \texttt{subject} just take text or
\LaTeX\ commands as inputs. \texttt{bgcolor} and \texttt{highlight} take 
names of colours, make sure to use a colour that has already been defined.
As the name suggests, \texttt{toplogo} and \texttt{botlogo} are interpreted
as names of image files. \texttt{toplogo} is the image in the top left 
corner and \texttt{botlogo} is the one that appears four times at the bottom.
\texttt{papersize} can have the values \texttt{a0paper}, \texttt{a1paper}, 
\texttt{a2paper}, \texttt{a3paper} or \texttt{a4paper}. Please note that the 
font size will be changed accordingly. You can use this template for posters
with up to three columns, \texttt{colcount} can be \texttt{3columns}, 
\texttt{2columns} or \texttt{1column}. If you use less than three columns, 
the test font size will be slightly increased.

The other options are activated by passing them to the class without any value.
The option \texttt{longtitle} decreases the font font the title and is 
particularly useful for posters with a long title, hence the name. 
\texttt{largecaption} does the opposite with the captions for tables and figures 
since I think that the default size of \SI{15}{pt} is quite small for a poster. 
Passing \texttt{draft} to the class disables the inclusion of graphics and draws 
a grid with \SI{2}{\centi\metre} spacing on top of the poster. The last option
is \texttt{nocrop}. This disables the cropmarks and produces output that 
exactly matches with dimensions specified in the \texttt{papesize} option.

\begin{table}
\caption{An example caption. Please note that there still is a bug that keeps the 
caption from being centered if it is only a single line. It should work with multiple
lines though}\label{tab:example}

\begin{tabular}{llr}
\toprule
\multicolumn{2}{c}{Item} \\
\cmidrule(r){1-2}
Animal    & Description & Price (\$) \\
\midrule
Gnat      & per gram    & 13.65      \\
          & each        & 0.01       \\
Gnu       & stuffed     & 92.50      \\
Emu       & stuffed     & 33.33      \\
Armadillo & frozen      & 8.99       \\
\bottomrule
\end{tabular}
\end{table}

\section{Special Environments}
Please use the custom environments \texttt{dtuposterhead} and 
\texttt{dtupostercontent}. You can format the list of authors and 
their affiliations with the new commands \texttt{\textbackslash dtuposterauthor}
and \texttt{\textbackslash dtuposteraffil}. Both commands have a starred version, 
\texttt{\textbackslash dtuposterauthor*} and \texttt{\textbackslash dtuposteraffil*}, 
which reduces the spacing. This can be used to show that two lines belong together, as
done in this example file.

Figures and tables have gotten their own custom tags \texttt{table} and 
\texttt{figure}, which define the caption style for you. With the \texttt{booktabs}
package, you can show tabular data like in the example \autoref{tab:example}.

For figures like \autoref{fig:example} and \autoref{fig:example2}, everything is 
similar, just that the caption should be below the figure.

\begin{figure}
\includegraphics[width=.5\linewidth,angle=90,origin=c]{tex_dtu_logo}
\caption{An example caption. This figure supports transparency and looks fine on 
the coloured background}\label{fig:example}
\end{figure}

If you have a figure with white background or one that does not support transparency,
you should draw a box around it to make it appear less alien on the coloured background.

\begin{figure}
  \begin{fadebox}\begin{center}
    \includegraphics[width=.5\linewidth,angle=-90,origin=c]{tex_dtu_logo}
  \end{center}\end{fadebox}
  \caption{An example caption. Figures can be wrapped in the \texttt{fadebox} 
environment to provide them with a white background. Note that you have to align the 
figure again.}\label{fig:example2}
\end{figure}

\end{dtupostercontent}

\end{document}

### 2021-06-13 code/SolarVehicle/lib/rf/RFfunctions.h
comments

@@ -6,45 +6,48 @@
 #include <nRF24L01.h>
 #include "TelemetryMessages.h"
 
-#define IRQ_PIN 2 
-#define CE_PIN 9
-#define CSN_PIN 10
+#define IRQ_PIN 2   // SPI interrupt pin
+#define CE_PIN 9    // SPI chip-enable (CE) pin
+#define CSN_PIN 10  // SPI chip-select (CSN) pin
 
 #define COUNT 5     // Number of transmission retries
 #define DELAY 15    // Delay between retries (= DELAY * 250 us + 250 us)
 
 #ifdef TEENSY40_BOARD
-bool radioNumber = 0;
+bool radioNumber = 0;       
 #elif defined(ARDUINO_BOARD)
 bool radioNumber = 1;
 #else
 bool radioNumber = 1;
 #endif
 
-RF24 radio(CE_PIN, CSN_PIN);                            // CE and CSN pins
-static const byte address[][6] = {"00001", "00002"};    // TX/RX byte addresses
+RF24 radio(CE_PIN, CSN_PIN);                                    // CE and CSN pins
+static const byte address[][6] = {"00001", "00002"};            // TX/RX byte addresses
 
-// RFinit() function
-// Takes no arguments.
+/**
+ * @brief Initializes and configures RF24 class object
+ * @param None 
+*/
 void RFinit()
 {
     if(!radio.begin()){Serial.println("Radio not working!");}
     radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate
     radio.enableDynamicPayloads();
-    //radio.setAutoAck(true);
+    radio.setAutoAck(true);
     //radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
-    //pinMode(IRQ_PIN, INPUT);
-    //attachInterrupt(digitalPinToInterrupt(IRQ_PIN), RFinterrupt, FALLING); 
-    //radio.maskIRQ(1, 1, 0);
     radio.openWritingPipe(address[!radioNumber]);
     radio.openReadingPipe(1, address[radioNumber]);
     radio.startListening();                                     // Starts RX mode
 }
 
-// First transmit() function
-// Takes char array and its size to send. size cannot be greater than 32 bytes (null-terminated)
+/**
+ * @brief Tries to transmit message to RX node. Returns true when message was succesfully
+ * transmitted, false otherwise. 
[…]

### 2021-06-13 code/SolarVehicle/src/support-vehicle/threads.h

@@ -15,8 +15,13 @@ struct threadBundle
     ThreadState *state;
 };
 
+// 2048 byte working stack for radioWorkerThread
 THD_WORKING_AREA(waRadioWorkerThread, 2048);
 
+/**
+ * @brief Thread function for radioWorkerThread
+ * @param arg, typecast to threadBundle pointer
+*/
 THD_FUNCTION(radioWorkerThread, arg)
 {
   threadBundle *bundle = (threadBundle*)arg; 
@@ -57,29 +62,32 @@ THD_FUNCTION(radioWorkerThread, arg)
   }
 }
 
+// 2048 byte working stack for serialWorkerThread
 THD_WORKING_AREA(waSerialWorkerThread, 2048);
 
+/**
+ * @brief Thread function for serialWorkerThread  
+ * @param arg, typecast to Security class pointer
+*/
 THD_FUNCTION(serialWorkerThread, arg)
 {
   Security *sec = (Security*)arg; 
   BaseTelemetryMsg message; 
 
-  while (true){
-
-    if (Serial.available()){
-      Serial.readBytes((char*)&message, 32);
-      //sec->encrypt((uint16_t*)&message, 32);   // Encrypt message before transmission
-      if (RFtransmit(&message, 32)){
+  while (true){                                 
+    if (Serial.available()){                      // Check if bytes are available on the serial interface
+      Serial.readBytes((char*)&message, 32);      // Read 32 bytes and cast them to a BaseTelemetryMessage 
+      //sec->encrypt((uint16_t*)&message, 32);    // Encrypt message before transmission
+      if (RFtransmit(&message, 32)){              // Check if message was transmitted succesfully 
         char str[128]; 
         message.toString(str, sizeof(str));
         Serial.print("Transmitted: "); 
         Serial.println(str); 
-        //Serial.println("Acknowledge received.");
-      } else {
-        Serial.println("Transmission failed or timed out.");
+      } else {      
+        Serial.println("Transmission failed or timed out.");  // Message was not transmitted
       }
     }  
-    chThdYield();  
+    chThdYield();                                 // Yield for same-priority thread                       
   }
 } 
 

### 2021-06-13 documentation/FinalPaper/paper.tex

@@
 
-\IEEEPARstart{I}{n} 2023, DTU Roadrunners Solar Team, abbreviated ROAST, is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in Northern Territory to Adelaide in South Australia.
-The solar car has limited opportunities for recharging during the race, so the main energy source during the race will be the sun.
+\IEEEPARstart{I}{n} 2023, the DTU Roadrunners Solar Team (abbreviated ROAST) is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in the Northern Territory to Adelaide in South Australia.
+The solar car has limited opportunities for recharging during the race, so the main energy source during the race will come from the sun.
 To this end, solar panels will cover the car, however, the solar car is only allowed to have a maximum of 4 square meters of solar panels.
@@
 This was a clear setback, as debugging could only be conducted by sending text over the serial connection.
+
 \section{Design}
@@
 The nRF24L01+ LoRa RF module is used to provide two-way communication between the solar car and support vehicle.
-It satisfies all the communications requirements a range exceeding 1000 m with and transmission rate of up to 2 Mbit/s.
-In addition, comprehensive driver libraries are provided, allowing us to raise the abstraction of our transmission protocol.
+The module includes a detachable antenna, power amplifier and low-nose amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
+In addition to this, comprehensive driver libraries are provided by the manufacturer, making it very easy to integrate with our project.
 \\
@@
 \subsection{Message protocol}
-The nRF24L01 module uses the Enhanced Shockburst packet structure to transmit messages over the 2.4 GHz ISM band. The package
+The nRF24L01+ module uses the Enhanced Shockburst $\copyright$ packet structure to transmit messages over the 2.4 GHz ISM band.
+Messages contain a data payload of up to 32 bytes, allowing us to transmit one CAN telemetry message per cycle.
+Further, the package structure includes a preamble, address, packet and CRC byte segments as shown in Figure X.
+These are all generated automatically when a call to the radio write function is made, and therefore fall outside the scope of this project.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/MessageTypes.pdf}
+    \caption{The }
+    \label{fig:my_label}
+\end{figure}
 
@@
 https://www.arduino.cc/reference/en/libraries/aunit/
+\bibitem{shockburst}
+nRF24 transmission protocol, URL: https://lastminuteengineers.com/nrf24l01-arduino-wireless-communication/
 \end{thebibliography}

### 2021-06-14 code/SolarVehicle/lib/rf/RFfunctions.h
comments on RFfunctions

@@ -1,3 +1,9 @@
+/********************************************************************************
+ * @file    RFfunctions.h                                                       *
+ * @author  Steffan Martin Kunoy                                                *
+ * @brief   Declaration of functions to operate and configure nRF24L01+ module. * 
+*********************************************************************************/
+
 #ifndef __RF_FUNCTIONS__
 #define __RF_FUNCTIONS__
 #include <Arduino.h>
@@ -23,11 +29,12 @@ bool radioNumber = 1;
 
 RF24 radio(CE_PIN, CSN_PIN);                                    // CE and CSN pins
 static const byte address[][6] = {"00001", "00002"};            // TX/RX byte addresses
+bool ack = true; 
 
-/**
- * @brief Initializes and configures RF24 class object
- * @param None 
-*/
+/**********************************************************
+ * @brief   Initializes and configures RF24 class object. *
+ * @param   None                                          *
+***********************************************************/
 void RFinit()
 {
     if(!radio.begin()){Serial.println("Radio not working!");}
@@ -35,58 +42,47 @@ void RFinit()
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate
     radio.enableDynamicPayloads();
     radio.setAutoAck(true);
-    //radio.enableAckPayload();
+    if (ack) radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
     radio.openWritingPipe(address[!radioNumber]);
     radio.openReadingPipe(1, address[radioNumber]);
     radio.startListening();                                     // Starts RX mode
 }
 
-/**
- * @brief Tries to transmit message to RX node. Returns true when message was succesfully
- * transmitted, false otherwise. 
- * @param buf Pointer to the data to be sent. 
- * @param len Number of bytes to be sent. Maximum 32 bytes (null-terminated).  
-*/
+/********************************************************************************
+ * @brief   Tries to transmit message to RX node.                               *
+ * @param   buf Pointer to the data to be sent.                                 *
+ * @param   len Number of bytes to be sent. Maximum 32 bytes (null-terminated). * 
+ * @return  True if message was succesfully transmitted, otherwise false.       *
+*********************************************************************************/
 bool RFtransmit(void *buf, uint8_t len)
 {
-    radio.stopListening();                                      // Starts TX mode
-    bool report = radio.write(buf, len);                       // Send message and wait for acknowledge
+    radio.stopListening();                              // Starts TX mode
[…]

### 2021-06-14 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp

@@ -40,23 +40,20 @@ void chSetup()
 {
   chSysInit();  // Initializes ChibiOS system
   threadBundle bundle = {.sec = &sec, .state = &rfState}; 
-  chThdCreateStatic(waRadioWorkerThread, sizeof(waRadioWorkerThread), NORMALPRIO + 1, radioWorkerThread, &bundle);
-  chThdCreateStatic(waSerialWorkerThread, sizeof(waSerialWorkerThread), NORMALPRIO + 1, serialWorkerThread, &sec);
+  chThdCreateStatic(WaReceiverThread, sizeof(WaReceiverThread), NORMALPRIO + 1, receiverThread, &bundle);
+  chThdCreateStatic(waTransmitterThread, sizeof(waTransmitterThread), NORMALPRIO + 1, transmitterThread, &sec);
 }  
 
 void setup()
 {
-  Serial.begin(921600); // initialize serial port
-  while(!Serial){}
-  RFinit(); 
+  Serial.begin(921600); // Initialize serial port
+  while(!Serial){}      // Wait until serial is available
+  RFinit();             // Initialize RF module
   chBegin(chSetup);     // Initialize and start all 4 threads
-  //rsa_print2();
-  while (true){}
+  while (true){}        
 }
 
 void loop()
 {
-  /*Serial.println(radio.available()); 
-  delay(1000);
   /* Not used */
 }

### 2021-06-14 code/SolarVehicle/src/support-vehicle/threads.h

@@ -6,8 +6,8 @@
 #include "ThreadState.h"
 #include "Encryption.h"
 
-Security sec; 
-ThreadState rfState;
+Security sec;         // Security global variable
+ThreadState rfState;  // ThreadState global variable
 
 struct threadBundle
 {
@@ -15,14 +15,14 @@ struct threadBundle
     ThreadState *state;
 };
 
-// 2048 byte working stack for radioWorkerThread
-THD_WORKING_AREA(waRadioWorkerThread, 2048);
+// 2048 byte working stack for receiverThread
+THD_WORKING_AREA(WaReceiverThread, 2048);
 
 /**
- * @brief Thread function for radioWorkerThread
- * @param arg, typecast to threadBundle pointer
+ * @brief       Thread function for receiverThread
+ * @param arg   Typecast to threadBundle pointer
 */
-THD_FUNCTION(radioWorkerThread, arg)
+THD_FUNCTION(receiverThread, arg)
 {
   threadBundle *bundle = (threadBundle*)arg; 
   Security *sec = bundle->sec; 
@@ -58,18 +58,18 @@ THD_FUNCTION(radioWorkerThread, arg)
         Serial.println("Could not receive message.");
       }
     }
-    chThdYield();
+    chThdYield();                                       // Yield for same-priority thread
   }
 }
 
-// 2048 byte working stack for serialWorkerThread
-THD_WORKING_AREA(waSerialWorkerThread, 2048);
+// 2048 byte working stack for transmitterThread
+THD_WORKING_AREA(waTransmitterThread, 2048);
 
 /**
- * @brief Thread function for serialWorkerThread  
- * @param arg, typecast to Security class pointer
+ * @brief     Thread function for transmitterThread  
+ * @param arg Typecast to Security class pointer
 */
-THD_FUNCTION(serialWorkerThread, arg)
+THD_FUNCTION(transmitterThread, arg)
 {
   Security *sec = (Security*)arg; 
   BaseTelemetryMsg message; 

### 2021-06-14 code/TelemetryUI/src/main/scala/telemetryui/types/CanFrame.scala
change message length to 16 bytes

@@ -11,7 +11,7 @@ case class CanFrame(id: BigInt, rtr: Boolean, len: Int, data: BigInt, stamp: Tim
     println(data.toByteArray.reverse.padTo(8,0.toByte).mkString(", "))
     (id.toByteArray.reverse.padTo(2,0.toByte) ++ Array((if(rtr)1 else 0).toByte) ++ Array(len.toByte) ++
       (if((data & 0x8000000000000000L) == 0) data.toByteArray.reverse.padTo(8,0.toByte) else data.toByteArray.drop(1).reverse.padTo(8,0.toByte)) ++
-      stamp.toByteArray).padTo(31,0.toByte)
+      stamp.toByteArray).padTo(15,0.toByte)
   }
 }
 

### 2021-06-14 code/SolarVehicle/src/support-vehicle/supportVehicleMain.cpp
update .ini file

@@ -1,48 +1,25 @@
+/****************************************************************************************************
+ * @file    supportVehicleMain.cpp                                                                  *
+ * @author  Steffan Martin Kunoy                                                                    *
+ * @brief   Includes setup() and loop() functions to be run on the support vehicle microcontroller. *
+ ****************************************************************************************************/
+
 #include <Arduino.h>
 #include <SPI.h>
 #include "support-vehicle/threads.h"
 #include "RSA.h"
 
-
-void rsa_print1()
-{
-  Security sec; 
-  BaseTelemetryMsg msg; char str[64]; 
-  msg.randomize(); 
-  msg.toString(str, sizeof(str)); 
-  Serial.println(str); 
-  sec.encrypt((uint16_t*)&msg, 32); 
-  msg.toString(str, sizeof(str)); 
-  Serial.println(str);
-  sec.decrypt((uint16_t*)&msg, 32); 
-  msg.toString(str, sizeof(str)); 
-  Serial.println(str);
-}
-
-void rsa_print2()
-{
-  RSA rsa; 
-  BaseTelemetryMsg msg; char str[64]; 
-  msg.randomize(); 
-  msg.toString(str, sizeof(str)); 
-  Serial.println(str); 
-  rsa.encrypt((uint8_t*)&msg, 32); 
-  msg.toString(str, sizeof(str)); 
-  Serial.println(str);
-  rsa.decrypt((uint8_t*)&msg, 32); 
-  msg.toString(str, sizeof(str)); 
-  Serial.println(str);
-}
-
-// ChibiOS setup function
-// Initializes all 4 threads with a working area, priority level, thread function and initial argument. 
+/*********************************************************
+ * @brief Initializes all threads with a working area,   *
+ * priority level, thread function and initial argument. *
+ *********************************************************/  
 void chSetup()
 {
   chSysInit();  // Initializes ChibiOS system
   threadBundle bundle = {.sec = &sec, .state = &rfState}; 
   chThdCreateStatic(WaReceiverThread, sizeof(WaReceiverThread), NORMALPRIO + 1, receiverThread, &bundle);
   chThdCreateStatic(waTransmitterThread, sizeof(waTransmitterThread), NORMALPRIO + 1, transmitterThread, &sec);
-}  
+}
 
 void setup()
 {

### 2021-06-14 code/SolarVehicle/src/telemetry/solar-car/CanReceiverThd.h
move src code into separate package

@@ -12,7 +12,7 @@
 #include "MutexLocker.h"
 #include "TelemetryMessages.h"
 
-#include "solar-car/Mutexes.h"
+#include "telemetry/solar-car/Mutexes.h"
 
 /**
  * A bundle used for passsing all relevant resources to the thread

### 2021-06-14 code/SolarVehicle/src/telemetry/solar-car/RfTxThd.h

@@ -15,7 +15,7 @@
 #include "Measure.h"
 #include "Encryption.h"
 
-#include "solar-car/Mutexes.h"
+#include "telemetry/solar-car/Mutexes.h"
 
 
 /**

### 2021-06-14 code/SolarVehicle/src/telemetry/solar-car/SystemThd.h

@@ -15,7 +15,7 @@
 #include "Measure.h"
 #include "Encryption.h"
 
-#include "solar-car/Mutexes.h"
+#include "telemetry/solar-car/Mutexes.h"
 
 /**
  * A bundle used for passsing all relevant resources to the radio thread

### 2021-06-14 code/SolarVehicle/src/telemetry/solar-car/blackBoxThd.h

@@ -12,7 +12,7 @@
 #include "MultiReaderFifo.h"
 #include "MutexLocker.h"
 
-#include "solar-car/Mutexes.h"
+#include "telemetry/solar-car/Mutexes.h"
 
 /**
  * A bundle used for passsing all relevant resources to the radio thread

### 2021-06-14 code/SolarVehicle/src/telemetry/solar-car/solarCarMain.cpp

@@ -21,11 +21,11 @@
 #include "MultiReaderFifo.h"
 #include "ThreadState.h"
 
-#include "solar-car/blackBoxThd.h"
-#include "solar-car/CanReceiverThd.h"
-#include "solar-car/Mutexes.h"
-#include "solar-car/RfTxThd.h"
-#include "solar-car/SystemThd.h"
+#include "telemetry/solar-car/blackBoxThd.h"
+#include "telemetry/solar-car/CanReceiverThd.h"
+#include "telemetry/solar-car/Mutexes.h"
+#include "telemetry/solar-car/RfTxThd.h"
+#include "telemetry/solar-car/SystemThd.h"
 
 
 #ifdef CH_CFG_TIME_QUANTUM

### 2021-06-14 code/SolarVehicle/src/telemetry/support-vehicle/supportVehicleMain.cpp

@@ -1,6 +1,6 @@
 #include <Arduino.h>
 #include <SPI.h>
-#include "support-vehicle/threads.h"
+#include "telemetry/support-vehicle/threads.h"
 #include "RSA.h"
 
 

### 2021-06-14 code/SolarVehicle/lib/message/TelemetryMessages.h
Update TelemetryMessages.h

@@ -53,7 +53,7 @@ class BaseTelemetryMsg {
 
   public: uint32_t toString(char *buf, uint32_t len){
     uint32_t curPos = snprintf(buf,len,"%02x:",cmd);
-    for(uint8_t i = 0; i < 31; i++){
+    for(uint8_t i = 0; i < BYTES_PER_MESSAGE-1; i++){
       curPos += snprintf(&(buf[curPos]),len-curPos," %02x",data[i]);
     }
     return curPos;

### 2021-06-14 code/SolarVehicle/src/telemetry/solar-car/SystemThd.h
Update SystemThd.h

@@ -52,8 +52,9 @@ THD_FUNCTION(systemThd, arg){
    
     radio.startListening();
     if(radio.available()){
-      RFreceive(&msg,32);
-      //sec->decrypt((uint8_t*)&msg,32);
+      uint16_t encrypted[16];
+      RFreceive(encrypted,BaseTelemetryMsg::length()<<1);
+      sec->decrypt(encrypted,(uint8_t*)&msg,32);
       uint32_t count = 0;
       for(uint32_t i = 0; i < 32; i++) {
         count += ((uint8_t*)&msg)[i];

### 2021-06-14 documentation/FinalPaper/paper.tex
Updates from Overleaf

@@
 The nRF24L01+ LoRa RF module is used to provide two-way communication between the solar car and support vehicle.
-The module includes a detachable antenna, power amplifier and low-nose amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
+The module includes a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
 In addition to this, comprehensive driver libraries are provided by the manufacturer, making it very easy to integrate with our project.
-\\
 
 \subsection{Message protocol}
+The radio connection is the pivot point of the telemetry system and a flexible way to pass commands and associated payloads both ways is needed. 
 The nRF24L01+ module uses the Enhanced Shockburst $\copyright$ packet structure to transmit messages over the 2.4 GHz ISM band.
-Messages contain a data payload of up to 32 bytes, allowing us to transmit one CAN telemetry message per cycle.
-Further, the package structure includes a preamble, address, packet and CRC byte segments as shown in Figure X.
+Messages contain a data payload of up to 32 bytes, allowing us comfortably to transmit one CAN telemetry message per packet.
+Further, the package structure includes a preamble, address, packet and CRC bytes segments as shown in Figure \ref{fig:shockburst}.
 These are all generated automatically when a call to the radio write function is made, and therefore fall outside the scope of this project.
-
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.png}
+    \caption{The Enhanced Shockburst $\copyright$ packet structure.}
+    \label{fig:shockburst}
+\end{figure}
 \begin{figure}
@@
 \subsubsection{Black Box} % tjark
+
+The black box functionality in the solar car is implemented using a SD card. The Teensy 3.6 used as a microcontroller in the telemetry module of the solar car already includes a 4-bit
+
 \subsubsection{CAN transceiver}

### 2021-06-16 code/SolarVehicle/lib/rf/RFfunctions.h
Update Solar Car Threads

@@ -12,9 +12,15 @@
 #include <nRF24L01.h>
 #include "TelemetryMessages.h"
 
+#ifdef TEENSY36_BOARD
+#define IRQ_PIN 2   // SPI interrupt pin
+#define CE_PIN 10    // SPI chip-enable (CE) pin
+#define CSN_PIN 9  // SPI chip-select (CSN) pin
+#else
 #define IRQ_PIN 2   // SPI interrupt pin
 #define CE_PIN 9    // SPI chip-enable (CE) pin
 #define CSN_PIN 10  // SPI chip-select (CSN) pin
+#endif
 
 #define COUNT 5     // Number of transmission retries
 #define DELAY 15    // Delay between retries (= DELAY * 250 us + 250 us)
@@ -29,7 +35,7 @@ bool radioNumber = 1;
 
 RF24 radio(CE_PIN, CSN_PIN);                                    // CE and CSN pins
 static const byte address[][6] = {"00001", "00002"};            // TX/RX byte addresses
-bool ack = true; 
+bool ack = false; 
 
 /**********************************************************
  * @brief   Initializes and configures RF24 class object. *
@@ -38,7 +44,7 @@ bool ack = true;
 void RFinit()
 {
     if(!radio.begin()){Serial.println("Radio not working!");}
-    radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level
+    radio.setPALevel(RF24_PA_MAX);                              // Set Power Amplifier level
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate
     radio.enableDynamicPayloads();
     radio.setAutoAck(true);

### 2021-06-16 code/SolarVehicle/src/telemetry/solar-car/solarCarMain.cpp

@@ -45,7 +45,7 @@ void setup(){
   
   // initialize serial port
   Serial.begin(921600);
-  while(!Serial){} //needs to be removed when headless!!!!!!!!!!!!!!!!!!!!
+  //while(!Serial){} //needs to be removed when headless!!!!!!!!!!!!!!!!!!!!
 
   // setup built in LED
   pinMode(LED_BUILTIN,OUTPUT);
@@ -57,6 +57,7 @@ void setup(){
   if(ACAN::can0.begin(settings) != 0){Serial.println("CAN setup failed!");}
 
   // setup radio module
+  Serial.println(radio.begin());
   RFinit();
 
   bb.init();

### 2021-06-16 code/TelemetryUI/src/main/scala/telemetryui/serial/SerialWorker.scala
Update Telemetry UI

@@ -42,10 +42,10 @@ class SerialWorker(port: SerialPort,
           assembler.add(b)
           if(assembler.hasMessage) {
             val msg = assembler.getMessage
-            cmdListener.foreach(_(msg.cmd))
-            if(msg.cmd == RECEIVED_CAN) canListener.foreach(_(msg.can.get))
+            cmdListener.foreach(f => synchronized{f(msg.cmd)})
+            if(msg.cmd == RECEIVED_CAN) canListener.foreach(f => synchronized{f(msg.can.get)})
           }
-          if(assembler.hadError) errListener.foreach(_())
+          if(assembler.hadError) errListener.foreach(f => synchronized{f()})
         }
       }
 

### 2021-06-16 code/TelemetryUI/src/main/scala/telemetryui/types/CanFrame.scala

@@ -7,8 +7,6 @@ import math.pow
 case class CanFrame(id: BigInt, rtr: Boolean, len: Int, data: BigInt, stamp: TimeStamp) {
   override def toString: String = s"($id: [$len]$data at $stamp)"
   def toByteArray : Array[Byte] = {
-    //TODO: padTo seems to create arrays that vary in size
-    println(data.toByteArray.reverse.padTo(8,0.toByte).mkString(", "))
     (id.toByteArray.reverse.padTo(2,0.toByte) ++ Array((if(rtr)1 else 0).toByte) ++ Array(len.toByte) ++
       (if((data & 0x8000000000000000L) == 0) data.toByteArray.reverse.padTo(8,0.toByte) else data.toByteArray.drop(1).reverse.padTo(8,0.toByte)) ++
       stamp.toByteArray).padTo(15,0.toByte)

### 2021-06-16 code/SolarVehicle/test/test_common_solar/test_solar.cpp
test_solar_update - added description+instructions

@@ -1,20 +1,14 @@
 /*
- Copyright (c) 2014-present PlatformIO <contact@platformio.org>
+ Test file for testing various functions made for the telemetry module for the ROAST solar car
+ @author Victor Alexander Hansen
+ @author Steffan Martin Kunoy
+ @author Tjark Petersen
 
- Licensed under the Apache License, Version 2.0 (the "License");
- you may not use this file except in compliance with the License.
- You may obtain a copy of the License at
-
-    http://www.apache.org/licenses/LICENSE-2.0
-
- Unless required by applicable law or agreed to in writing, software
- distributed under the License is distributed on an "AS IS" BASIS,
- WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
- See the License for the specific language governing permissions and
- limitations under the License.
+ To run the test, open a pio terminal and type "pio test" in PlatformIO
+ Be sure to have the correct board connected for the corresponding environment in the .ini file.
+ If the test keeps uploading but nothing happens, try to comment out the other environments in
+ the .ini file
 **/
-
-// Maybe use pio --without-building?
 
 #include <unity.h>
 #include <ACAN.h>
@@ -29,13 +23,14 @@
 #include "RFfunctions.h"
 #include "MultiReaderFifo.h"
 #include "ThreadState.h"
-#include "solar-car/blackBoxThd.h"
-#include "solar-car/CanReceiverThd.h"
-#include "solar-car/Mutexes.h"
-#include "solar-car/RfTxThd.h"
-#include "solar-car/SystemThd.h"
+#include "telemetry/solar-car/blackBoxThd.h"
+#include "telemetry/solar-car/CanReceiverThd.h"
+#include "telemetry/solar-car/Mutexes.h"
+#include "telemetry/solar-car/RfTxThd.h"
+#include "telemetry/solar-car/SystemThd.h"
 
 CANMessage frame;
+CanTelemetryMsg CANmsg;
 BlackBox box;
 Security security;
 MultiReaderFifo<CanTelemetryMsg> canFifo(64,2);
@@ -71,20 +66,69 @@ void test_can_x_begin(void) {
 
[…]

### 2021-06-16 code/SolarVehicle/lib/rf/RFfunctions.h
implement TX FIFO in thread

@@ -29,23 +29,23 @@ bool radioNumber = 1;
 
 RF24 radio(CE_PIN, CSN_PIN);                                    // CE and CSN pins
 static const byte address[][6] = {"00001", "00002"};            // TX/RX byte addresses
-bool ack = true; 
+bool ack = false; 
 
 /**********************************************************
  * @brief   Initializes and configures RF24 class object. *
  * @param   None                                          *
 ***********************************************************/
 void RFinit()
-{
+{ 
     if(!radio.begin()){Serial.println("Radio not working!");}
-    radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level
-    radio.setDataRate(RF24_1MBPS);                              // Set Data Rate
-    radio.enableDynamicPayloads();
-    radio.setAutoAck(true);
+    radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level, choose between MIN, LOW, HIGH, MAX (higher PA level improves range)
+    radio.setDataRate(RF24_1MBPS);                              // Set Data Rate, choose between RF24_250KBS, RF24_1MBS, RF24_2MBS (higher Data Rates may cause data loss)
+    radio.enableDynamicPayloads();                              // Enable variable data payloads
+    radio.setAutoAck(true);                                     // RX node sends an automatic ack packet
     if (ack) radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry
-    radio.openWritingPipe(address[!radioNumber]);
-    radio.openReadingPipe(1, address[radioNumber]);
+    radio.openWritingPipe(address[!radioNumber]);               // Open writing pipe at address 
+    radio.openReadingPipe(1, address[radioNumber]);             // Open reading pipe at address
     radio.startListening();                                     // Starts RX mode
 }
 

### 2021-06-16 code/SolarVehicle/src/telemetry/support-vehicle/RfThd.h

@@ -1,218 +1,0 @@
-#ifndef __RF_THD_H__
-#define __RF_THD_H__        
-
-#include <ChRt.h>
-#include "ThreadState.h"
-#include "RFfunctions.h"
-#include "Fifo.h"
-
-// Two FIFO buffers for radio & serial I/O
-Fifo<BaseTelemetryMsg> RFoutbox(32), RFinbox(32);
- 
-// Thread states for all 4 active threads
-ThreadState radioReceiverState, radioTransmitterState, serialReceiverState, serialTransmitterState; 
-ThreadState radioWorkerState, serialWorkerState; 
-// A bundle used for passsing all relevant resources to each thread
-struct threadBundle
-{
-    Fifo<BaseTelemetryMsg> *fifo[2];
-    ThreadState *state;
-};
-/*
-// 1024 byte working stack for radioReceiverThread
-THD_WORKING_AREA(waRadioReceiverThread, 1024);
-
-// Thread function for radioReceiverThread 
-THD_FUNCTION(radioReceiverThread, arg){
-
-  threadBundle *radioReceiverBundle = (threadBundle*) arg;
-  ThreadState *radioReceiverState = radioReceiverBundle->state;
-  Fifo<BaseTelemetryMsg> *radioReceiverFifo = radioReceiverBundle->fifo;
-  Serial.println("radioReceiver");
-  //chThdSleepMilliseconds(100);
-  while(!radioReceiverState->terminate){
-    //Serial.println("radioReceiverThread");
-
-    if (radioReceiverFifo->full()){
-      radioReceiverFifo->waitForSpace();
-    }
-
-    if (radioReceiverState->pause){
-      radioReceiverState->suspend();
-    }
-
-    if (RFreceive(radioReceiverFifo->tail())){
-      Serial.println("radioReceiverThread");
-      //char str[64];
-      //Serial.println((radioReceiverFifo->tail())->toString(str,sizeof(str)));
-      radioReceiverFifo->moveTail();
-      radioReceiverFifo->signalData();
-    }
-    chThdYield();  
-  }
-}
-
-// 1024 byte working stack for radioTransmitterThread
-THD_WORKING_AREA(waRadioTransmitterThread, 1024);
-
-// Thread function for radioReceiverThread 
-THD_FUNCTION(radioTransmitterThread, arg)
-{
-  threadBundle *radioTransmitterBundle = (threadBundle*) arg;
[…]

### 2021-06-16 code/SolarVehicle/src/telemetry/support-vehicle/supportVehicleMain.cpp

@@ -16,9 +16,9 @@
 void chSetup()
 {
   chSysInit();  // Initializes ChibiOS system
-  threadBundle bundle = {.sec = &sec, .state = &rfState}; 
-  chThdCreateStatic(WaReceiverThread, sizeof(WaReceiverThread), NORMALPRIO + 1, receiverThread, &bundle);
-  chThdCreateStatic(waTransmitterThread, sizeof(waTransmitterThread), NORMALPRIO + 1, transmitterThread, &sec);
+  chThdCreateStatic(WaReceiverThread, sizeof(WaReceiverThread), NORMALPRIO + 1, receiverThread, &sec);
+  threadBundle transmitterBundle = {.security = &sec, .fifo = &TXfifo}; 
+  chThdCreateStatic(waTransmitterThread, sizeof(waTransmitterThread), NORMALPRIO + 1, transmitterThread, &transmitterBundle);
 }
 
 void setup()
@@ -26,6 +26,7 @@ void setup()
   Serial.begin(921600); // Initialize serial port
   while(!Serial){}      // Wait until serial is available
   RFinit();             // Initialize RF module
+  sec.encryption_key(); // Create message encryption key
   chBegin(chSetup);     // Initialize and start all 4 threads
   while (true){}        
 }

### 2021-06-16 code/SolarVehicle/src/telemetry/support-vehicle/threads.h

@@ -1,8 +1,8 @@
-/**
- * @file    threads.h
- * @author  Steffan Martin Kunoy
- * @brief   ChibiOS threads to be run on the Support Vehicle
- */
+/**************************************************************
+ * @file    threads.h                                         *
+ * @author  Steffan Martin Kunoy                              *
+ * @brief   ChibiOS threads to be run on the Support Vehicle. *
+ **************************************************************/
 
 #ifndef __RF_THD_H__
 #define __RF_THD_H__        
@@ -10,61 +10,58 @@
 #include <ChRt.h>
 #include "RFfunctions.h"
 #include "ThreadState.h"
+#include "Fifo.h"
 #include "Encryption.h"
 
-Security sec;         // Security global variable
-ThreadState rfState;  // ThreadState global variable
+Security sec;                       // Security global variable
+Fifo<BaseTelemetryMsg> TXfifo(32);  // Fifo global variable
 
-struct threadBundle
+// A struct used to pass arguments to transmitterThread
+struct threadBundle 
 {
-    Security *sec; 
-    ThreadState *state;
+  Security *security;             // Security class member
+  Fifo<BaseTelemetryMsg> *fifo;   // Fifo class member
 };
 
 // 2048 byte working stack for receiverThread
 THD_WORKING_AREA(WaReceiverThread, 2048);
 
-/***************************************************
- * @brief   Thread function for receiverThread.    *
- * @param   arg, typecast to threadBundle pointer. *
-****************************************************/
+/*****************************************************
+ * @brief   Thread function for receiverThread.      *
+ * @param   arg, typecast to Security class pointer. *
+******************************************************/
 THD_FUNCTION(receiverThread, arg)
 {
-  threadBundle *bundle = (threadBundle*)arg; 
-  Security *sec = bundle->sec; 
-  ThreadState *state = bundle->state; 
-  BaseTelemetryMsg received; 
-  chThdSleepMicroseconds(100); 
+  Security *sec = (Security*)arg;                                   // Cast input argument to Security class pointer
+  BaseTelemetryMsg received;                                        // BaseTelemetryMsg object for received messages
+  uint16_t encrypted[16];                                           // Buffer to contain encrypted message
+  char tempStr[256];                                                // Buffer to print received message to serial
+  chThdSleepMicroseconds(100);                                      // Sleep to allow for creation of secondary thread
 
-  while (!state->terminate){
-
-    if (state->pause){
[…]

### 2021-06-16 documentation/FinalPaper/paper.tex

@@
 \usepackage{graphicx}
+\usepackage{comment}
 \usepackage{textcomp}
@@
 \hyphenation{op-tical net-works semi-conduc-tor}
+
+\newcommand{\todo}[1]{{\color{olive} TODO: #1}}
 
@@
 \IEEEPARstart{I}{n} 2023, the DTU Roadrunners Solar Team (abbreviated ROAST) is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in the Northern Territory to Adelaide in South Australia.
-The solar car has limited opportunities for recharging during the race, so the main energy source during the race will come from the sun.
+The solar car has limited opportunities for recharging during the race, so its the main energy source will be the sun.
 To this end, solar panels will cover the car, however, the solar car is only allowed to have a maximum of 4 square meters of solar panels.
-This emphasizes the need to create an energy-efficient solar car to win the race \cite{wsc}.\\
-Throughout the race the solar car will be monitored by a support vehicle, which can monitor the car's condition and should be able to issue commands to the solar car's internal systems.
+This emphasizes the need to create an energy-efficient solar car to win the race \cite{wsc}.
+
+Throughout the race the solar car will be monitored by a support vehicle, which can monitor the car's condition and should be able to issue commands to the solar cars internal systems.
 This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
-The support car will therefore be able to analyze and react to the data stream being transmitted from the solar car, even when the driver is preoccupied by driving the car and navigating traffic.
-\\
-The main function of the telemetry module is to read data from the solar car's on-board CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support vehicle, while simultaneously logging data in a local storage acting as a "black box".
-The support vehicle is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus.
-As a consequence, the telemetry module plays a vital role in the communication between both vehicles, considering that the distance between them may be up to 1 km depending on traffic conditions.\\
-Therefore, the telemetry module must both serve the purpose of fulfilling the necessary specifications for communication, while also being as energy-efficient as possible.
-This project will mainly focus on the former, that is, ensuring that the module meets the specifications with a solid solution.
+The support vehicle will therefore be able to analyze and react to the data stream being transmitted from the solar car, even when the driver is preoccupied with driving the car.
+
+Additionally, valuable data can be collected during the race which can help to identify possible areas of improvement of the solar car.
+Therefore a "black box" system is needed to capture all system activity in a local storage.
+
+These key functionalities require a system which is able to read and write data from the solar cars on-board CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support vehicle, while simultaneously logging data locally.
+The support vehicle is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus via RF.
+Thus, the capabilities of such a telemetry module play a vital role in the communication between both vehicles, considering that the distance between them may be up to 1 km depending on traffic conditions.
+
+The telemetry system must both serve the purpose of fulfilling the necessary specifications for communication, while also being as energy-efficient as possible.
+This project aims at implementing a functioning prototype for a telemetry system while mainly focusing on the former of the two objectives, that is, ensuring that the system meets the specifications with a solid solution.
+
+This paper begins with giving a more precise problem statement and introducing the methodology used throughout the project. Thereafter, the design... \todo{Skal laves når strukturen er slået fast}
 
@@
 The following sprints would then mostly have software development objectives, such as achieving a first transmission and receiving of data.
-Finally, a last hardware sprint was conducted where we soldered our hardware to a perf board.\\
+Finally, a last hardware sprint was conducted where we soldered our hardware to a perf board.
+
 During the development of the software functions, we also verified and validated our code design.
 In verification the code and its functionality was discussed.
-To validate our code we made unit tests that test the actual behaviour of our code against the intended behaviour.\\
+To validate our code we made unit tests that test the actual behaviour of our code against the intended behaviour.
+
 During the software development phase, we attempted to make our debugging more efficient with the use of a debugger.
@@
 \section{Design}
-Overall specification?
+- Block Diagram and description of overall data flow
 
-Block diagram
+\section{Components}
+\subsection{CAN bus}
+% MCP
+\subsection{RF network} % Steffan
+The nRF24L01+ LoRa RF module is used to provide two-way communication between the solar car and support vehicle.
+The module includes a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
+In addition to this, comprehensive driver libraries are provided by the manufacturer, making it very easy to integrate with our project.
+The nRF24L01+ module uses the Enhanced Shockburst $\copyright$ packet structure to transmit messages over the 2.4 GHz ISM band.
+Messages contain a data payload of up to 32 bytes, allowing us comfortably to transmit one CAN telemetry message per packet \cite{shockburst}.
+Further, the package structure includes a preamble, address, packet and CRC bytes segments as shown in Figure \ref{fig:shockburst}.
+\begin{comment}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.png}
+    \caption{The Enhanced Shockburst $\copyright$ packet structure.}
+    \label{fig:shockburst}
+\end{figure}
+\end{comment}
+
+\subsection{SD card}
+\subsection{Microcontrollers} % Steffan
+
+\section{Software Stack}
+
+Several of the key functionalities of the telemetry are implemented in software.
+This together with the aim of designing a system which is easy to extend in the future makes the choice of software frameworks very important, since they affect reusability and maintainability of the developed code.
+
+\subsection{Build system}
+As a build system, platformIO \cite{platformIO} was chosen for the project.
+It is a version control friendly, cross-platform embedded build tool which includes library management and works by defining \textit{environments} to allow development with different embedded platforms on the same code base.
+Other project groups working on the solar car in parallel integrated their software into the platformIO ecosystem as well, making extensive code reuse in the future possible.
+
+\subsection{Real Time Operating System}
+We chose to build the software for the telemetry system on top of a real time operating system.
+The small kernel running behind the scenes when using a RTOS allows a step up in abstraction from bare metal programming.
+Different tasks can be spawned as threads running concurrently.
+These threads get delegated time slices to run on the processor by the real time kernel based on their priority.
+
+This comes at the cost of introducing overhead when switching between threads, but the great flexibility in code organization following with an RTOS was evaluated to make the overhead a worthwhile trade-off.
+
+As a specific RTOS, we chose ChibiOS \cite{chibios}, since it supported both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.
+
+\subsection{Other Libraries}
+% ACAN.h
+% RF24.h
+%
+
+\section{Internal Communication}
+The main function of the telemetry system is to bridge data streams between different end points.
+As such, a flexible way to pass commands and associated data payloads both ways through the system is needed.
+
+The majority of messages passed between the two telemetry subsystems will contain CAN messages as part of the stream from the solar car to the support vehicle.
+Therefore, the message protocol is designed around 16 byte messages, which are large enough to comfortably fit a CAN message and a time stamp from when the message was received.
+A diagram of the byte layout of a message is shown in figure \ref{fig:messageTypes}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/MessageTypes.pdf}
+    \caption{The different message types used in the telemetry system.\todo{delete uint16\_t part in fig and make cmd to broadcast can}}
+    \label{fig:messageTypes}
+\end{figure}
+
+\todo{make clearer distinction between message types and commands. Several commands can share the same message type}
+
+The first byte, the command, always identifies the message type and indicates how the remaining 15 bytes should be interpreted.
+This eases the decoding of messages on the receiving end and makes for a expandable framework which has enough room for future expansions.
+
+As an example of a specific message type, figure \ref{fig:messageTypes} shows a message at the bottom with the command \texttt{BROADCAST\_CAN} which is encoded by \texttt{0x01}.
+This message carries a CAN message as a payload and commands the solar car subsystem to inject the message into the solar car CAN bus.
+Even though some message types are only sent one direction between the telemetry modules, a unified design was chosen due to the inherent simplicity.
+
+% message protcol
+% command system
+% security -> encryption
+
+\section{\todo{Final product}}
+\subsection{Solar Car}
+% threads 
+\subsection{Support Vehicle}
+% threads
+\subsection{\todo{Graphical User Interface}}
+% GUI
 
 \subsection{Components}
+The primary components for our project has been two Teensy boards, a Teensy 3.6 board and 4.0 board, two nRF24 antennas, and a CAN transceiver.
+These components make up the chain of communication from reading/writing data from/to the CAN bus, through the Teensy boards CAN ports, to the antenna for transmission to the other antenna and Teensy board.
+
+In addition to these components, we have used a 12V to 5V DC converter, a logic converter from 5V to 3V, a molex board connector, a 100\(\mu\)F capacitor, a coin cell battery holder and a perf board for implementing our solution with the above mentioned components.
+We have also made use of Arduino Uno boards during the development to generate random CAN data.
 
@@
 \subsubsection{PlatformIO}
-The Teensy micro-processor line-up is integrated into the Arduino framework and implements as such the Arduino standard API.
-As a cross-platform build tool, PlatformIO is a more potent alternative to the standard Arduino IDE including more advanced features such as a static code analyzer and even a remote unit testing framework.
 
@@
 An open-source RTOS for embedded systems including Teensy. It supports threads, semaphores, data-sharing and timers and more to ease the scheduling of different tasks by the program kernel.
+
+\subsubsection{ACAN}
+The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards.
+The library has been important for the connection between the CAN network and the ports on the Teensy board.
+We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD.
+The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.
 
@@
 The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer.
-This is possible with the special structure of the CAN message.\\
+This is possible with the special structure of the CAN message.
+
 \begin{figure}[h]
@@
     \label{fig:CANframe}
-\end{figure}\\
+\end{figure}
+
 The frame uses an identifier to tell which messages goes to which unit.
@@
 The rest of the fields in fig.
-\ref{fig:CANframe} is not important concerning the nodes ability to communicate across the CAN network.\\
+\ref{fig:CANframe} is not important concerning the nodes ability to communicate across the CAN network.
+
 The CAN bus itself is two twisted wires, CAN High and CAN Low.
@@
 A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.
-\subsection{RF module} % Steffan 
-The nRF24L01+ LoRa RF module is used to provide two-way communication between the solar car and support vehicle.
-The module includes a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
-In addition to this, comprehensive driver libraries are provided by the manufacturer, making it very easy to integrate with our project.
 
-\subsection{Message protocol}
-The radio connection is the pivot point of the telemetry system and a flexible way to pass commands and associated payloads both ways is needed. 
-The nRF24L01+ module uses the Enhanced Shockburst $\copyright$ packet structure to transmit messages over the 2.4 GHz ISM band.
-Messages contain a data payload of up to 32 bytes, allowing us comfortably to transmit one CAN telemetry message per packet.
-Further, the package structure includes a preamble, address, packet and CRC bytes segments as shown in Figure \ref{fig:shockburst}.
-These are all generated automatically when a call to the radio write function is made, and therefore fall outside the scope of this project.
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.png}
-    \caption{The Enhanced Shockburst $\copyright$ packet structure.}
-    \label{fig:shockburst}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{documentation/images/MessageTypes.pdf}
-    \caption{The }
-    \label{fig:my_label}
-\end{figure}
+Since the most transmitted message in the telemetry system will be CAN messages forwarded from the solar cars CAN bus, it is chosen to build the message protocol around 16 byte packets...
+\todo{talk about cmd and message types and more}\\
+\todo{Should encryption be included in message protocol or a separate subsection?}
+As an extra feature not included in our original specification, we have made an encryption protocol for our telemetry module.
+The encryption is based on the RSA encryption method, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
+When the message is encrypted, the type is converted from a unsigned 8-bit integer to an unsigned 16-bit integer, to avoid overflow problems.
+The encrypted message is then decrypted at the other module and converted back to unsigned 8-bit type.
+This effectively halves the amount of data we can send in our transmissions, which is the downside of our encryption solution.\\
+The combinations of two prime numbers suited for the encryption protocol can be seen in the below table,
 
@@
 As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design.
-We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.\\
+We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.
+
 We performed software validation as we went along with the software coding, to make sure that the actual behaviour of the code was as the expected behaviour.
@@
 We used this to validate our code design with unit tests, testing both edge cases and random cases.
-We chose only to validate the code and functions we made ourselves, and not those that was imported from external libraries.\\
+We chose only to validate the code and functions we made ourselves, and not those that was imported from external libraries.
+
 Doing this provided us a test bench for our project, which greatly improved the credibility of our work, proving that our design works in general cases and not just the case we represent in our report.
@@
 \section{Discussion}
+% variable length messages would reduce overhead -> but most messages cotain CAN and are optimized
+%
 
@@
 
-% use section* for acknowledgment
 \section*{Acknowledgment}
-
 The authors would like to thank Martin Schoeberl and Christian Kampp Kruse for their mentoring and assistance throughout the project.
-The authors would also like to thank Claus from the department of Mechanical Engineering at DTU.
+The authors would also like to thank Claus Suldrup Nielsen from the department of Mechanical Engineering at DTU.
 
@@
 https://www.learn-cpp.org/
-\bibitem{Aunit}
+\bibitem{ACAN}
 Documentation,
-https://www.arduino.cc/reference/en/libraries/aunit/
+https://github.com/pierremolinaro/acan
 \bibitem{shockburst}

### 2021-06-16 code/SolarVehicle/src/telemetry/solar-car/SystemThd.h
saving and restoring pre-sleep state

@@ -46,6 +46,8 @@ THD_FUNCTION(systemThd, arg){
   rfWorkerState->pause = false;
   canReceiverState->pause = false;
 
+  uint8_t oldBlackBoxState = 0, oldRfTxState = 0;
+
   chThdSleepMicroseconds(100); // release in order to allow creation of other threads
 
   while(true){
@@ -122,8 +124,10 @@ THD_FUNCTION(systemThd, arg){
             if(!canReceiverState->pause){
               Serial.println("SystemThd:\tPutting system to sleep");
               canReceiverState->pause = true;
-              if(!rfWorkerState->pause) rfWorkerState->pause = true;
-              if(!blackBoxWorkerState->pause) blackBoxWorkerState->pause = true;
+              oldBlackBoxState = blackBoxWorkerState->pause;
+              oldRfTxState = rfWorkerState->pause;
+              rfWorkerState->pause = true;
+              blackBoxWorkerState->pause = true;
             } else {
               Serial.println("SystemThd:\tSystem is already asleep");
             }
@@ -134,6 +138,8 @@ THD_FUNCTION(systemThd, arg){
             if(canReceiverState->pause){
               Serial.println("SystemThd:\tWaking up the system");
               canReceiverState->wakeUp();
+              if(oldBlackBoxState) blackBoxWorkerState->wakeUp();
+              if(oldRfTxState) rfWorkerState->wakeUp();
             } else {
               Serial.println("SystemThd:\tThe system is already awake");
             }

### 2021-06-16 code/TelemetryUI/src/main/scala/telemetryui/udp/UdpServer.scala

@@ -36,6 +36,7 @@ class UdpServer extends Thread {
         case "end" =>
           println("Stopping UDP stream")
           hasListener = false
+        case _ =>
       }
 
     }

### 2021-06-16 code/SolarVehicle/lib/rf/RFfunctions.h
comments

@@ -12,18 +12,18 @@
 #include <nRF24L01.h>
 #include "TelemetryMessages.h"
 
-#ifdef TEENSY36_BOARD
-#define IRQ_PIN 2   // SPI interrupt pin
-#define CE_PIN 10    // SPI chip-enable (CE) pin
-#define CSN_PIN 9  // SPI chip-select (CSN) pin
+#ifdef TEENSY36_BOARD   // SPI pins are configured differently on Teensy 3.6
+#define IRQ_PIN 2       // SPI interrupt (IRQ) pin
+#define CE_PIN 10       // SPI chip-enable (CE) pin
+#define CSN_PIN 9       // SPI chip-select (CSN) pin
 #else
-#define IRQ_PIN 2   // SPI interrupt pin
-#define CE_PIN 9    // SPI chip-enable (CE) pin
-#define CSN_PIN 10  // SPI chip-select (CSN) pin
+#define IRQ_PIN 2       // SPI interrupt (IRQ) pin 
+#define CE_PIN 9        // SPI chip-enable (CE) pin
+#define CSN_PIN 10      // SPI chip-select (CSN) pin
 #endif
 
-#define COUNT 5     // Number of transmission retries
-#define DELAY 15    // Delay between retries (= DELAY * 250 us + 250 us)
+#define COUNT 5         // Number of transmission retries
+#define DELAY 15        // Delay between retries (= DELAY * 250 us + 250 us)
 
 #ifdef TEENSY40_BOARD
 bool radioNumber = 0;       
@@ -33,9 +33,9 @@ bool radioNumber = 1;
 bool radioNumber = 1;
 #endif
 
-RF24 radio(CE_PIN, CSN_PIN);                                    // CE and CSN pins
-static const byte address[][6] = {"00001", "00002"};            // TX/RX byte addresses
-bool ack = false; 
+RF24 radio(CE_PIN, CSN_PIN);                                    // RF24 global variable declaration
+static const byte address[][6] = {"00001", "00002"};            // Byte addresses for reading/writing pipes
+bool ack = false;                                               // Enable ack packages by setting this true
 
 /**********************************************************
  * @brief   Initializes and configures RF24 class object. *

### 2021-06-16 code/SolarVehicle/lib/message/RSA.h
added return value RFinit()

@@ -1,53 +1,0 @@
-#ifndef __RSA_H__
-#define __RSA_H__
-
-#include <stdlib.h>
-#include <math.h>
-#include <stdio.h>
-#include <stdint.h>
-
-class RSA {
-    private : 
-        double p = 11, q = 17;                 // Choose two prime numbers p & q
-        double n = p * q, phi = (p-1) * (q-1); // Calculate n = p * q and phi(n) = (p-1) * (q-1)
-        double e = 7;                          // Select e such that 1 < e < phi(n), and gcd(e, phi(n)) = 1
-        double d = fmod(1/e, phi);          // Compute d as the multiplicative inverse of e (mod phi(n))
-
-        /**
-         * 
-         * 
-        */
-        int gcd(int a, int b)               // Helper function to compute the greatest common denominator of two integers a and b
-        {
-            if (!(a%b)){
-                return b; 
-            } else {
-                return gcd(b, a%b); 
-            }
-        }
-
-    public: 
-
-        /**
-         *  
-         *  
-        */
-        void encrypt(uint8_t *arr, int len)  
-        {
-            for (int i = 0; i < len; i++){
-                Serial.println(fmod(arr[i]*e,n), HEX);
-                arr[i] = (uint8_t) fmod(arr[i]*e,n);  
-            }
-        }
-
-        void decrypt(uint8_t *arr, int len)
-        {
-            for (int i = 0; i < len; i++){
-                Serial.println(fmod(arr[i]*d,n), HEX);
-                arr[i] = (uint8_t) fmod(arr[i]*d,n); 
-            }
-        }
-
-};
-
-#endif

### 2021-06-16 code/SolarVehicle/lib/rf/RFfunctions.h

@@ -37,13 +37,14 @@ RF24 radio(CE_PIN, CSN_PIN);                                    // RF24 global v
 static const byte address[][6] = {"00001", "00002"};            // Byte addresses for reading/writing pipes
 bool ack = false;                                               // Enable ack packages by setting this true
 
-/**********************************************************
- * @brief   Initializes and configures RF24 class object. *
- * @param   None                                          *
-***********************************************************/
-void RFinit()
+/****************************************************************************
+ * @brief   Initializes and configures RF24 class object.                   *
+ * @param   None                                                            *
+ * @return  True if radio object was properly initialized, otherwise false. *
+*****************************************************************************/
+bool RFinit()
 { 
-    if(!radio.begin()){Serial.println("Radio not working!");}
+    if(!radio.begin()) return false; 
     radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level, choose between MIN, LOW, HIGH, MAX (higher PA level improves range)
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate, choose between RF24_250KBS, RF24_1MBS, RF24_2MBS (higher Data Rates may cause data loss)
     radio.enableDynamicPayloads();                              // Enable variable data payloads
@@ -53,6 +54,7 @@ void RFinit()
     radio.openWritingPipe(address[!radioNumber]);               // Open writing pipe at address 
     radio.openReadingPipe(1, address[radioNumber]);             // Open reading pipe at address
     radio.startListening();                                     // Starts RX mode
+    return true; 
 }
 
 /********************************************************************************

### 2021-06-16 code/SolarVehicle/src/telemetry/support-vehicle/supportVehicleMain.cpp

@@ -5,9 +5,7 @@
  ****************************************************************************************************/
 
 #include <Arduino.h>
-#include <SPI.h>
 #include "telemetry/support-vehicle/threads.h"
-#include "RSA.h"
 
 /*********************************************************
  * @brief Initializes all threads with a working area,   *
@@ -15,7 +13,7 @@
  *********************************************************/  
 void chSetup()
 {
-  chSysInit();  // Initializes ChibiOS system
+  chSysInit();          // Initializes ChibiOS system
   chThdCreateStatic(WaReceiverThread, sizeof(WaReceiverThread), NORMALPRIO + 1, receiverThread, &sec);
   threadBundle transmitterBundle = {.security = &sec, .fifo = &TXfifo}; 
   chThdCreateStatic(waTransmitterThread, sizeof(waTransmitterThread), NORMALPRIO + 1, transmitterThread, &transmitterBundle);
@@ -27,6 +25,7 @@ void setup()
   while(!Serial){}      // Wait until serial is available
   RFinit();             // Initialize RF module
   sec.encryption_key(); // Create message encryption key
+  TXfifo.clear();       // Clears TXfifo 
   chBegin(chSetup);     // Initialize and start all 4 threads
   while (true){}        
 }

### 2021-06-16 documentation/FinalPaper/paper.tex

@@
 
-\usepackage{cite}
 \usepackage[pdftex]{graphicx}
-\usepackage{cite}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \usepackage{graphicx}
+\usepackage{float}
 \usepackage{comment}
@@
 \usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-
+\usepackage{hyperref}
+\usepackage[style=ieee]{biblatex}
+\addbibresource{paper.bib}
 \usepackage{url}
@@
 \section{Design}
-- Block Diagram and description of overall data flow
+The overall system consists of three separate modules; the Solar Car Unit which communicates with the Support Vehicle Unit via a wireless RF connection, which in turn interacts with the Support Vehicle Laptop through a serial USB connection.
+The schematic for the system is outlined in Figure \ref{fig:schematic}.
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.5\textwidth]{documentation/images/schematic.png}
+    \caption{Overview of the system's components in block diagram form.}
+    \label{fig:schematic}
+\end{figure}
 
@@
 \subsection{RF network} % Steffan
-The nRF24L01+ LoRa RF module is used to provide two-way communication between the solar car and support vehicle.
-The module includes a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
-In addition to this, comprehensive driver libraries are provided by the manufacturer, making it very easy to integrate with our project.
+The system uses two nRF24L01+ transceivers to provide a bidirectional connection between the Solar Car and Support Vehicle over the 2.4 GHz ISM band.
+The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
+
+An advantage of using an off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^{\copyright}$.
+
 The nRF24L01+ module uses the Enhanced Shockburst $\copyright$ packet structure to transmit messages over the 2.4 GHz ISM band.
@@
 Further, the package structure includes a preamble, address, packet and CRC bytes segments as shown in Figure \ref{fig:shockburst}.
-\begin{comment}
+
 \begin{figure}[h]
@@
 \end{figure}
-\end{comment}
 
@@
 \section{Discussion}
-% variable length messages would reduce overhead -> but most messages cotain CAN and are optimized
+% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
 %
@@
 
-\bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,../bib/paper}
-
-\section{References}
-\begingroup
-\renewcommand{\section}[2]{}%
-%\renewcommand{\chapter}[2]{}% for other classes
-\begin{thebibliography}{}
-\bibitem{wsc}
-Bridgestone World Solar Challenge, 
-URL: https://worldsolarchallenge.org/
-\bibitem{ROAST}
-Reports and source files provided by DTU ROAST
-\bibitem{teensy}
-Teensy microcontroller,
-URL: https://www.pjrc.com/teensy/
-\bibitem{C++}
-Website for learning C++
-https://www.learn-cpp.org/
-\bibitem{ACAN}
-Documentation,
-https://github.com/pierremolinaro/acan
-\bibitem{shockburst}
-nRF24 transmission protocol, URL: https://lastminuteengineers.com/nrf24l01-arduino-wireless-communication/
-\end{thebibliography}
-\endgroup
+\printbibliography
 

### 2021-06-17 code/SolarVehicle/lib/rf/RFfunctions.h
cc

@@ -10,7 +10,7 @@
 #include <SPI.h>
 #include <RF24.h>
 #include <nRF24L01.h>
-#include "TelemetryMessages.h"
+#include "printf.h"
 
 #ifdef TEENSY36_BOARD   // SPI pins are configured differently on Teensy 3.6
 #define IRQ_PIN 2       // SPI interrupt (IRQ) pin
@@ -44,10 +44,10 @@ bool ack = false;                                               // Enable ack pa
 *****************************************************************************/
 bool RFinit()
 { 
-    if(!radio.begin()) return false; 
+    if(!radio.begin()) return false;                            // Begin radio class object, return false if not succeeded. 
     radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level, choose between MIN, LOW, HIGH, MAX (higher PA level improves range)
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate, choose between RF24_250KBS, RF24_1MBS, RF24_2MBS (higher Data Rates may cause data loss)
-    radio.enableDynamicPayloads();                              // Enable variable data payloads
+    radio.setPayloadSize(BaseTelemetryMsg::length()<<1);        // Set constant payload size to improve transmission time
     radio.setAutoAck(true);                                     // RX node sends an automatic ack packet
     if (ack) radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry

### 2021-06-17 code/SolarVehicle/lib/util/Fifo.h

@@ -10,16 +10,16 @@ class Fifo {
     private: 
         semaphore_t spaceSem, dataSem;
         T *fifo;
-        size_t _head, _tail;  
+        size_t _size, _head, _tail;  
         uint32_t size;
 
     public: Fifo(uint32_t size){
-        this->size  = size;
+        this->_size  = size;
         this->_head   = 0;
         this->_tail   = 0;
-        spaceSem    = _SEMAPHORE_DATA(spaceSem,(cnt_t)size);
+        spaceSem    = _SEMAPHORE_DATA(spaceSem,(cnt_t)_size);
         dataSem     = _SEMAPHORE_DATA(dataSem,0);
-        fifo        = new T[size];
+        fifo        = new T[_size];
     }
 
     public: ~Fifo(){
@@ -43,12 +43,12 @@ class Fifo {
 
     // returns a pointer index to the next fifo slot
     public: uint32_t advance(uint32_t index){
-        return index < (size - 1) ? index + 1 : 0;  
+        return index < (_size - 1) ? index + 1 : 0;  
     }
 
     // advances the index pointer to the next fifo slot
     public: void advance(uint32_t *index){
-        *index = *index < (size - 1) ? *index + 1 : 0;
+        *index = *index < (_size - 1) ? *index + 1 : 0;
     }
 
     // signals that an element has been read and that the slot is free for new data
@@ -63,17 +63,17 @@ class Fifo {
     
     // returns the size of the fifo
     public: uint32_t getSize(){
-        return size;
+        return _size;
     }
 
     // set all fifo entries to 0
     public: void clear(){
-        memset(fifo,0,size*sizeof(T));
+        memset(fifo, 0, _size * sizeof(T));
     }
     
     // Prints the contents of the fifo buffer for debugging
     public: void printContents(){
-        for (int i = 0; i < size; i++){
+        for (int i = 0; i < _size; i++){
             Serial.print(i + " : " + fifo[i]);
         }
         Serial.println();

### 2021-06-17 code/SolarVehicle/src/telemetry/support-vehicle/supportVehicleMain.cpp

@@ -1,7 +1,7 @@
 /****************************************************************************************************
  * @file    supportVehicleMain.cpp                                                                  *
  * @author  Steffan Martin Kunoy                                                                    *
- * @brief   Includes setup() and loop() functions to be run on the support vehicle microcontroller. *
+ * @brief   Includes setup() and loop() functions to be run on the Support Vehicle microcontroller. *
  ****************************************************************************************************/
 
 #include <Arduino.h>
@@ -22,8 +22,11 @@ void chSetup()
 void setup()
 {
   Serial.begin(921600); // Initialize serial port
-  while(!Serial){}      // Wait until serial is available
-  RFinit();             // Initialize RF module
+  while (!Serial){}     // Wait until serial is available
+  if (!RFinit()){       // Initialize and configure RF24 module
+    Serial.println("Radio not working!"); 
+    radio.printDetails(); 
+  }
   sec.encryption_key(); // Create message encryption key
   TXfifo.clear();       // Clears TXfifo 
   chBegin(chSetup);     // Initialize and start all 4 threads

### 2021-06-17 code/SolarVehicle/src/telemetry/support-vehicle/threads.h

@@ -9,6 +9,7 @@
 
 #include <ChRt.h>
 #include "RFfunctions.h"
+#include "TelemetryMessages.h"
 #include "ThreadState.h"
 #include "Fifo.h"
 #include "Encryption.h"
@@ -36,32 +37,32 @@ THD_FUNCTION(receiverThread, arg)
   BaseTelemetryMsg received;                                        // BaseTelemetryMsg object for received messages
   uint16_t encrypted[16];                                           // Buffer to contain encrypted message
   char tempStr[256];                                                // Buffer to print received message to serial
-  chThdSleepMicroseconds(100);                                      // Sleep to allow for creation of secondary thread
+  chThdSleepMicroseconds(100);                                        // Sleep to allow for creation of secondary thread
 
   while (true){
   
-    if (radio.available()){                                         // Check if TX node is transmitting data
-      if (RFreceive(encrypted, BaseTelemetryMsg::length()<<1)){     // Check if message was received
+    if (radio.available()){                                           // Check if TX node is transmitting data
+      if (RFreceive(encrypted, BaseTelemetryMsg::length()<<1)){       // Check if message was received
         Serial.print("Received: ");
-        sec->decrypt(encrypted, (uint8_t*)&received, 16);           // Decrypt received message
-        switch (received.cmd){                                      // Switch based on the message cmd field 
-          case RECEIVED_CAN:{                                       // Standard case, cmd matches RECEIVED_CAN
-            CanTelemetryMsg *ptr = (CanTelemetryMsg*)&received;     // Cast received to CanTelemetryMsg pointer
+        sec->decrypt(encrypted, (uint8_t*)&received, 16);             // Decrypt received message
+        switch (received.cmd){                                        // Switch based on the message cmd field 
+          case RECEIVED_CAN:{                                         // Standard case, cmd matches RECEIVED_CAN
+            CanTelemetryMsg *ptr = (CanTelemetryMsg*)&received;       // Cast received to CanTelemetryMsg pointer
             ptr->toJSON(tempStr, sizeof(tempStr));                  
-            Serial.println(tempStr);                                // Print message formatted as JSON
+            Serial.println(tempStr);                                  // Print message formatted as JSON
           }
             break;
-          default:{                                                 // Default case, cmd does not match RECEIVED_CAN
+          default:{                                                   // Default case, cmd does not match RECEIVED_CAN
             received.toString(tempStr, sizeof(tempStr));            
-            Serial.println(tempStr);                                // Print message formatted as string
[…]

### 2021-06-17 documentation/FinalPaper/paper.tex

@@
 \end{figure}
+%
 
@@
 \subsection{CAN bus}
+The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer.
+This is possible with the special structure of the CAN message.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
+    \caption{Structure of a CAN frame}
+    \label{fig:CANframe}
+\end{figure}
+
+The frame uses an identifier to tell which messages goes to which unit.
+When two messages try to access the bus, the one with the highest priority, lowest ID, will be granted access to the bus, while the other message has to wait.
+After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a node.
+The control field, sometimes called DLC, holds the amount of bytes used in the following frame field containing the actual data.
+The rest of the fields in fig.
+\ref{fig:CANframe} serve error checking and coordination of communication on the bus.
+
+The physical CAN bus itself consists of a differential pair of twisted wires, CAN High and CAN Low.
+These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively.
+A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.
+
+The Teensy 3.6 used in the solar car features a CAN controller handling the communication with the CAN bus.
+In order to interface with the physical CAN bus though, a transceiver is needed.
+Here the MCP2551 was chosen, which maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.
 % MCP
@@
 The system uses two nRF24L01+ transceivers to provide a bidirectional connection between the Solar Car and Support Vehicle over the 2.4 GHz ISM band.
-The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data rates of up to 2 Mbit/s.
+The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data transfer rates of up to 2 Mbit/s.
 
-An advantage of using an off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^{\copyright}$.
-
-The nRF24L01+ module uses the Enhanced Shockburst $\copyright$ packet structure to transmit messages over the 2.4 GHz ISM band.
-Messages contain a data payload of up to 32 bytes, allowing us comfortably to transmit one CAN telemetry message per packet \cite{shockburst}.
-Further, the package structure includes a preamble, address, packet and CRC bytes segments as shown in Figure \ref{fig:shockburst}.
+An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}.
+Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Figure \ref{fig:shockburst}.
 
@@
     \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.png}
-    \caption{The Enhanced Shockburst $\copyright$ packet structure.}
+    \caption{The Enhanced Shockburst$^\copyright$ packet structure.}
     \label{fig:shockburst}
@@
 \subsection{SD card}
+
+We chose a SD card as a permanent local storage solution for the solar cars "black box" due to its ease of use, its interchangeability and since the Teensy 3.6 has an integrated SD card slot on its PCB.
+Furthermore, this solution makes it easy to increase the storage capacity to more than sufficient sizes with new SD card sizes exceeding 1 TiB.
+
 \subsection{Microcontrollers} % Steffan
+Finally, Teensy 3.6 and 4.0 microcontrollers are used as connection hubs in the Solar Car and Support Vehicle Units respectively.
+The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data.
+Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.
 
@@
 
-Several of the key functionalities of the telemetry are implemented in software.
+Several of the key functionalities of the telemetry module are implemented in software.
 This together with the aim of designing a system which is easy to extend in the future makes the choice of software frameworks very important, since they affect reusability and maintainability of the developed code.
@@
 \subsection{Build system}
-As a build system, platformIO \cite{platformIO} was chosen for the project.
+As a build system, PlatformIO \cite{platformIO} was chosen for the project.
 It is a version control friendly, cross-platform embedded build tool which includes library management and works by defining \textit{environments} to allow development with different embedded platforms on the same code base.
@@
 \subsection{Other Libraries}
+
+In order to interface with the components presented in the previous section, a set of libraries were used.
+
+The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards.
+The library has been important for the connection between the CAN network and the ports on the Teensy board.
+We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD.
+The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.
 % ACAN.h
@@
 
-\subsection{Software Stack}
-
-In this section the main software tools and libraries used throughout the project are introduced.
-
-\subsubsection{PlatformIO}
-
-\subsubsection{ChibiOS}
-An open-source RTOS for embedded systems including Teensy. It supports threads, semaphores, data-sharing and timers and more to ease the scheduling of different tasks by the program kernel.
-
-\subsubsection{ACAN}
-The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards.
-The library has been important for the connection between the CAN network and the ports on the Teensy board.
-We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD.
-The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.
-
-\subsubsection{Other Libraries}
-
 \subsection{CAN} % Victor
-The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer.
-This is possible with the special structure of the CAN message.
-
-\begin{figure}[h]
-    %\centering
-    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
-    \caption{Structure of a CAN frame}
-    \label{fig:CANframe}
-\end{figure}
-
-The frame uses an identifier to tell which messages goes to which unit.
-When two messages tries to access the bus, the one with the highest priority, lowest ID, will be granted access to the bus, while the other message has to wait.
-After the ID comes the RTR bit which determines whether the frame is set to be at sent or received by the node.
-The control field, sometimes called DLC, tells the amount of bytes used in the following frame field with actual data.
-The rest of the fields in fig.
-\ref{fig:CANframe} is not important concerning the nodes ability to communicate across the CAN network.
-
-The CAN bus itself is two twisted wires, CAN High and CAN Low.
-These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively.
-A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.
 
@@
 
+\href{https://youtu.be/-iwST3REn40}{video link}
 \section{Discussion}
 % variable length messages would reduce overhead -> but most messages contain CAN and are optimized
-%
+% range test
+% throughput limitations
 

### 2021-06-17 documentation/FinalPaper/paper.tex
Updates from Overleaf

@@
 \usepackage{url}
+\usepackage[colorlinks = true,
+            linkcolor = blue,
+            urlcolor  = blue,
+            citecolor = blue,
+            anchorcolor = blue]{hyperref}
+\newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}
 
@@
 
-% Regler ti skrivning: 
+% Regler til skrivning: 
 % 1. Altid nutid
@@
 It is a version control friendly, cross-platform embedded build tool which includes library management and works by defining \textit{environments} to allow development with different embedded platforms on the same code base.
-Other project groups working on the solar car in parallel integrated their software into the platformIO ecosystem as well, making extensive code reuse in the future possible.
+It comes with a VScode plugin for ideal IDE support.
+Other project groups working on the solar car in parallel integrated their software into the PlatformIO ecosystem as well, making extensive code reuse in the future possible.
 
@@
 The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.
+
 % ACAN.h
@@
 
+As an extra feature not included in our original specification, we have made an encryption protocol for our telemetry module.
+The encryption is based on the RSA algorithm, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
+The pairings of prime numbers suited for the encryption protocol can be seen below in Table \ref{}.
+When a message is encrypted, each of the 16 bytes is extended to a 16-bit unsigned integer to avoid overflow problems.
+Once transmitted to the receiving node, the message is then decrypted and converted back to unsigned 8-bit type.
+This essentially doubles the size of the data payloads being sent in our RF transmission, which is the downside of our encryption solution.
+
 % message protcol
@@
 
-\subsection{Components}
-The primary components for our project has been two Teensy boards, a Teensy 3.6 board and 4.0 board, two nRF24 antennas, and a CAN transceiver.
-These components make up the chain of communication from reading/writing data from/to the CAN bus, through the Teensy boards CAN ports, to the antenna for transmission to the other antenna and Teensy board.
-
-In addition to these components, we have used a 12V to 5V DC converter, a logic converter from 5V to 3V, a molex board connector, a 100\(\mu\)F capacitor, a coin cell battery holder and a perf board for implementing our solution with the above mentioned components.
-We have also made use of Arduino Uno boards during the development to generate random CAN data.
-
-\subsection{CAN} % Victor
-
-Since the most transmitted message in the telemetry system will be CAN messages forwarded from the solar cars CAN bus, it is chosen to build the message protocol around 16 byte packets...
-\todo{talk about cmd and message types and more}\\
-\todo{Should encryption be included in message protocol or a separate subsection?}
-As an extra feature not included in our original specification, we have made an encryption protocol for our telemetry module.
-The encryption is based on the RSA encryption method, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
-When the message is encrypted, the type is converted from a unsigned 8-bit integer to an unsigned 16-bit integer, to avoid overflow problems.
-The encrypted message is then decrypted at the other module and converted back to unsigned 8-bit type.
-This effectively halves the amount of data we can send in our transmissions, which is the downside of our encryption solution.\\
-The combinations of two prime numbers suited for the encryption protocol can be seen in the below table,
-
-\section{Implementation}
-
-\subsection{Solar Car} 
-\subsubsection{Black Box} % tjark
-
-The black box functionality in the solar car is implemented using a SD card. The Teensy 3.6 used as a microcontroller in the telemetry module of the solar car already includes a 4-bit
-
-\subsubsection{CAN transceiver}
-\subsection{Support Vehicle}
-\subsubsection{GUI} % tjark
-\subsubsection{RF transceiver}
-
 \section{Verification and validation} % Victor
@@
 
-\href{https://youtu.be/-iwST3REn40}{video link}
 \section{Discussion}
@@
 % range test
+In order to test the connection between the Solar Car and Support Vehicle modules we performed a range test in a suitable outdoor environment.
+In the end, we achieved a transmission range of around 160 m before the connection became too unreliable.
+This is obviously a big step down from the 1000 m claimed by the manufacturer.
+However, A video documenting the test can be viewed with \MYhref{https://youtu.be/-iwST3REn40}{this link}.
 % throughput limitations
@@
 \section{Conclusion}
-The conclusion goes here.
+%The conclusion goes here.
 

### 2021-06-17 documentation/FinalPaper/paper.tex

@@
 \usepackage{xcolor}
-\usepackage{hyperref}
 \usepackage[style=ieee]{biblatex}
@@
 \usepackage{url}
-\usepackage[colorlinks = true,
-            linkcolor = blue,
-            urlcolor  = blue,
-            citecolor = blue,
-            anchorcolor = blue]{hyperref}
+\usepackage[linkcolor = blue]{hyperref}
 \newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}
@@
 \section{Design}
-The overall system consists of three separate modules; the Solar Car Unit which communicates with the Support Vehicle Unit via a wireless RF connection, which in turn interacts with the Support Vehicle Laptop through a serial USB connection.
-The schematic for the system is outlined in Figure \ref{fig:schematic}.
+The overall system consists of three separate modules; the Solar Car Unit (SCU) which is connected to the CAN bus and communicates with the support vehicle unit via a wireless RF connection.
+The Support Vehicle Unit (SVU) in turn interacts with the Support Vehicle laptop through a serial UART connection.
+The GUI program running on the laptop broadcasts the CAN bus message stream on a UDP port on the local network.
+A block diagram for the system is outlined in Fig.
+\ref{fig:schematic}.
 \begin{figure}[H]
     \centering
-    \includegraphics[width=0.5\textwidth]{documentation/images/schematic.png}
+    \includegraphics[width=0.5\textwidth]{documentation/images/SystemSchematic.pdf}
     \caption{Overview of the system's components in block diagram form.}
@@
 \subsection{CAN bus}
-The CAN network in the solar car enables the micro-controllers at each electrical unit in the car to send and receive data without a host computer.
-This is possible with the special structure of the CAN message.
+The CAN network in the SCU enables the all the car's processing nodes to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.
 
@@
 The frame uses an identifier to tell which messages goes to which unit.
-When two messages try to access the bus, the one with the highest priority, lowest ID, will be granted access to the bus, while the other message has to wait.
+When two messages try to access the bus, the one with the highest priority (lowest ID) will be granted access to the bus, while the other message have to wait.
 After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a node.
@@
 These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively.
-A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.5V and the CAN Low wire is driven at 2.5V.
+A 1 is seen when the voltage in both wires is 2.5V while a 0 is seen when the CAN High wire is driven by 3.75V and the CAN Low wire is driven at 1.25V.
 
@@
 An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}.
-Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Figure \ref{fig:shockburst}.
+Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Fig.
+\ref{fig:shockburst}.
+Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and CRC segments are generated automatically and therefore fall beyond the scope of this project.
 
@@
 
-We chose a SD card as a permanent local storage solution for the solar cars "black box" due to its ease of use, its interchangeability and since the Teensy 3.6 has an integrated SD card slot on its PCB.
-Furthermore, this solution makes it easy to increase the storage capacity to more than sufficient sizes with new SD card sizes exceeding 1 TiB.
+We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accesibility through the Teensy 3.6 integrated SD card slot.
+Furthermore, this solution scales very easily for larger storage capacity, with new SD card sizes exceeding 1 TiB.
 
 \subsection{Microcontrollers} % Steffan
-Finally, Teensy 3.6 and 4.0 microcontrollers are used as connection hubs in the Solar Car and Support Vehicle Units respectively.
+Finally, Teensy 3.6 and 4.0 microcontrollers are used as connection hubs in the SCU and SVU respectively.
 The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data.
@@
 Several of the key functionalities of the telemetry module are implemented in software.
-This together with the aim of designing a system which is easy to extend in the future makes the choice of software frameworks very important, since they affect reusability and maintainability of the developed code.
+This, together with the aim of designing a system to be expanded upon in the future, makes the choice of software frameworks very important.
+Thus, an emphasis has been placed on the reusability and maintainability of the developed code.
 
@@
 
-\section{Verification and validation} % Victor
+\section{Testing and Validation} % Victor
 As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design.
@@
 
+%Finally, we were able to test all three modules
+
 \section{Discussion}
-% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
+
 % range test
@@
 In the end, we achieved a transmission range of around 160 m before the connection became too unreliable.
+A video documenting the test can be viewed with \MYhref{https://youtu.be/-iwST3REn40}{this link}.
 This is obviously a big step down from the 1000 m claimed by the manufacturer.
-However, A video documenting the test can be viewed with \MYhref{https://youtu.be/-iwST3REn40}{this link}.
+However, considering the crude implementation of our hardware components, as well as the interference from buildings and other vehicles, that range was highly unattainable in the first place.
+Further enhancements in the hardware, such as a stable power supply for the RF modules, would likely improve the range beyond 400 m.
+
 % throughput limitations
+% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
 

### 2021-06-17 documentation/FinalPaper/paper.tex

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.png}
+    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.pdf}
     \caption{The Enhanced Shockburst$^\copyright$ packet structure.}
@@
 
-As a specific RTOS, we chose ChibiOS \cite{chibios}, since it supported both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.
+As a specific RTOS, we chose ChibiOS \cite{chibios}, since it is supported both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.
 
@@
     \includegraphics[width=\linewidth]{documentation/images/MessageTypes.pdf}
-    \caption{The different message types used in the telemetry system.\todo{delete uint16\_t part in fig and make cmd to broadcast can}}
+    \caption{The different message types used in the telemetry system.}
     \label{fig:messageTypes}
@@
 The encryption is based on the RSA algorithm, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
-The pairings of prime numbers suited for the encryption protocol can be seen below in Table \ref{}.
+The pairings of prime numbers suited for the encryption protocol can be seen below in Table \ref{fig:prime_tabel}.
 When a message is encrypted, each of the 16 bytes is extended to a 16-bit unsigned integer to avoid overflow problems.
@@
 This essentially doubles the size of the data payloads being sent in our RF transmission, which is the downside of our encryption solution.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/prime_tabel.pdf}
+    \caption{Valid combinations of prime numbers, marked by OOOOOO}
+    \label{fig:prime_tabel}
+\end{figure}
 
@@
 \subsection{Solar Car}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/SCU_CircuitDiagram.pdf}
+    \caption{The circuit diagram for the SCU.}
+    \label{fig:SCU_circuit}
+\end{figure}
 % threads 

### 2021-06-17 documentation/FinalPaper/paper.tex
Updates from Overleaf

@@
 \section{Components}
+
+\subsection{Microcontrollers} % Steffan
+We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU respectively.
+The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data.
+Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.
+
 \subsection{CAN bus}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.pdf}
+    \includegraphics[trim={0 2cm 0 0 },clip]{documentation/images/EnhancedShockburst.pdf}
     \caption{The Enhanced Shockburst$^\copyright$ packet structure.}
@@
 Furthermore, this solution scales very easily for larger storage capacity, with new SD card sizes exceeding 1 TiB.
-
-\subsection{Microcontrollers} % Steffan
-Finally, Teensy 3.6 and 4.0 microcontrollers are used as connection hubs in the SCU and SVU respectively.
-The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data.
-Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.
 
@@
     \centering
-    \includegraphics[width=\linewidth]{documentation/images/prime_tabel.pdf}
+    \includegraphics[width=\linewidth]{documentation/images/prime_tabel.png}
     \caption{Valid combinations of prime numbers, marked by OOOOOO}

### 2021-06-17 documentation/FinalPaper/paper.tex

@@
     \centering
-    \includegraphics[trim={0 2cm 0 0 },clip]{documentation/images/EnhancedShockburst.pdf}
+    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.pdf}
     \caption{The Enhanced Shockburst$^\copyright$ packet structure.}
@@
     \includegraphics[width=\linewidth]{documentation/images/prime_tabel.png}
-    \caption{Valid combinations of prime numbers, marked by OOOOOO}
+    \caption{Valid combinations of prime numbers, marked by O}
     \label{fig:prime_tabel}
@@
 \subsection{Support Vehicle}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/SVU_CircuitDiagram.pdf}
+    \caption{The circuit diagram for the SVU.}
+    \label{fig:SCU_circuit}
+\end{figure}
 % threads

### 2021-06-17 documentation/FinalPaper/paper.tex
Updates from Overleaf

@@
 \usepackage{comment}
+\usepackage{movie15}
 \usepackage{textcomp}
@@
 \begin{abstract}
-Abstract
+This paper presents the telemetry project for the DTU ROAST Solar car.
+The development of the module is done through design sprints, where hardware and software is gradually developed along side each other.
+First a Hardware prototype on breadboards, then the software to run the modules satisfactorily was developed and finally the final product was made when the hardware was soldered to perf boards.\\
+The software developed for the module provided the module to have two-way communication with CAN message structure, encryption of the data during transmission between the module in the solar car and a module in a support car, the ability to give the solar car module commands from the support car and streaming received messages in the support car to Matlab.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-ROAST, telemetry, CAN
+ROAST, telemetry, CAN, RSA encryption, Matlab
 \end{IEEEkeywords}
@@
 
+\begin{figure}[b]
+    \centering
+    \includegraphics[width=0.5\textwidth]{documentation/images/SolarChallengeLogo.pdf}
+    \caption{Bridgestone World Solar Challenge logo \cite{wsc}.}
+    \label{fig:solar}
+\end{figure}
+
 The telemetry system must both serve the purpose of fulfilling the necessary specifications for communication, while also being as energy-efficient as possible.
@@
 
-This paper begins with giving a more precise problem statement and introducing the methodology used throughout the project. Thereafter, the design... \todo{Skal laves når strukturen er slået fast}
+This paper begins with giving a more precise problem statement and introducing the methodology used throughout the project.
+Thereafter, the overall system design is presented and the following paragraphs discuss the hardware components and software used for these.
+After this comes a paragraph discussing the implementation of the communication between the two modules, both internal and during transmission.
+This concludes with a section of the final product, which is followed up by our method method for testing and validating our solution, as well as discussion and conclusion of the solution.
 
@@
 The overall system consists of three separate modules; the Solar Car Unit (SCU) which is connected to the CAN bus and communicates with the support vehicle unit via a wireless RF connection.
-The Support Vehicle Unit (SVU) in turn interacts with the Support Vehicle laptop through a serial UART connection.
-The GUI program running on the laptop broadcasts the CAN bus message stream on a UDP port on the local network.
+The Support Vehicle Unit (SVU) in turn interacts with the Support Vehicle laptop through a serial connection.
+The GUI program running on the laptop broadcasts the CAN bus message stream on the local network.
 A block diagram for the system is outlined in Fig.
@@
 \end{figure}
-%
 
@@
 \subsection{Microcontrollers} % Steffan
-We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU respectively.
+We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU, respectively.
+The Teensy line-up has been used as the main microcontroller platform in most ecocar projects and we decided to use them as well after considering the performance of newer Teensy boards.
 The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data.
@@
 \subsection{CAN bus}
-The CAN network in the SCU enables the all the car's processing nodes to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.
+The CAN network in the SCU enables all the car's processing nodes to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.
 
@@
     \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
-    \caption{Structure of a CAN frame}
+    \caption{Structure of a CAN frame.}
     \label{fig:CANframe}
@@
 
-The Teensy 3.6 used in the solar car features a CAN controller handling the communication with the CAN bus.
+The Teensy 3.6 used in the solar car features a CAN controller handling message assembly and receiving from the CAN bus.
 In order to interface with the physical CAN bus though, a transceiver is needed.
-Here the MCP2551 was chosen, which maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.
-% MCP
+Here the MCP2551 was chosen since it has been used extensively in other ecocar projects.
+The component maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.
+
 \subsection{RF network} % Steffan
-The system uses two nRF24L01+ transceivers to provide a bidirectional connection between the Solar Car and Support Vehicle over the 2.4 GHz ISM band.
-The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m at data transfer rates of up to 2 Mbit/s.
+Two radio-frequency (RF) transceivers are required to provide a stable connection between the SCU and SVU at longer distances.
+There are many options for radio modules supporting long range communication at the expense of achievable data rate.
+We chose a trade-off between both by selecting the nRF24L01+ operating on the 2.4 GHz ISM band.
+The module features a detachable antenna, power amplifier and low-noise amplifier to provide a theoretical range in excess of 1000 m and data transfer rates of up to 2 Mbit/s.
 
 An advantage of using the off-the-shelf RF module is the automated message protocol, known as Enhanced Shockburst$^\copyright$ \cite{shockburst}.
-Package structures contain a preamble, address, packet, data payload and CRC bytes segments as shown in Fig.
+Messages are transmitted in packages, structured in flexible byte segments, as shown in Fig.
 \ref{fig:shockburst}.
@@
 We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accesibility through the Teensy 3.6 integrated SD card slot.
-Furthermore, this solution scales very easily for larger storage capacity, with new SD card sizes exceeding 1 TiB.
+Furthermore, this solution scales very well and easily in terms storage capacity, with new SD card sizes exceeding 1 TiB.
 
@@
 
-As a specific RTOS, we chose ChibiOS \cite{chibios}, since it is supported both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.
+As a specific RTOS, we chose ChibiOS \cite{chibios}, since it is supported on both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.
+
+\subsection{Graphical User Interface} % Tjark
+
+The GUI included in the telemetry system serves a proof-of-concept purpose.
+Therefore, scala-swing \cite{scala-swing} was chosen as a framework, due to the effectiveness of the scala \cite{scala} language and the high level of abstraction used when describing a user interface in scala-swing.
+Furthermore, a GUI implementation in scala can run on any system with a java virtual machine and is therefore cross-platform compatible.
+
+In order to send and receive data from the SVU via serial connection, the jSerialComm \cite{jSerialComm} library was used.
+
+\subsection{Data Processing} % Steffan
+Due to the wishes of DTU ROAST to use the telemetry data in applications of predictive modelling and performance optimization we decided to implement a simple UDP network stream hosted on the Support Vehicle laptop.
+This would allow all devices on the local network to monitor the data being sent over the serial connection for use in further data processing.
+For a proof-of-concept case we created a simple Matlab script to sample CAN messages from the UDP port and print them formatted to the terminal.
 
 \subsection{Other Libraries}
+In order to interface with the components presented in the previous section, a set of libraries were included in the project.
+These include a CAN network driver to support use of the Teensy 3.6's integrated CAN controller \cite{ACAN}, an RF OSI layer providing seamless integration with the nRF24L01+ modules, and an SD card driver to control data logging in the black box memory.
 
-In order to interface with the components presented in the previous section, a set of libraries were used.
+%In order to interface with the components presented in the previous section, a set of libraries were used.
 
-The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards.
+%The ACAN library by Pierre Molinaro is a CAN network driver for the Teensy 3.6 board, as well as older Teensy series 3 boards.
 The library has been important for the connection between the CAN network and the ports on the Teensy board.
 We have only used the classic CAN configuration, but in addition to this, the library supports Flex CAN and CANFD.
-The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s\cite{ACAN}.
-
-% ACAN.h
-% RF24.h
-%
+The driver supports bit rates ranging from 62.5 kbit/s up to 1 Mbit/s.
 
@@
 
-\todo{make clearer distinction between message types and commands. Several commands can share the same message type}
+The system is built around \textit{commands} and \textit{message types}.
+All message types begin with a command byte.
+The remaining 15 bytes are interpreted based on the message type.
+As an example, a \texttt{CanTelemetryMsg} message type is shown in figure \ref{fig:messageTypes}.
 
-The first byte, the command, always identifies the message type and indicates how the remaining 15 bytes should be interpreted.
+One message type can be connected to different commands.
+For instance, the \texttt{CanTelemetryMsg} message type can be used to send data from the SCU to the SVU as part of the data stream setting the command to \texttt{RECEIVED\_CAN} = \texttt{0x00} or it can be used to inject a CAN message from the SVU into the solar car's CAN bus setting the command to \texttt{BROADCAST\_CAN} = \texttt{0x01}.
+
+When decoding messages, only the first byte needs to be considered and it will unambiguously decide how the rest of the message should be interpreted.
 This eases the decoding of messages on the receiving end and makes for a expandable framework which has enough room for future expansions.
 
-As an example of a specific message type, figure \ref{fig:messageTypes} shows a message at the bottom with the command \texttt{BROADCAST\_CAN} which is encoded by \texttt{0x01}.
-This message carries a CAN message as a payload and commands the solar car subsystem to inject the message into the solar car CAN bus.
 Even though some message types are only sent one direction between the telemetry modules, a unified design was chosen due to the inherent simplicity.
@@
 The encryption is based on the RSA algorithm, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
-The pairings of prime numbers suited for the encryption protocol can be seen below in Table \ref{fig:prime_tabel}.
-When a message is encrypted, each of the 16 bytes is extended to a 16-bit unsigned integer to avoid overflow problems.
-Once transmitted to the receiving node, the message is then decrypted and converted back to unsigned 8-bit type.
+The pairings of prime numbers suited for the encryption protocol can be seen in Fig.
+\ref{fig:prime_tabel}.
+
+When a message is encrypted, each of the 16 bytes is extended to a 16-bit unsigned integer to avoid overflow problems increasing the total message size to 32 bytes.
+Once transmitted to the receiving node, the message is then decrypted and the overflow protection bytes can be removed leaving us with 16 bytes again.
 This essentially doubles the size of the data payloads being sent in our RF transmission, which is the downside of our encryption solution.
@@
 
-% message protcol
-% command system
-% security -> encryption
+\section{\todo{Final product}}
+\subsection{Solar Car Unit}
 
-\section{\todo{Final product}}
-\subsection{Solar Car}
+% actual circuit
+% - powered through molex connector 12V
+% - converted to 5V and teensy internal voltage regulator to 3.3
+% - Teensy has RTC which can keep time when powered by an external 2.9V battery
+
+% on the software side
+% - threads:
+%  -
 
@@
 % threads 
-\subsection{Support Vehicle}
+\subsection{Support Vehicle Unit}
 \begin{figure}
@@
 
-\section{Testing and Validation} % Victor
+\section{Verification and Validation} % Victor
 As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design.
@@
 
-We performed software validation as we went along with the software coding, to make sure that the actual behaviour of the code was as the expected behaviour.
+We performed software validation concurrently with writing our code to make sure that the actual behaviour of the program matched the expected behaviour.
 This was done through the testing environment in PlatformIO.
@@
 Doing this provided us a test bench for our project, which greatly improved the credibility of our work, proving that our design works in general cases and not just the case we represent in our report.
-
-%Finally, we were able to test all three modules
 
@@
 % variable length messages would reduce overhead -> but most messages contain CAN and are optimized
+% UDP works only one way for now. If automated control from matlab should be possible, udp listeners should be able to send as well
 

### 2021-06-17 code/SolarVehicle/lib/rf/RFfunctions.h
new

@@ -47,7 +47,7 @@ bool RFinit()
     if(!radio.begin()) return false;                            // Begin radio class object, return false if not succeeded. 
     radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level, choose between MIN, LOW, HIGH, MAX (higher PA level improves range)
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate, choose between RF24_250KBS, RF24_1MBS, RF24_2MBS (higher Data Rates may cause data loss)
-    radio.setPayloadSize(BaseTelemetryMsg::length()<<1);        // Set constant payload size to improve transmission time
+    radio.setPayloadSize(32);                                   // Set constant payload size to improve transmission time
     radio.setAutoAck(true);                                     // RX node sends an automatic ack packet
     if (ack) radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry

### 2021-06-17 code/SolarVehicle/lib/util/Fifo.h

@@ -11,7 +11,6 @@ class Fifo {
         semaphore_t spaceSem, dataSem;
         T *fifo;
         size_t _size, _head, _tail;  
-        uint32_t size;
 
     public: Fifo(uint32_t size){
         this->_size  = size;
@@ -74,9 +73,10 @@ class Fifo {
     // Prints the contents of the fifo buffer for debugging
     public: void printContents(){
         for (int i = 0; i < _size; i++){
-            Serial.print(i + " : " + fifo[i]);
+            char str[64]; 
+            snprintf(str, sizeof(str), "%-2d : %d", i, fifo[i]); 
+            Serial.println(str);
         }
-        Serial.println();
     }
 
     // Returns a pointer to element at head index

### 2021-06-17 code/SolarVehicle/src/telemetry/support-vehicle/SystemThds.h
added SVU threads

@@ -1,5 +1,5 @@
 /**************************************************************
- * @file    threads.h                                         *
+ * @file    SystemThds.h                                      *
  * @author  Steffan Martin Kunoy                              *
  * @brief   ChibiOS threads to be run on the Support Vehicle. *
  **************************************************************/
@@ -17,21 +17,21 @@
 Security sec;                       // Security global variable
 Fifo<BaseTelemetryMsg> TXfifo(32);  // Fifo global variable
 
-// A struct used to pass arguments to transmitterThread
-struct threadBundle 
+// A struct used to pass arguments to transmitterThd
+struct thdBundle 
 {
   Security *security;             // Security class member
   Fifo<BaseTelemetryMsg> *fifo;   // Fifo class member
 };
 
-// 2048 byte working stack for receiverThread
-THD_WORKING_AREA(WaReceiverThread, 2048);
+// 2048 byte working stack for receiverThd
+THD_WORKING_AREA(WaReceiverThd, 2048);
 
 /*****************************************************
- * @brief   Thread function for receiverThread.      *
+ * @brief   Thread function for receiverThd.      *
  * @param   arg, typecast to Security class pointer. *
 ******************************************************/
-THD_FUNCTION(receiverThread, arg)
+THD_FUNCTION(receiverThd, arg)
 {
   Security *sec = (Security*)arg;                                   // Cast input argument to Security class pointer
   BaseTelemetryMsg received;                                        // BaseTelemetryMsg object for received messages
@@ -62,20 +62,20 @@ THD_FUNCTION(receiverThread, arg)
         Serial.println("Could not receive message.");                 // Message was not received 
       }
     }
-    chThdYield();                                                     // Yield for same-priority thread
+    chThdYield();                                                     // Yield for same-priority Thd
   }
 }
 
-// 2048 byte working stack for transmitterThread
-THD_WORKING_AREA(waTransmitterThread, 2048);
+// 2048 byte working stack for transmitterThd
+THD_WORKING_AREA(waTransmitterThd, 2048);
 
 /*****************************************************
- * @brief   Thread function for transmitterThread.   *
- * @param   arg, typecast to threadBundle pointer.   *
+ * @brief   Thread function for transmitterThd.   *
+ * @param   arg, typecast to thdBundle pointer.   *
 ******************************************************/
-THD_FUNCTION(transmitterThread, arg)
+THD_FUNCTION(transmitterThd, arg)
 {
[…]

### 2021-06-17 code/SolarVehicle/src/telemetry/support-vehicle/supportVehicleMain.cpp

@@ -5,7 +5,7 @@
  ****************************************************************************************************/
 
 #include <Arduino.h>
-#include "telemetry/support-vehicle/threads.h"
+#include "telemetry/support-vehicle/SystemThds.h"
 
 /*********************************************************
  * @brief Initializes all threads with a working area,   *
@@ -14,9 +14,9 @@
 void chSetup()
 {
   chSysInit();          // Initializes ChibiOS system
-  chThdCreateStatic(WaReceiverThread, sizeof(WaReceiverThread), NORMALPRIO + 1, receiverThread, &sec);
-  threadBundle transmitterBundle = {.security = &sec, .fifo = &TXfifo}; 
-  chThdCreateStatic(waTransmitterThread, sizeof(waTransmitterThread), NORMALPRIO + 1, transmitterThread, &transmitterBundle);
+  chThdCreateStatic(WaReceiverThd, sizeof(WaReceiverThd), NORMALPRIO + 1, receiverThd, &sec);
+  thdBundle transmitterBundle = {.security = &sec, .fifo = &TXfifo}; 
+  chThdCreateStatic(waTransmitterThd, sizeof(waTransmitterThd), NORMALPRIO + 1, transmitterThd, &transmitterBundle);
 }
 
 void setup()

### 2021-06-17 documentation/FinalPaper/paper.tex
Updates from Overleaf

@@
 \usepackage{textcomp}
+\usepackage{siunitx}
 \usepackage{xcolor}
@@
 The development of the module is done through design sprints, where hardware and software is gradually developed along side each other.
-First a Hardware prototype on breadboards, then the software to run the modules satisfactorily was developed and finally the final product was made when the hardware was soldered to perf boards.\\
+First a Hardware prototype on breadboards, then the software to run the modules satisfactorily was developed and finally the final product was made when the hardware was soldered to perf boards.
+
 The software developed for the module provided the module to have two-way communication with CAN message structure, encryption of the data during transmission between the module in the solar car and a module in a support car, the ability to give the solar car module commands from the support car and streaming received messages in the support car to Matlab.
+
+The final product supports the features above and a range test of the solar car module and support car module had a succesful and reliable transmission distance up to about 200m.
 \end{abstract}
@@
 \IEEEPARstart{I}{n} 2023, the DTU Roadrunners Solar Team (abbreviated ROAST) is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in the Northern Territory to Adelaide in South Australia.
-The solar car has limited opportunities for recharging during the race, so its the main energy source will be the sun.
+The solar car has limited opportunities for recharging during the race, so its main energy source will be the sun.
 To this end, solar panels will cover the car, however, the solar car is only allowed to have a maximum of 4 square meters of solar panels.
@@
 
-Throughout the race the solar car will be monitored by a support vehicle, which can monitor the car's condition and should be able to issue commands to the solar cars internal systems.
+Throughout the race the solar car will be monitored by a support vehicle, which can monitor the car's condition and should be able to issue commands to the solar car's internal systems.
 This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
@@
 After this comes a paragraph discussing the implementation of the communication between the two modules, both internal and during transmission.
-This concludes with a section of the final product, which is followed up by our method method for testing and validating our solution, as well as discussion and conclusion of the solution.
+The paper finishes with a section of the final product, which is followed up by our method for testing and validating our solution, as well as a discussion and conclusion of the solution.
 
@@
 This will ease the burden on the driver and leverage the expertise of the supporting crew.
-For these reasons we have chosen to focus our project on the following points:
+For these reasons we have chosen to focus our project on the following guiding questions:
 \begin{itemize}
@@
     \item How can CAN data be transmitted securely over a distance of up to one kilometer?
-    \item How does the support-car module process data upon receiving it from the solar car?
+    \item How does the unit sitting in the support vehicle process data upon receiving it from the solar car?
     \item What is the best method for storing data locally (black box) while being accessible at a later time?
@@
 \section{Methodology}
-The project methodology has largely consisted of making a preliminary design to be completed in a series of design sprints.
+The project was executed under a sprint methodology where a series of design and implementation sprints were conducted, each resulting in an increasingly finished product.
 At first a detailed specification was drafted which then served as a reevaluation tool and an overall guidance each time a sprint was completed and when the goal for the next sprint had to be selected.
@@
 We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU, respectively.
-The Teensy line-up has been used as the main microcontroller platform in most ecocar projects and we decided to use them as well after considering the performance of newer Teensy boards.
+The Teensy line-up has been used as the main microcontroller platform in most DTU Roadrunners projects and we decided to use them as well after considering the performance of newer Teensy boards.
 The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data.
@@
 In order to interface with the physical CAN bus though, a transceiver is needed.
-Here the MCP2551 was chosen since it has been used extensively in other ecocar projects.
+Here the MCP2551 was chosen since it has been used extensively in other DTU Roadrunners projects.
 The component maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.
@@
 We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accesibility through the Teensy 3.6 integrated SD card slot.
-Furthermore, this solution scales very well and easily in terms storage capacity, with new SD card sizes exceeding 1 TiB.
+Furthermore, this solution scales very well and easily in terms of storage capacity, with new SD card sizes exceeding 1 TiB.
 
@@
 It comes with a VScode plugin for ideal IDE support.
-Other project groups working on the solar car in parallel integrated their software into the PlatformIO ecosystem as well, making extensive code reuse in the future possible.
+Other project groups working on the solar car in parallel integrated their software into the PlatformIO ecosystem as well, easing extensive code reuse in the future.
 
@@
 We chose to build the software for the telemetry system on top of a real time operating system.
-The small kernel running behind the scenes when using a RTOS allows a step up in abstraction from bare metal programming.
+The small kernel running on the microcontroller behind the scenes when using a RTOS allows a step up in abstraction from bare metal programming.
 Different tasks can be spawned as threads running concurrently.
@@
 \subsection{Data Processing} % Steffan
-Due to the wishes of DTU ROAST to use the telemetry data in applications of predictive modelling and performance optimization we decided to implement a simple UDP network stream hosted on the Support Vehicle laptop.
+Due to the wishes of DTU ROAST to use the telemetry data in applications of predictive modelling and performance optimization, we needed a way to retain the data being sent on the CAN bus, both locally via logging and remotely in real time via streaming.
+
+We log all the messages received on the CAN bus to the external "black box" memory as \texttt{.csv} files.
+Each line in the file corresponds to a message with date and time stamps, ID, RTR, length and 64-bit data fields.
+Log file switching was also implemented.
+
+For remote access we decided to implement a simple UDP network stream hosted on the Support Vehicle laptop.
 This would allow all devices on the local network to monitor the data being sent over the serial connection for use in further data processing.
@@
 In order to interface with the components presented in the previous section, a set of libraries were included in the project.
-These include a CAN network driver to support use of the Teensy 3.6's integrated CAN controller \cite{ACAN}, an RF OSI layer providing seamless integration with the nRF24L01+ modules, and an SD card driver to control data logging in the black box memory.
+These include a CAN network driver to support use of the Teensy 3.6's integrated CAN controller \cite{ACAN}, an RF OSI layer providing seamless integration with the nRF24L01+ modules \cite{nrfLib}, and an SD card driver \cite{sdfatlib} to control data logging in the black box memory.
 
@@
 For instance, the \texttt{CanTelemetryMsg} message type can be used to send data from the SCU to the SVU as part of the data stream setting the command to \texttt{RECEIVED\_CAN} = \texttt{0x00} or it can be used to inject a CAN message from the SVU into the solar car's CAN bus setting the command to \texttt{BROADCAST\_CAN} = \texttt{0x01}.
+Even though such message type and command combinations are only sent one direction between the telemetry modules, a unified message system for both directions was chosen due to the inherent simplicity.
 
@@
 This eases the decoding of messages on the receiving end and makes for a expandable framework which has enough room for future expansions.
-
-Even though some message types are only sent one direction between the telemetry modules, a unified design was chosen due to the inherent simplicity.
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/SCU_CircuitDiagram.pdf}
+    \caption{The circuit diagram for the SCU.}
+    \label{fig:SCU_circuit}
+\end{figure}
+
+The SCU is built around the Teensy 3.6 and includes the before mentioned CAN transceiver and RF module.
+In the solar car, all modules receive power through a molex connector at $\SI{12}{V}$.
+The molex connector also includes the differential pair of wires of the CAN bus.
+
+The Teensy requires a voltage between $\SI{3.6}{V}$ and $\SI{6}{V}$ to be powered, while the CAN transceiver requires $\SI{5}{V}$ and the RF module requires $\SI{3.3}{V}$.
+Therefore, a linear voltage regulator is chosen to step down the external voltage from $\SI{12}{V}$ to $\SI{5}{V}$ to power the Teensy and the CAN transceiver.
+The Teensy 3.6 has itself an integrated voltage regulator stepping down the input voltage to $\SI{3.3}{V}$ which we use to power the RF module.
+
+The Teensy 3.6 includes a real time clock (RTC) which can keep the time while the Teensy is powered off by connecting a coin cell battery to the Teensy's $V_{bat}$ terminal.
+This allows the generated time stamps associated with a received CAN message to always be valid, once the RTC is synchronized.
+
+A complete circuit diagram of the SCU is shown in figure \ref{fig:SCU_circuit}.
+The final circuit contains decoupling capacitors for the Teensy as well as the RF module and an optional CAN bus termination which can be seen to the right of the diagram.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/SCU_threads.pdf}
+    \caption{The execution flow of the 4 threads running in the SCU.}
+    \label{fig:SCU_threads}
+\end{figure}
+
+On the software side, the SCU is executing four threads to handle its tasks. A diagram of the execution flow of the threads is shown in figure \ref{fig:SCU_threads}.
+
+There are two threads running at the lowest priority.
+Since the right to execute is based on priority, the two threads need to use a technique called \textit{cooperative multithreading} to solve the problem of arbitrating execution slots between them.
+Cooperative multithreading is based on the two threads voluntarily giving up their execution right in reasonable time intervals therby giving the other thread at the same priority the right to execute.
+This execution flow is coordinated by calling the \texttt{yield} method, which invokes the real time kernel which then switches threads.
+
 % actual circuit
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{documentation/images/SCU_CircuitDiagram.pdf}
-    \caption{The circuit diagram for the SCU.}
-    \label{fig:SCU_circuit}
-\end{figure}
 % threads 

### 2021-06-17 documentation/FinalPaper/paper.tex

@@
 
+The system thread and the CAN receiver thread having the lowest priority always run when no other high priority thread wants to execute.
+In the system thread, messages are received via RF, decoded and the action connected to the message is executed.
+This includes for instance injecting a message on the CAN bus.
+The CAN receiver thread tries to receive new messages from the CAN bus.
+These messages then need to be passed to the two high priority threads, which process them.
+
+This inter-thread communication is performe
+
 % actual circuit
@@
     \caption{The circuit diagram for the SVU.}
-    \label{fig:SCU_circuit}
+    \label{fig:SVU_circuit}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[height=4.5cm]{documentation/images/SVU_threads.pdf}
+    \caption{The execution flow of the 2 threads running in the SVU.}
+    \label{fig:SVU_threads}
 \end{figure}

### 2021-06-17 documentation/FinalPaper/paper.tex

@@
 First a Hardware prototype on breadboards, then the software to run the modules satisfactorily was developed and finally the final product was made when the hardware was soldered to perf boards.
+The hardware used is Teensy boards, nRF24 antennas and
 
@@
 
-The final product supports the features above and a range test of the solar car module and support car module had a succesful and reliable transmission distance up to about 200m.
+The final product supports the features above and a range test of the solar car module and support car module had a successful and reliable transmission distance up to about 200m.
 \end{abstract}
@@
 \section{Methodology}
-The project was executed under a sprint methodology where a series of design and implementation sprints were conducted, each resulting in an increasingly finished product.
-At first a detailed specification was drafted which then served as a reevaluation tool and an overall guidance each time a sprint was completed and when the goal for the next sprint had to be selected.
+The project was executed under a series of design and implementation sprints, each resulting in an increasingly finished product.
+At first a detailed specification was drafted which then served as a reevaluation tool for future sprint procedures.
 Using the specification, we made a prototype setup of the hardware on breadboards as our first sprint.
@@
 
-During the development of the software functions, we also verified and validated our code design.
+Subsequently, we created our first proof of concept (POC) model with the hardware setup on two breadboards and basic software functions.
+Our aim was to test the basic functionality of the system, such as receiving CAN messages and sending data over the RF channels.
+
+During the following design sprints we focused mostly on software development objectives, such as multithreading and data handling.
+developed a minimum viable product (MVP) with support for all the core
+
+During the development of the system software, we also verified and validated our code design.
 In verification the code and its functionality was discussed.
@@
 Cooperative multithreading is based on the two threads voluntarily giving up their execution right in reasonable time intervals therby giving the other thread at the same priority the right to execute.
-This execution flow is coordinated by calling the \texttt{yield} method, which invokes the real time kernel which then switches threads.
+This execution flow is coordinated by calling the \texttt{yield} method, which invokes the real time kernel which then switches threads \cite{chibiOsYield}.
 
@@
 
-This inter-thread communication is performe
+This inter-thread communication is performed using two First-In-First-Out (FIFO) buffers, one for each consumer thread, which decouple the producer and the consumer.
+The two high priority threads only need to execute when new data is written into the buffer.
+To achieve this, a counting semaphore is used.
+This construct allows one thread to hand out execution rights by incrementing the semaphore counter.
+A thread waiting for this execution right can then execute if its priority allows it to \cite{chibiOsSemaphore}.
+
+This technique is used to let the producer, the can receiver thread, signal to the consumer, the black box thread or the RF thread, that new data is available in the FIFO.
+The real time kernel then gives the execution right to the consumer threads due to their higher priority.
+
+The black box thread takes a CAN message, converts it into a csv-file line and writes it into a buffer, which is emptied at a certain level to the actual SD card.
+When a predefined number of CAN messages have been logged to one file, a new file is opened.
+
+The RF thread encrypts the CAN messages and tries to send them to the SVU. If unsuccessful this operation has no effect
 

### 2021-06-18 code/SolarVehicle/src/telemetry/solar-car/RfTxThd.h
Fix fifo semaphore problems

@@ -22,7 +22,7 @@
  * A bundle used for passsing all relevant resources to the radio thread
  */
 struct rfTxWorkerBundle{
-    MultiReaderFifo<CanTelemetryMsg> *fifo;
+    Fifo<CanTelemetryMsg> *fifo;
     ThreadState *state;
     Security *sec;
 };
@@ -34,20 +34,18 @@ THD_FUNCTION(rfWorker, arg){
 
   rfTxWorkerBundle *bundle = (rfTxWorkerBundle*) arg;
   ThreadState *state = bundle->state;
-  MultiReaderFifo<CanTelemetryMsg> *fifo = bundle->fifo;
+  Fifo<CanTelemetryMsg> *fifo = bundle->fifo;
   Security *sec = bundle->sec;
 
   CanTelemetryMsg *msg;
 
 
-  uint8_t readerId = 1;
-
   WITH_MTX(serialMtx){Serial.println("RfTxThd:\tStarting");}
 
   while(!state->terminate){
-   
+
     //Serial.println("RFTxThd:\tWaiting for data");
-    fifo->waitForData(readerId);
+    fifo->waitForData();
 
     if(state->pause){
       state->suspend();
@@ -55,45 +53,45 @@ THD_FUNCTION(rfWorker, arg){
 
     MEASURE("RfTxThd:\t"){
 
-      msg = fifo->head(readerId);
+      msg = fifo->head();
       
       chSysLock();
       uint16_t encrypted[16];
       char tempString[128];
 
-      WITH_MTX(serialMtx){
+      /*WITH_MTX(serialMtx){
         Serial.print("RfTxThd:\t");
         msg->toMessage()->toString(tempString,sizeof(tempString));
         Serial.println(tempString);
+      }*/
+      sec->encrypt((uint8_t*)msg,encrypted,BaseTelemetryMsg::length());
 
-        sec->encrypt((uint8_t*)msg,encrypted,BaseTelemetryMsg::length());
+      /*WITH_MTX(serialMtx){
 
         Serial.print("RfTxThd:\t");
         for(uint32_t i = 0; i < BaseTelemetryMsg::length(); i++) {
           Serial.print(encrypted[i],HEX); Serial.print(" ");
         }
         Serial.println();
-
-        
-      }
+      }*/
       WITH_MTX(rfMTX){
         if(!RFtransmit(encrypted,BaseTelemetryMsg::length()<<1)){
           Serial.println("RfTxThd:\tMessage not send or incorrect ACK");
         } else {
[…]

### 2021-06-20 documentation/FinalPaper/paper.tex
Updates from Overleaf

--- 2021-06-17 documentation/FinalPaper/paper.tex
@@
-\documentclass[conference]{IEEEtran}
+\documentclass[a4paper,conference]{IEEEtran}
 \IEEEoverridecommandlockouts
@@
 \usepackage{comment}
-\usepackage{movie15}
 \usepackage{textcomp}
 \usepackage{siunitx}
+\usepackage{listings}
 \usepackage{xcolor}
@@
 \usepackage{url}
+\usepackage{fancyhdr}
 \usepackage[linkcolor = blue]{hyperref}
 \newcommand{\MYhref}[3][blue]{\href{#2}{\color{#1}{#3}}}
-
-% correct bad hyphenation here
+\textheight = 250mm
+% correct bad hyphenation herehttps://www.overleaf.com/project/602635b948e4260c4d50d582
 \hyphenation{op-tical net-works semi-conduc-tor}
-
 \newcommand{\todo}[1]{{\color{olive} TODO: #1}}
-
+\pagestyle{fancy}
+\fancypagestyle{default}{
+    \fancyhf{}
+    \renewcommand{\headrulewidth}{0pt}
+    \renewcommand{\footrulewidth}{0pt}
+    \rfoot{7.\thepage}
+}
 % Regler til skrivning: 
@@
 % 
-
+\pagestyle{fancy}
 \begin{document}
-
-\title{Telemetry Module for Solar Vehicle}
+\pagestyle{default}
+\title{Telemetry Module for the DTU Roadrunners Solar Car}
 
@@
 \IEEEauthorblockA{\textit{Department of Electrical Engineering} \\
-\textit{Technical Uniservity of Denmark}\\
+\textit{Technical University of Denmark}\\
 Kgs. Lyngby, Denmark \\
@@
 \begin{abstract}
-This paper presents the telemetry project for the DTU ROAST Solar car.
-The development of the module is done through design sprints, where hardware and software is gradually developed along side each other.
-First a Hardware prototype on breadboards, then the software to run the modules satisfactorily was developed and finally the final product was made when the hardware was soldered to perf boards.
+This paper presents the telemetry project for the DTU ROAST solar car.
+The software and hardware developed for the telemetry system provides a two-way communication between the solar car and a support vehicle.
+The solar car module can read data from the CAN bus, store it locally in a black box and stream it to the support vehicle, and in return the support vehicle can send commands to the solar car.
+All messages are secured with an RSA encryption.
+A graphical user interface enables the support vehicle to stream the CAN data and send commands.
+In addition, support for a simple UDP network stream was also implemented which can for instance be accessed in Matlab.
 
-The software developed for the module provided the module to have two-way communication with CAN message structure, encryption of the data during transmission between the module in the solar car and a module in a support car, the ability to give the solar car module commands from the support car and streaming received messages in the support car to Matlab.
-
-The final product supports the features above and a range test of the solar car module and support car module had a succesful and reliable transmission distance up to about 200m.
+A provisional hardware setup of the solar car and support vehicle modules was implemented using two Teensy microcontrollers on perfboards.
+A range test of the two modules had a successful and reliable transmission distance of up to around 160 m.
+While this does not satisfy the original goal of 400 m - 1000 m, the final solution is concluded to have a good communication chain design between the solar car and support vehicle modules, and can be readily improved upon to support a longer transmission distance.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-ROAST, telemetry, CAN, RSA encryption, Matlab
+Telemetry, CAN, RSA encryption, Matlab
 \end{IEEEkeywords}
@@
 
-\IEEEPARstart{I}{n} 2023, the DTU Roadrunners Solar Team (abbreviated ROAST) is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in the Northern Territory to Adelaide in South Australia.
-The solar car has limited opportunities for recharging during the race, so its main energy source will be the sun.
+\IEEEPARstart{I}{n} 2023, the DTU Roadrunners Solar Team (abbreviated ROAST) is set to participate in the Bridgestone World Solar Challenge, a race spanning over 3000 km across Australia from Darwin in the Northern Territory to Adelaide in South Australia \cite{wsc}.
+%The solar car has limited opportunities for recharging during the race, so its main energy source will be the sun.
 To this end, solar panels will cover the car, however, the solar car is only allowed to have a maximum of 4 square meters of solar panels.
 This emphasizes the need to create an energy-efficient solar car to win the race \cite{wsc}.
+During the race the car will be mainly powered by its solar panels which emphasizes the need to create an energy-efficient car.
 
-Throughout the race the solar car will be monitored by a support vehicle, which can monitor the car's condition and should be able to issue commands to the solar car's internal systems.
-This is a crucial function as the solar car will be navigating Australia's busy highways, where the heat and tire pressure may cause a detriment to the vehicle.
-The support vehicle will therefore be able to analyze and react to the data stream being transmitted from the solar car, even when the driver is preoccupied with driving the car.
+Throughout the race the solar car will be accompanied by a support vehicle, which can monitor the car's condition and should be able to issue commands to the solar car's internal systems.
+This is a crucial function as the solar car will be navigating Australia's busy highways, where for instance the heat and tire pressure may cause a detriment to the vehicle.
+The support vehicle will therefore be able to analyze and react to the data stream being transmitted from the solar car, even when the driver is preoccupied driving the car.
 
 Additionally, valuable data can be collected during the race which can help to identify possible areas of improvement of the solar car.
-Therefore a "black box" system is needed to capture all system activity in a local storage.
+Therefore a "black box" system is needed to capture all system activity on a local storage.
 
-These key functionalities require a system which is able to read and write data from the solar cars on-board CAN (Controller Area Network) bus and transmit it via an RF-transceiver to the support vehicle, while simultaneously logging data locally.
+These key functionalities require a system which is able to read and write data from the solar car's on-board CAN (Controller Area Network) bus and transmit it via a radio frequency (RF) transceiver to the support vehicle, while simultaneously logging data locally.
 The support vehicle is equipped with a similar module such that the support crew can react to the data manually or automatically by sending commands back to the CAN bus via RF.
@@
 
-The telemetry system must both serve the purpose of fulfilling the necessary specifications for communication, while also being as energy-efficient as possible.
+%The telemetry system must fulfill the specifications for communication, while also being as energy-efficient as possible.
 This project aims at implementing a functioning prototype for a telemetry system while mainly focusing on the former of the two objectives, that is, ensuring that the system meets the specifications with a solid solution.
 
-This paper begins with giving a more precise problem statement and introducing the methodology used throughout the project.
+This project aims at implementing a functioning prototype for a telemetry system while ensuring that the system meets the specifications with a solid solution.
+
+%This paper begins with giving a more precise problem statement and introducing the methodology used throughout the project.
 Thereafter, the overall system design is presented and the following paragraphs discuss the hardware components and software used for these.
@@
 \subsection{Problem Statement}
-Remote sensing of data from the solar car will give the ROAST team a competitive advantage as the support vehicle will be able to take on a more active role in managing and optimising the car's performance.
+Remote sensing of data from the solar car will give the ROAST team a competitive advantage as the support vehicle will be able to take on a more active role in managing and optimizing the car's performance.
 This will ease the burden on the driver and leverage the expertise of the supporting crew.
@@
 \section{Methodology}
-The project was executed under a sprint methodology where a series of design and implementation sprints were conducted, each resulting in an increasingly finished product.
-At first a detailed specification was drafted which then served as a reevaluation tool and an overall guidance each time a sprint was completed and when the goal for the next sprint had to be selected.
-Using the specification, we made a prototype setup of the hardware on breadboards as our first sprint.
+The project was executed under a series of design and implementation sprints, each resulting in an increasingly finished product.
+At first a detailed specification was drafted which then served as a reevaluation tool for future sprint procedures.
+%Using the specification, we made a prototype setup of the hardware on breadboards as our first sprint.
 The following sprints would then mostly have software development objectives, such as achieving a first transmission and receiving of data.
-Finally, a last hardware sprint was conducted where we soldered our hardware to a perf board.
+Finally, a last hardware sprint was conducted where we soldered our hardware to a perfboard.
+Subsequently, we created our first proof of concept (POC) model with the hardware setup on two breadboards.
+Our aim was to test the basic functionality of the system, such as receiving CAN messages and sending data over the RF channels.
 
-During the development of the software functions, we also verified and validated our code design.
-In verification the code and its functionality was discussed.
-To validate our code we made unit tests that test the actual behaviour of our code against the intended behaviour.
+During the following design sprints we focused mostly on software development objectives until we had achieved a minimum viable product (MVP).
+At this point our system worked as intended, albeit with reliability issues and software bugs.
+Hence, the code was validated thoroughly with unit tests that determine the behavior of our code against its intended behavior.
+The final hardware sprint resulted in soldering of our entire setup to a perfboard for better reliability.
+Real-life tests were then conducted to prove the efficacy of our product.
+Upon applying some final touches, the final product was completed and handed over to DTU ROAST.
 
-During the software development phase, we attempted to make our debugging more efficient with the use of a debugger.
+%During the software development phase, we attempted to make our debugging more efficient with the use of a debugger.
 However, this proved quite troublesome, as the Teensy 3.6 board does not have an on-board debugger and in general is not very suited for debuggers.
@@
 \section{Design}
-The overall system consists of three separate modules; the Solar Car Unit (SCU) which is connected to the CAN bus and communicates with the support vehicle unit via a wireless RF connection.
-The Support Vehicle Unit (SVU) in turn interacts with the Support Vehicle laptop through a serial connection.
-The GUI program running on the laptop broadcasts the CAN bus message stream on the local network.
+The overall system consists of three separate modules.
+The Solar Car Unit (SCU) is connected to the CAN bus and communicates with the Support Vehicle Unit (SVU) via a wireless RF connection.
+The SVU in turn interacts with a laptop in the Support Vehicle through a serial connection.
+The TelemetryUI program running on the laptop broadcasts the CAN bus message stream on the local network.
 A block diagram for the system is outlined in Fig.
 \ref{fig:schematic}.
-\begin{figure}[H]
+\begin{figure}
     \centering
-    \includegraphics[width=0.5\textwidth]{documentation/images/SystemSchematic.pdf}
+    \includegraphics[width=\linewidth]{documentation/images/SystemSchematic.pdf}
     \caption{Overview of the system's components in block diagram form.}
@@
 \subsection{Microcontrollers} % Steffan
-We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU, respectively.
+We use Teensy 3.6 and 4.0 microcontrollers as connection hubs in the SCU and SVU respectively.
 The Teensy line-up has been used as the main microcontroller platform in most DTU Roadrunners projects and we decided to use them as well after considering the performance of newer Teensy boards.
-The boards can interact with the various components through an SPI interface to gather, process and transmit large amounts of data.
 Being based on the Arduino framework, both Teensys offer support for numerous protocols and software libraries to ease the implementation of our design.
@@
 \subsection{CAN bus}
-The CAN network in the SCU enables all the car's processing nodes to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.
+The CAN network in the SCU enables all the car's engine control units (ECU) to send and receive data without a host computer. This is possible thanks to the special structure of the CAN message.
 
-\begin{figure}[h]
+\begin{figure}
     \centering
-    \includegraphics[scale=0.35]{documentation/images/detailed-can-data-frame-architecture.jpg}
-    \caption{Structure of a CAN frame.}
+    \includegraphics[width=\linewidth]{documentation/images/detailed-can-data-frame-architecture.jpg}
+    \caption{Structure of a CAN frame, non-extended.}
     \label{fig:CANframe}
@@
 
-The frame uses an identifier to tell which messages goes to which unit.
+Fig.
+\ref{fig:CANframe} shows a detailed structure of the CAN frame.
+The fields of importance for our project is the message identifier, RTR, control field and data field.
+The rest of the fields serve as error checking and coordination of communication on the bus.
+
+Each ECU has a unique ID, which the message identifier uses to determine where the CAN message should go to.
 When two messages try to access the bus, the one with the highest priority (lowest ID) will be granted access to the bus, while the other message have to wait.
-After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a node.
-The control field, sometimes called DLC, holds the amount of bytes used in the following frame field containing the actual data.
-The rest of the fields in fig.
-\ref{fig:CANframe} serve error checking and coordination of communication on the bus.
+After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a ECU.
+The control field, in other literature referred to as "Data Length Code" (DLC), holds the amount of bytes used in the following frame field containing the actual data\cite{DS_ISO}\cite{css}.
 
-The physical CAN bus itself consists of a differential pair of twisted wires, CAN High and CAN Low.
+%The physical CAN bus itself consists of a differential pair of twisted wires, CAN High and CAN Low.
 These control the bits in the CAN message by setting a dominant and recessive voltage, which corresponds to a 0 and 1 respectively.
@@
 
-The Teensy 3.6 used in the solar car features a CAN controller handling message assembly and receiving from the CAN bus.
-In order to interface with the physical CAN bus though, a transceiver is needed.
+The Teensy 3.6 used in the solar car features a CAN controller receiving data from the CAN bus as well as handling message assembly.
+However, in order to interface with the physical CAN bus, a transceiver is needed.
 Here the MCP2551 was chosen since it has been used extensively in other DTU Roadrunners projects.
-The component maintains one input and one output serial stream to the CAN controller on the Teensy and drives the CAN bus when a message is transmitted.
+The component maintains one input and one output serial stream to the CAN controller and drives the CAN bus when a message is transmitted.
 
 \subsection{RF network} % Steffan
-Two radio-frequency (RF) transceivers are required to provide a stable connection between the SCU and SVU at longer distances.
+Two radio-frequency (RF) transceivers are required to provide a wireless connection between the SCU and SVU.
 There are many options for radio modules supporting long range communication at the expense of achievable data rate.
@@
 \ref{fig:shockburst}.
-Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and CRC segments are generated automatically and therefore fall beyond the scope of this project.
+Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and CRC segments are generated automatically by the RF module.
 
-\begin{figure}[h]
+\begin{figure}
     \centering
-    \includegraphics[width=0.5\textwidth]{documentation/images/EnhancedShockburst.pdf}
+    \includegraphics[width=\linewidth]{documentation/images/EnhancedShockburst.pdf}
     \caption{The Enhanced Shockburst$^\copyright$ packet structure.}
@@
 
-We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accesibility through the Teensy 3.6 integrated SD card slot.
-Furthermore, this solution scales very well and easily in terms of storage capacity, with new SD card sizes exceeding 1 TiB.
+We chose a SD card as a permanent local storage solution for the SCU's "black box" due to its ease of use and accessibility through the Teensy 3.6 integrated SD card slot.
+Furthermore, this solution scales easily in terms of storage capacity, with new SD card sizes exceeding 1 TB.
 
@@
 
-Several of the key functionalities of the telemetry module are implemented in software.
+Several key functionalities of the telemetry module are implemented in software.
 This, together with the aim of designing a system to be expanded upon in the future, makes the choice of software frameworks very important.
@@
 Different tasks can be spawned as threads running concurrently.
-These threads get delegated time slices to run on the processor by the real time kernel based on their priority.
+These threads get delegated time slots to run on the processor by the real time kernel based on their priority.
 
-This comes at the cost of introducing overhead when switching between threads, but the great flexibility in code organization following with an RTOS was evaluated to make the overhead a worthwhile trade-off.
+This comes at the cost of introducing overhead when switching between threads, but the great flexibility in code organization possible through the use of an RTOS was determined to make the overhead a worthwhile trade-off.
 
-As a specific RTOS, we chose ChibiOS \cite{chibios}, since it is supported on both the Teensy 3.6 and 4.0 and since it attempts to keep the memory footprint of the real time kernel as small as possible.
+As a specific RTOS, we chose ChibiOS \cite{chibios}, since it is supported on both the Teensy 3.6 and 4.0 and because it attempts to keep the memory footprint of the real time kernel as small as possible.
 
@@
 
-The GUI included in the telemetry system serves a proof-of-concept purpose.
+The graphical user interface (GUI) included in the telemetry system serves a proof-of-concept purpose.
 Therefore, scala-swing \cite{scala-swing} was chosen as a framework, due to the effectiveness of the scala \cite{scala} language and the high level of abstraction used when describing a user interface in scala-swing.
 Furthermore, a GUI implementation in scala can run on any system with a java virtual machine and is therefore cross-platform compatible.
-
-In order to send and receive data from the SVU via serial connection, the jSerialComm \cite{jSerialComm} library was used.
 
@@
 We log all the messages received on the CAN bus to the external "black box" memory as \texttt{.csv} files.
-Each line in the file corresponds to a message with date and time stamps, ID, RTR, length and 64-bit data fields.
-Log file switching was also implemented.
+Each line in the file corresponds to a message with date and time stamps, ID, RTR, DLC and 64-bit data fields as displayed in Listing \ref{lst:logFile}.
 
-For remote access we decided to implement a simple UDP network stream hosted on the Support Vehicle laptop.
-This would allow all devices on the local network to monitor the data being sent over the serial connection for use in further data processing.
-For a proof-of-concept case we created a simple Matlab script to sample CAN messages from the UDP port and print them formatted to the terminal.
+\lstset{
+  basicstyle=\fontsize{9}{13}\selectfont\ttfamily
+}
+\begin{lstlisting}[caption={Example of a \texttt{.csv} log file with random data.},captionpos=b,label={lst:logFile}]
+"time","id","rtr","len","data"
+"18/06/2021 07-56-23",851,0,4,4266193389
+"18/06/2021 07-56-24",55,0,6,24023809739244
+"18/06/2021 07-56-24",205,0,4,3791082139
+"18/06/2021 07-56-25",4,0,6,251964400399878
+"18/06/2021 07-56-26",1279,0,5,65365801097
+\end{lstlisting}
+
+For remote access to the data stream we decided to implement a simple UDP network stream being hosted by the TelemetryUI.
+This allows all devices on the local network to monitor the data being sent over the serial connection for use in further data processing.
+For a POC case we created a simple Matlab script to sample CAN messages from the UDP port and print them formatted to the terminal.
 
@@
 In order to interface with the components presented in the previous section, a set of libraries were included in the project.
-These include a CAN network driver to support use of the Teensy 3.6's integrated CAN controller \cite{ACAN}, an RF OSI layer providing seamless integration with the nRF24L01+ modules \cite{nrfLib}, and an SD card driver \cite{sdfatlib} to control data logging in the black box memory.
+These include a CAN network driver to support use of the Teensy 3.6's integrated CAN controller \cite{ACAN}, an RF OSI layer providing seamless integration with the nRF24L01+ modules \cite{nrfLib}, and an SD card driver to control data logging in the black box memory \cite{sdfatlib}.
 
@@
 Therefore, the message protocol is designed around 16 byte messages, which are large enough to comfortably fit a CAN message and a time stamp from when the message was received.
-A diagram of the byte layout of a message is shown in figure \ref{fig:messageTypes}.
+A diagram of the byte layout of a message is shown in Fig.
+\ref{fig:messageTypes}.
 
@@
 The remaining 15 bytes are interpreted based on the message type.
-As an example, a \texttt{CanTelemetryMsg} message type is shown in figure \ref{fig:messageTypes}.
+As an example, a \texttt{CanTelemetryMsg} message type is shown in Fig.
+\ref{fig:messageTypes}.
 
@@
 When decoding messages, only the first byte needs to be considered and it will unambiguously decide how the rest of the message should be interpreted.
-This eases the decoding of messages on the receiving end and makes for a expandable framework which has enough room for future expansions.
+This eases the decoding of messages on the receiving end and makes for a expandable framework.
 
 As an extra feature not included in our original specification, we have made an encryption protocol for our telemetry module.
-The encryption is based on the RSA algorithm, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
-The pairings of prime numbers suited for the encryption protocol can be seen in Fig.
-\ref{fig:prime_tabel}.
+The feature can be activated by transmitting an \texttt{ENABLE\_ENCRYPTION} command to the SCU.
+The encryption is based on the RSA algorithm \cite{rsa}, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
+The pairings of prime numbers suited for the encryption protocol can be seen in a table uploaded to DTU ROAST BitBucket \cite{prime}.
 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{documentation/images/prime_tabel.png}
-    \caption{Valid combinations of prime numbers, marked by O}
-    \label{fig:prime_tabel}
-\end{figure}
+The CAN bus used in the solar car is driven at $\SI{125}{kbit/s}$.
+With the shortest possible CAN message being $\SI{47}{bits}$ and the longest being $\SI{111}{bits}$, a theoretical throughput of $\SI{1126}{s^{-1}}$ to $\SI{2659}{s^{-1}}$ messages can be achieved at full usage of the bandwidth.% The Teensy 3.6 being clocked at $\SI{180}{MHz}$ can thus complete between $\SI{62680}{}$ and $\SI{148016}{}$ instructions between two messages assuming it can complete one instruction per clock cycle.
 
-\section{\todo{Final product}}
+The RF connection transmits Shockburst$^\copyright$ messages of worst case $\SI{392}{bits}$ length at $\SI{1}{Mbit/s}$ and is thus able to achieve a throughput of $\SI{2551}{s^{-1}}$ messages which is slightly less than the peak CAN bus throughput.
+This is unlikely to occur in reality, since all CAN messages need to have an empty data field to achieve the peak throughput.
+
+A serial connection of up to $\SI{6}{Mbit/s}$ can be achieved using the Teensy 4.0.
+When sending from the SVU to the TelemetryUI, the message types are converted into JSON strings (see Listing \ref{lst:json}), which are worst case $\SI{117}{bytes}$ long, giving a throughput of $\SI{6410}{s^{-1}}$ messages.
+The JSON string can be directly used to set the state of an object inside scala which then can be used for further processing.
+
+\lstset{
+  basicstyle=\fontsize{9}{13}\selectfont\ttfamily
+}
+\begin{lstlisting}[caption={Example of a JSON string.},captionpos=b,label={lst:json}]
+{"cmd":0,"can":{"id":1018,"rtr":false,"len":5,
+"data":831599620407,
+"stamp":{"hour":8,"minute":58,"second":27}}}
+\end{lstlisting}
+
+\section{Implementation}
 \subsection{Solar Car Unit}
@@
 
-The SCU is built around the Teensy 3.6 and includes the before mentioned CAN transceiver and RF module.
+The SCU is built around the Teensy 3.6 and includes the aforementioned CAN transceiver and RF module.
 In the solar car, all modules receive power through a molex connector at $\SI{12}{V}$.
@@
 
-The Teensy requires a voltage between $\SI{3.6}{V}$ and $\SI{6}{V}$ to be powered, while the CAN transceiver requires $\SI{5}{V}$ and the RF module requires $\SI{3.3}{V}$.
+Different voltages are required to operate the hardware setup correctly.
 Therefore, a linear voltage regulator is chosen to step down the external voltage from $\SI{12}{V}$ to $\SI{5}{V}$ to power the Teensy and the CAN transceiver.
@@
 
-The Teensy 3.6 includes a real time clock (RTC) which can keep the time while the Teensy is powered off by connecting a coin cell battery to the Teensy's $V_{bat}$ terminal.
-This allows the generated time stamps associated with a received CAN message to always be valid, once the RTC is synchronized.
+A real time clock (RTC) is included on the Teensy 3.6 which can keep the time while the Teensy is powered off by connecting a coin cell battery to the its $V_{bat}$ terminal.
+This ensures that the generated time stamps associated with a received CAN message are always valid, provided the RTC is synchronized.
 
-A complete circuit diagram of the SCU is shown in figure \ref{fig:SCU_circuit}.
-The final circuit contains decoupling capacitors for the Teensy as well as the RF module and an optional CAN bus termination which can be seen to the right of the diagram.
+A complete circuit diagram of the SCU is shown in Fig.
+\ref{fig:SCU_circuit}.
+The final circuit contains decoupling capacitors for the Teensy as well as the RF module and an optional CAN bus termination which can be activated by a switch.
 
@@
 
-On the software side, the SCU is executing four threads to handle its tasks. A diagram of the execution flow of the threads is shown in figure \ref{fig:SCU_threads}.
+On the software side, the SCU is executing four threads to handle its tasks. A diagram of the execution flow of the threads is shown in Fig. \ref{fig:SCU_threads}.
 
 There are two threads running at the lowest priority.
-Since the right to execute is based on priority, the two threads need to use a technique called \textit{cooperative multithreading} to solve the problem of arbitrating execution slots between them.
-Cooperative multithreading is based on the two threads voluntarily giving up their execution right in reasonable time intervals therby giving the other thread at the same priority the right to execute.
-This execution flow is coordinated by calling the \texttt{yield} method, which invokes the real time kernel which then switches threads.
+Since the right to execute is based on priority, these threads use a technique called \textit{cooperative multithreading} to solve the problem of arbitrating execution slots between them.
+Cooperative multithreading is based on the two threads voluntarily giving up their execution right by calling \texttt{yield} in reasonable time intervals, thereby giving the other thread at the same priority the right to execute \cite{chibiOsYield}.
 
-% actual circuit
-% - powered through molex connector 12V
-% - converted to 5V and teensy internal voltage regulator to 3.3
-% - Teensy has RTC which can keep time when powered by an external 2.9V battery
+In the system thread, messages are received via RF, decrypted, decoded and the action connected to the message is executed.
+The CAN receiver thread tries to receive new messages from the CAN bus.
+These messages then need to be passed to the two high priority threads, which process them.
 
-% on the software side
-% - threads:
-%  -
+This inter-thread communication is performed using two First-In-First-Out (FIFO) buffers, one for each consumer thread, which decouple the producer and the consumer.
+The two high priority threads only need to execute when new data is available.
+To achieve this, a counting semaphore is used which allows one thread to hand out execution rights by incrementing the semaphore counter and another thread to wait for execution rights \cite{chibiOsSemaphore}.
+This technique is used to let the producing CAN receiver thread activate the consumer threads when new data is available in the FIFO.
 
-% threads 
+The black box thread takes a CAN message, converts it into a \texttt{.csv} file line and writes it into a buffer, which is emptied to the actual SD card when filled.
+When a predefined number of CAN messages have been logged to one file, a new file is opened.
+The RF thread encrypts the CAN messages and tries to send them to the SVU.
+If unsuccessful the RF thread goes to sleep for a predefined amount of time and wakes up periodically to reattempt transmission.
+
+The system thread always runs in order to receive commands from the SVU. All other threads can be paused from the system thread making it easy to stop logging or streaming of the CAN data.
+
 \subsection{Support Vehicle Unit}
+The SVU is anchored around a Teensy 4.0 microcontroller connected to an RF module.
+The unit receives power and serial data through a micro USB cable connected to a laptop.
+Furthermore, the RF module is bridged via the $\SI{3.3}{V}$ supply and SPI pins.
+A complete circuit diagram for the SVU is shown in Fig.
+\ref{fig:SVU_circuit}.
+
+On the software side, the SVU core executes two threads during operation; a \textit{receiver} thread and a \textit{transmitter} thread are both run on the same priority level with thread switching enabled by the \texttt{yield} method \cite{chibiOsYield}.
+The execution flow of both threads is outlined in Fig.
+\ref{fig:SVU_threads}.
+
+During execution, the receiver thread listens for incoming messages on the RF channel, decodes and relays the message to the TelemetryUI.
+Meanwhile, the transmitter thread waits for serial input from the TelemetryUI to transmit via the RF channel.
+The incoming bytes are interpreted as telemetry messages with a \texttt{cmd} field specifying the command being sent to or received from the SCU.
+Messages are buffered in a FIFO queue to ensure proper transmission.
+
 \begin{figure}
@@
     \caption{The circuit diagram for the SVU.}
-    \label{fig:SCU_circuit}
+    \label{fig:SVU_circuit}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[height=4.5cm]{documentation/images/SVU_threads.pdf}
+    \caption{The execution flow of the 2 threads running in the SVU.}
+    \label{fig:SVU_threads}
 \end{figure}
 % threads
-\subsection{\todo{Graphical User Interface}}
-% GUI
+\subsection{TelemetryUI}
 
-\section{Verification and Validation} % Victor
-As mentioned in the methodology section, we spent time thoroughly verifying and validating our code design.
-We undertook the software verification after finishing the prototype and software coding, reviewing the naming, functionality and commenting the code accordingly.
+The user interface written in scala-swing is divided into three interface components: one presenting received CAN messages, one allowing the user to inject CAN messages into the solar car CAN bus, and finally one containing command buttons controlling the state of the SCU.
+A screenshot of the interface is shown in Fig.
+\ref{fig:gui_screenshot}.
 
-We performed software validation concurrently with writing our code to make sure that the actual behaviour of the program matched the expected behaviour.
+In the background of the program, multithreading is used to catch new messages received via the serial connection to the SVU, which then are broadcasted to UI elements and a separate thread streaming data to the local network via a UDP port.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{documentation/images/swing.png}
+    \caption{A screenshot of the GUI.}
+    \label{fig:gui_screenshot}
+\end{figure}
+
+\section{Testing and Validation} % Victor
+As mentioned in the methodology section, we spent time thoroughly testing and validating our code design.
+Software validation was performed concurrently with writing our code to make sure that the actual behaviour of the program matched the expected behaviour.
 This was done through the testing environment in PlatformIO.
 We used this to validate our code design with unit tests, testing both edge cases and random cases.
-We chose only to validate the code and functions we made ourselves, and not those that was imported from external libraries.
 
@@
 
+Once all the subsystems' functionalities had been validated, testing of the final product began.
+Initially, we were interested in testing the efficacy of our transmission protocol by sending and receiving asynchronous messages between the SCU and SVU.
+Once this was demonstrated, a range test was performed in a suitable outdoor environment.
+In the end, we achieved a transmission range of around 160 m before the connection became too unreliable \cite{rangeTest}.
+
 \section{Discussion}
+\subsection{Range limitation}
+The measured operating range of the two RF modules turned out to be well short of the 400 m that was targeted, let alone the 1000 m claimed by the manufacturer.
+While this was disappointing, there are several indicators for potential range improvement.
 
-% range test
-In order to test the connection between the Solar Car and Support Vehicle modules we performed a range test in a suitable outdoor environment.
-In the end, we achieved a transmission range of around 160 m before the connection became too unreliable.
-A video documenting the test can be viewed with \MYhref{https://youtu.be/-iwST3REn40}{this link}.
-This is obviously a big step down from the 1000 m claimed by the manufacturer.
-However, considering the crude implementation of our hardware components, as well as the interference from buildings and other vehicles, that range was highly unattainable in the first place.
-Further enhancements in the hardware, such as a stable power supply for the RF modules, would likely improve the range beyond 400 m.
+First of all, the hardware setup could be significantly upgraded to ensure a stable power supply to both RF modules, which seemed to be the main error source.
+A PCB implementation would certainly have been a possibility, given its energy-efficiency, but due to the project time frame this was not achievable.
 
-% throughput limitations
-% variable length messages would reduce overhead -> but most messages contain CAN and are optimized
-% UDP works only one way for now. If automated control from matlab should be possible, udp listeners should be able to send as well
+Secondly, time constraints prevented us in testing all of the RF module's many software configurations to enhance the range.
+As such, our RF modules were programmed with the default configuration and therefore not optimized for long range transmission.
+
+\subsection{Throughput limitation}
+When testing the throughput capabilities of the system, it was found that the library code used to interface with the RF modules took up to $\SI{1200}{\mu s}$ to send one message, taking around $\SI{200000}{}$ clock cycles.
+
+Combining the execution time spent on receiving CAN messages and writing to the black box, the SCU spent approximately $\SI{2}{ms}$ on each received CAN message, giving the system a throughput of $\SI{500}{s^{-1}}$ messages falling short of the theoretical $\SI{2659}{s^{-1}}$ messages the CAN bus can handle at peak usage by a large margin.
+
+This shows that there is a lot of performance potential to be gained in optimizing the library code by narrowing it down to only include functionality needed for our use case.
+Additionally, the message protocol used internally can be optimized to reduce the number of bits sent per message, at the cost of increasing the processing workload.
+
+One goal was to allow for automated control from the support vehicle.
+As for now, the CAN data stream is only broadcasted via UDP, but a UDP client can not send CAN messages the other way.
+Since a two way communication is already used for very basic server-client handshaking, this functionality can be integrated into the TelemetryUI relatively easily.
 
 \section{Conclusion}
-%The conclusion goes here.
+The project set out to create a telemetry module for the ROAST solar car and support car, and is successful in most of the points listed in the problem statement.
+The solar car module can read and write data or commands to and from the CAN bus.
+The solution for the processing of data in the support car uses both a GUI and streaming of data directly to Matlab.
+The issue of storing the data locally was solved using a SD card in the Teensy 3.6 as a black box.
+In addition, a feature for securely encrypting and decrypting the message was implemented.
+
+The only features for which the final product does not meet the specification are the maximum reliable transmission distance and the maximum achievable system throughput.
+These issues need further enhancement in the hardware and software, which can be the basis for future projects.
+
+However, considering all of the above, the final product still stands as a satisfactory solution to the issues listed in the problem statement.
 

### 2021-06-20 code/SolarVehicle/lib/rf/RFfunctions.h
tidy up code

@@ -47,8 +47,8 @@ bool RFinit()
     if(!radio.begin()) return false;                            // Begin radio class object, return false if not succeeded. 
     radio.setPALevel(RF24_PA_LOW);                              // Set Power Amplifier level, choose between MIN, LOW, HIGH, MAX (higher PA level improves range)
     radio.setDataRate(RF24_1MBPS);                              // Set Data Rate, choose between RF24_250KBS, RF24_1MBS, RF24_2MBS (higher Data Rates may cause data loss)
-    //radio.setPayloadSize(32);                                   // Set constant payload size to improve transmission time
-    radio.enableDynamicPayloads(); 
+    radio.setPayloadSize(32);                                   // Set constant payload size to improve transmission time
+    //radio.enableDynamicPayloads(); 
     radio.setAutoAck(true);                                     // RX node sends an automatic ack packet
     if (ack) radio.enableAckPayload();
     radio.setRetries(DELAY, COUNT);                             // Sets number of retries and delay between each retry

### 2021-06-20 code/SolarVehicle/src/telemetry/support-vehicle/SystemThds.h

@@ -28,7 +28,7 @@ struct thdBundle
 THD_WORKING_AREA(WaReceiverThd, 2048);
 
 /*****************************************************
- * @brief   Thread function for receiverThd.      *
+ * @brief   Thread function for receiverThd.         *
  * @param   arg, typecast to Security class pointer. *
 ******************************************************/
 THD_FUNCTION(receiverThd, arg)
@@ -40,33 +40,30 @@ THD_FUNCTION(receiverThd, arg)
   chThdSleepMicroseconds(100);                                      // Sleep to allow for creation of secondary thread
 
   while (true){
-    bool flag = false; 
-    if (radio.available()){                                           // Check if TX node is transmitting data
-      if (sec->activate){
-        if (RFreceive(encrypted, BaseTelemetryMsg::length()<<1)){     // Check if message was received
-          sec->decrypt(encrypted, (uint8_t*)&received, 16);           // Decrypt received message
-          flag = true; 
-        } 
-      } else {
-        flag = RFreceive(&received, BaseTelemetryMsg::length()); 
-      }
-      if (flag){
+    if (radio.available()){                                         // Check if TX node is transmitting data
+      if (RFreceive(encrypted, BaseTelemetryMsg::length()<<1)){     // Check if message was received
+        if (sec->activate){                                         // Decrypt received message if encryption has been activated
+          sec->decrypt(encrypted, (uint8_t*)&received, 16);         
+        } else {
+          received.cmd = (cmd_t)encrypted[0];
+          for (int i = 1; i < 16; i++){
+            received.data[i] = (uint8_t)encrypted[i]; 
+          }  
+        }
         Serial.print("Received: ");
-        switch (received.cmd){                                        // Switch based on the message cmd field 
-          case RECEIVED_CAN:{                                         // Standard case, cmd matches RECEIVED_CAN
-            CanTelemetryMsg *ptr = (CanTelemetryMsg*)&received;       // Cast received to CanTelemetryMsg pointer
+        switch (received.cmd){                                      // Switch based on the message cmd field 
+          case RECEIVED_CAN:{                                       // Standard case, cmd matches RECEIVED_CAN
+            CanTelemetryMsg *ptr = (CanTelemetryMsg*)&received;     // Cast received to CanTelemetryMsg pointer
             ptr->toJSON(tempStr, sizeof(tempStr));                  
-            Serial.println(tempStr);                                  // Print message formatted as JSON
+            Serial.println(tempStr);                                // Print message formatted as JSON
           }
             break;
[…]

### 2021-06-20 documentation/FinalPaper/paper.tex

@@
 % 
-\pagestyle{fancy}
+%\pagestyle{fancy}
 \begin{document}
 \pagestyle{default}
-\title{Telemetry Module for the DTU Roadrunners Solar Car}
+\title{Telemetry Module for DTU Roadrunners' Solar Car}
 
@@
 Kgs. Lyngby, Denmark \\
-31015 Introductory project - electrotechnology\\
+31015 Introductory Project - Electrotechnology (Group 7)\\
 s194027@student.dtu.dk, s194006@student.dtu.dk, s186083@student.dtu.dk}}
-%\author{Victor Alexander Hansen, Steffan Martin Kunoy, Tjark Petersen}
 
@@
 All messages are secured with an RSA encryption.
-A graphical user interface enables the support vehicle to stream the CAN data and send commands.
+A graphical user interface enables the user to stream the CAN data and send commands.
 In addition, support for a simple UDP network stream was also implemented which can for instance be accessed in Matlab.
@@
 \ref{fig:schematic}.
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 
-\begin{figure}
+\begin{figure}[t]
     \centering
@@
 The fields of importance for our project is the message identifier, RTR, control field and data field.
-The rest of the fields serve as error checking and coordination of communication on the bus.
+The rest of the fields serve error checking and coordination of communication on the bus.
 
 Each ECU has a unique ID, which the message identifier uses to determine where the CAN message should go to.
-When two messages try to access the bus, the one with the highest priority (lowest ID) will be granted access to the bus, while the other message have to wait.
+When two messages try to access the bus, the one with the highest priority (lowest ID) will be granted access to the bus, while the other message has to wait.
 After the ID comes the RTR bit which determines whether the frame remotely requests or actually sends data to a ECU.
@@
 \subsection{RF network} % Steffan
-Two radio-frequency (RF) transceivers are required to provide a wireless connection between the SCU and SVU.
+Two RF transceivers are required to provide a wireless connection between the SCU and SVU.
 There are many options for radio modules supporting long range communication at the expense of achievable data rate.
@@
 \ref{fig:shockburst}.
-Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and CRC segments are generated automatically by the RF module.
+Only the data payload and address needs to be specified prior to sending a package; the preamble, packet and cyclic redundancy check (CRC) segments are generated automatically by the RF module.
 
-\begin{figure}
+\begin{figure}[h]
     \centering
@@
 The graphical user interface (GUI) included in the telemetry system serves a proof-of-concept purpose.
-Therefore, scala-swing \cite{scala-swing} was chosen as a framework, due to the effectiveness of the scala \cite{scala} language and the high level of abstraction used when describing a user interface in scala-swing.
+Therefore, scala-swing \cite{scala-swing} was chosen as a framework, due to the effectiveness of the scala language \cite{scala} and the high level of abstraction used when describing a user interface in scala-swing.
 Furthermore, a GUI implementation in scala can run on any system with a java virtual machine and is therefore cross-platform compatible.
@@
 
-\begin{figure}
+\begin{figure}[b]
     \centering
@@
 The remaining 15 bytes are interpreted based on the message type.
+All message types can be worked with in an uninterpreted state as a \texttt{BaseTelemetryMsg}.
 As an example, a \texttt{CanTelemetryMsg} message type is shown in Fig.
@@
 The encryption is based on the RSA algorithm \cite{rsa}, which essentially generates an encryption key based on the modulus and totient function of two prime numbers.
-The pairings of prime numbers suited for the encryption protocol can be seen in a table uploaded to DTU ROAST BitBucket \cite{prime}.
+The pairings of prime numbers suited for the encryption protocol can be seen in a table uploaded to the DTU ROAST BitBucket \cite{prime}.
 
@@
 A serial connection of up to $\SI{6}{Mbit/s}$ can be achieved using the Teensy 4.0.
-When sending from the SVU to the TelemetryUI, the message types are converted into JSON strings (see Listing \ref{lst:json}), which are worst case $\SI{117}{bytes}$ long, giving a throughput of $\SI{6410}{s^{-1}}$ messages.
-The JSON string can be directly used to set the state of an object inside scala which then can be used for further processing.
+When sending from the SVU to the TelemetryUI, the message types are converted into JSON \cite{json} strings (see Listing \ref{lst:json}), which are worst case $\SI{117}{bytes}$ long, giving a throughput of $\SI{6410}{s^{-1}}$ messages.
+The JSON string can be directly used to set the state of an object inside scala which then in turn can be used for further processing.
 
@@
 The SCU is built around the Teensy 3.6 and includes the aforementioned CAN transceiver and RF module.
-In the solar car, all modules receive power through a molex connector at $\SI{12}{V}$.
+In the solar car, all modules receive power through a 4-pin molex connector at $\SI{12}{V}$.
 The molex connector also includes the differential pair of wires of the CAN bus.
@@
 
-A real time clock (RTC) is included on the Teensy 3.6 which can keep the time while the Teensy is powered off by connecting a coin cell battery to the its $V_{bat}$ terminal.
+A real time clock (RTC) is included on the Teensy 3.6 which can keep the time while the Teensy is powered off when a coin cell battery is connected to its $V_{bat}$ terminal.
 This ensures that the generated time stamps associated with a received CAN message are always valid, provided the RTC is synchronized.
@@
 \end{figure}
-\begin{figure}
+\begin{figure}[t]
     \centering
@@
 \end{figure}
-% threads
+
 \subsection{TelemetryUI}
@@
 
-In the background of the program, multithreading is used to catch new messages received via the serial connection to the SVU, which then are broadcasted to UI elements and a separate thread streaming data to the local network via a UDP port.
+In the background of the program, multithreading is used to catch new messages received via the serial connection to the SVU, which then are broadcasted to UI elements and spawn a separate thread streaming data to the local network via a UDP port.
 
-\begin{figure}
+\begin{figure}[b]
     \centering
@@
 \subsection{Throughput limitation}
-When testing the throughput capabilities of the system, it was found that the library code used to interface with the RF modules took up to $\SI{1200}{\mu s}$ to send one message, taking around $\SI{200000}{}$ clock cycles.
+When testing the throughput capabilities of the system, it was found that the library code used to interface with the RF modules took up to $\SI{1200}{\mu s}$ to send one message, taking around $\SI{200000}{}$ clock cycles on the Teensy 3.6.
 
@@
 
-This shows that there is a lot of performance potential to be gained in optimizing the library code by narrowing it down to only include functionality needed for our use case.
+This shows that there is a lot of performance potential to be gained in optimizing the library code by narrowing down its functionality and directly working with interrupts.
 Additionally, the message protocol used internally can be optimized to reduce the number of bits sent per message, at the cost of increasing the processing workload.
@@
 \section{Conclusion}
-The project set out to create a telemetry module for the ROAST solar car and support car, and is successful in most of the points listed in the problem statement.
+The project set out to create a telemetry module for the ROAST solar car, and is successful in most of the points listed in the problem statement.
 The solar car module can read and write data or commands to and from the CAN bus.
-The solution for the processing of data in the support car uses both a GUI and streaming of data directly to Matlab.
+The solution for the processing of data in the support vehicle uses both a GUI and streaming of data directly to Matlab.
 The issue of storing the data locally was solved using a SD card in the Teensy 3.6 as a black box.
@@
 
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 The authors would like to thank Martin Schoeberl and Christian Kampp Kruse for their mentoring and assistance throughout the project.
-The authors would also like to thank Claus Suldrup Nielsen from the department of Mechanical Engineering at DTU.
+The authors would also like to thank Claus Suldrup Nielsen and the rest of the DTU Roadrunners team for their outstanding cooperation throughout the project.
 

### 2021-06-21 code/SolarVehicle/lib/message/Encryption.h
Removed debug lines from Encryption

@@ -21,7 +21,7 @@ class Security {
         int t = (pub_key - 1)*(prv_key - 1);                   // Totient function
         int n = pub_key * prv_key;                             // Modulus of prime numbers
         long int i, flag;                             
-        long int e[256], d[256], temp[256];
+        long int e[256], d[256];
         bool activate = false; 
     /**
      * @brief Function to check for prime number
@@ -78,7 +78,9 @@ class Security {
 
         /**
          * @brief Encrypts input message
-         * @param Message as char pointer array
+         * @param message message pointer to be encrypted
+         * @param array outgoing encrypted message pointer
+         * @param len length of message
         */
         void encrypt(uint8_t *message, uint16_t *array, int len){
             long int pt, k, enkey = e[0];
@@ -91,12 +93,7 @@ class Security {
                     k = k * pt;
                     k = k % n;
                 }
-                //temp[i] = k;                       // Array used for encryption and decryption
-                //ct = k + 96;
                 ct = k + 2*96;
-                /*while (ct < 0){
-                    ct += 96;
-                }*/
                 array[i] = ct;
                 i++;
             }
@@ -104,18 +101,16 @@ class Security {
 
         /**
          * @brief Decrypts input message
-         * @param Message as char pointer array
+         * @param array incoming encrypted message pointer
+         * @param message the decrypted message pointer
+         * @param len length of message
         */
         void decrypt(uint16_t *array, uint8_t *message, int len){
             long int ct, dekey = d[0], k;
             uint8_t pt;
             i = 0;
             while(i < len){
-                //ct = temp[i];                      // Array used for encryption and decryption
                 ct = array[i] - 2*96;
-                /*while (ct > 255){
-                    ct -= 96;
-                }*/
                 k = 1;
                 for (int j = 0; j < dekey; j++){
                     k = k * ct;

### 2021-06-21 code/SolarVehicle/test/test_common_support/test_support.cpp
test_support

@@ -17,37 +17,121 @@
 // Maybe use pio --without-building?
 
 #include <unity.h>
+#include <sys/stat.h>
+#include <ctime>
+#include <fstream>
+#include <ChRt.h>
 
-// void setUp(void) {
-// // set stuff up here
-// }
+#include "TelemetryMessages.h"
+#include "Encryption.h"
+#include "RFfunctions.h"
+#include "MultiReaderFifo.h"
+#include "ThreadState.h"
 
-// void tearDown(void) {
-// // clean stuff up here
-// }
+CANMessage frame;
+CanTelemetryMsg CANmsg;
+Security security;
 
-void test_add(void) {
+void test_sanity(void) {
     TEST_ASSERT_EQUAL(32, 25 + 7);
-}
-
-void test_subtract(void) {
     TEST_ASSERT_EQUAL(20, 23 - 3);
-}
-
-void test_multiplication(void) {
     TEST_ASSERT_EQUAL(50, 25 * 2);
+    TEST_ASSERT_EQUAL(32, 96 / 3);
 }
 
-void test_division(void) {
-    TEST_ASSERT_EQUAL(32, 96 / 3);
+// ------------------------------ Encryption.h ----------------------------- //
+
+// Should test that an encryption array from Security class is made
+// Test does not take into account that an encryption key could be all zero
+// void encryption_key()
+void test_encryption_key(void){
+    int cnte, cntd = 0;
+    for (int i = 0; i < 256; i++){
+        if (security.e[i] != 0){
+            cnte++;
+        }
+        if (security.d[i] != 0){
+            cntd++;
+        }
+    }
+    TEST_ASSERT_TRUE(cnte);
+    TEST_ASSERT_TRUE(cntd);
+}
+
+// Should test that an input message is encrypted
+//void encrypt(uint16_t *message, int len)
+//void decrypt(uint16_t *message, int len)
+void test_encrypt_decrypt(void){
+    uint8_t msg[16], msgcon[16];
+    uint16_t array[16];
+    int cnten = 0;
+    security.encryption_key();
+    // Random test
+    for(int j = 0; j < 100; j++){
+        for(int i = 0; i < 16; i++){
+            msg[i] = random(0,256);
+            msgcon[i] = msg[i];
+        }
+        security.encrypt(msg, array, 16);
+        for (int i = 0; i < 16; i++){
+            if (msg[i] != msgcon[i]){
+                cnten++;
+            }
+        }
[…]
